---
title: "Best_models"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Best models

These are the best models for all the traits.

## 1. Derived woodiness

### a. Best model 
```{r}
#construct best model
woody_best_model2<- woody ~ 1 + monthCountByTemp10 + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio5
```

### b. Model optimization 
To optimize the model, the phyloglm function is used. 
```{r}
#load required package
library(phylolm)

#fill in arguments and execute function
model_woody2 <- phyloglm(formula = woody_best_model2, data = GIS_woody, phy = tree_woody, method = "logistic_MPLE", btol = 100, log.alpha.bound = 8)

summary(model_woody2)
```

### c. Final model
```{r}
final_model_woody <- woody ~ 1.78041 - 2.20071*monthCountByTemp10 + 0.28208*PETWettestQuarter + 1.28682*OrganicCarbon - 1.07275*PhCaCL - 0.98976*Slope + 1.48267*bio5
```

### d. Model performance
The model performance package is used, to see whether the model sticks to the assumptions we've made.
```{r}
#load required package 
library(performance)
library(see)
library(gridExtra)

#construct model without phylogeny
model_check_woody <- glm(woody ~ 1 + monthCountByTemp10 + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio5, data = GIS_woody, family = binomial())

#check model
check_model(model_check_woody)
```

## 2. Domestication (staple crops)

### a. Best model formula
```{r}
#construct best model
dom_best_model2 <- domestication ~ 1 + ClayPercentage + PETWarmestQuarter 
```

### b. Model optimization 
To optimize the model, the phyloglm function is used. 
```{r}
#load required package
library(phylolm)

#fill in the arguments and execute it
model_dom2 <- phyloglm(formula = dom_best_model2, data = GIS_dom, phy = tree_dom, method = "logistic_MPLE", btol = 100, log.alpha.bound = 8)

summary(model_dom2) 
```

### c. Final model
```{r}
final_model_dom <- domestication ~ - 1.90666 + 2.32367*ClayPercentage + 2.00246*PETWarmestQuarter 
```

### d. Model performance
The model performance package is used, to see whether the model sticks to the assumptions we've made.
```{r}
#load required package 
library(performance)
library(see)
library(gridExtra)

#construct model without phylogeny
model_check_dom <- glm(domestication ~ 1 + ClayPercentage + PETWarmestQuarter, data = GIS_dom, family = binomial())

#execute function
check_model(model_check_dom)
```

## 3. Mycorrhiza

### a. Best model formula
```{r}
#construct model
myco_best_model <- mycorrhiza ~ 1 + ClayPercentage + bio2 + bio15
```

### b. Model optimization 
To optimize the model, the phyloglm function is used. 
```{r}
#load required package
library(phylolm)

#fill in the arguments and execute it
model_myco <- phyloglm(formula = myco_best_model, data = GIS_myco, phy = tree_myco, method = "logistic_MPLE", btol = 100, log.alpha.bound = 8)

summary(model_myco)
```

### c. Final model
```{r}
final_model_myco <- mycorrhiza ~ - 0.58705 - 0.57211*ClayPercentage -0.64992*bio2 + 0.55083*bio15
```

### d. Model performance
This is needed to see whether the best models we constructed, are still in compliance with the assumptions that have to be met when performing a GLM. There is a package that can do this for you, called 'performance'. We'll do a comprehensive model check, with function 'check_model()'. However, this function is not appropriate for models that include a phylogeny, so in this case, a phyloGLM is not seen as a subclass of GLMs. We will still do it while keeping out the phylogeny, but keep in mind this might result in skewed interpretation. 

```{r}
#install and load required package 
library(performance)
library(see)
library(gridExtra)

#construct model without phylogeny
model_check_myco <- glm(mycorrhiza ~ 1 + ClayPercentage + bio2 + bio15, data = GIS_myco, family = binomial())

#execute function
check_model(model_check_myco)
```

## Compare all the models
Just for fun, see which model performs best. 

```{r}
# Compare models
compare_performance(final_model_woody, final_model_dom, final_model_myco, rank = TRUE)
```

