---
title: "Best_models"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Best models

These are the best models for all the traits.

## 1. Derived woodiness

### a. Best model formula
```{r}
# 24a. With bio5 (for both BM and OU)
woody_formula_24a <- woody ~ Aspect + BulkDensity + ClayPercentage + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio5 + bio14 + bio15 + bio18 + bio19
```

### b. Model performance
The model performance package is used, to see whether the model sticks to the assumptions we've made.
```{r}
#install and load required package 
library(performance)
library(see)

# BM
model_BM_woody <- glm(woody_formula_24a, data = GIS_woody, family = binomial(link = "logit"))
check_model(model_BM_woody)

# OU
model_OU_woody <- glm(woody_formula_24a, data = GIS_woody, family = binomial(link = "logit"))
check_model(model_OU_woody)

# Compare models
compare_performance(model_BM_woody, model_OU_woody)
```

### c. Model optimization 
To optimize the model, the phyloglm function is used. 
```{r}
# BM
phyloglm(formula = woody_formula_24a, data = GIS_woody, phy = tree_woody, method = "logistic_IG10")

#OU
phyloglm(formula = , data = GIS_woody, phy = tree_woody, method = , btol = , log.alpha.bound = , start.beta = , start.alpha = , boot = , full.matrix = )
```

### Conclusion: The best models
```{r}
# 24a. With bio5 for BM
woodiness_BM <- woody ~ 0.83475898 + 0.15198885*BulkDensity -0.48410553*monthCountByTemp10 + 0.08058603*PETWettestQuarter + 0.43981669*OrganicCarbon -0.30052534*PhCaCL -0.19695033*Slope + 0.48659487*bio5 + 0.20864413*bio19

# 24a. With bio5 for OUfixedRoot
woodiness_OU <- woody ~ 0.79869004 -0.32765328*monthCountByTemp10 -0.16620515*PETDriestQuarter + 0.18423198*OrganicCarbon -0.17396725*PhCaCL -0.12842927*Slope + 0.34782999*bio5 -0.08450621*bio14 + 0.07455151*bio19 
```

## 2. Domestication (staple crops)

### a. Best model formula
```{r}
# 11a. With aridityIndexThornthwaite (BM)
crops_formula_11a <- domestication ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + continentality + minTempWarmest + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio19

# 18. With monthCountByTemp10 (OU)
crops_formula_18 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio14 + bio19
```

### b. Model performance
The model performance package is used, to see whether the model sticks to the assumptions we've made.
```{r}
#install and load required package 
library(performance)
library(see)

#BM
model_BM_crops <- glm(crops_formula_11a, data = GIS_crops, family = binomial(link = "logit"))
check_model(model_BM_crops)

#OU
model_OU_crops <- glm(crops_formula_18, data = GIS_crops, family = binomial(link = "logit"))
check_model(model_OU_crops)

# Compare models
compare_performance(model_BM_crops, model_OU_crops, rank = TRUE)
```

### c. Model optimization 
To optimize the model, the phyloglm function is used. 
```{r}
# BM
phyloglm(formula = , data = GIS_crops, phy = tree_crops, method = , btol = , log.alpha.bound = , start.beta = , start.alpha = , boot = , full.matrix = )

#OU
phyloglm(formula = , data = GIS_crops, phy = tree_crops, method = , btol = , log.alpha.bound = , start.beta = , start.alpha = , boot = , full.matrix = )
```

### Conclusion: The best model
```{r}
# BM
domestication_BM <- domestication ~ 0.2718999 + 0.7254957*Aspect + 0.4467063*ClayPercentage + 0.4136109*PhCaCL + 0.2569151*Slope 

# OUfixedRoot
domestication_OU <- domestication ~ 0.2812481 + 0.7397210*Aspect -0.2509282*BulkDensity + 0.4460845*ClayPercentage + 0.3330660*PhCaCL + 0.1744375*Slope
```

## 3. Mycorrhiza

### a. Best model formula
```{r}
# 10c. With growingDegDays5 for BM
myco_formula_10c <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + OrganicCarbon + PhCaCL + Slope + bio3 + bio8 + bio9 + bio14 + bio15 + bio19

# ?. With ... for OUfixedRoot: 

```

### b. Model performance
The model performance package is used, to see whether the model sticks to the assumptions we've made.
```{r}
#install and load required package 
library(performance)
library(see)

# BM
model_BM_myco <- glm(myco_formula_10c, data = GIS_myco, family = binomial(link = "logit"))
check_model(model_BM_myco)

# OU
model_OU_myco <- glm(myco_formula_??, data = GIS_myco, family = binomial(link = "logit"))
check_model(model_OU_myco)

# Compare models
compare_performance(model_BM_myco, model_OU_myco, rank = TRUE)
```

### c. Model optimization 
To optimize the model, the phyloglm function is used. 
```{r}
# BM
phyloglm(formula = , data = GIS_myco, phy = tree_myco, method = , btol = , log.alpha.bound = , start.beta = , start.alpha = , boot = , full.matrix = )

#OU
phyloglm(formula = , data = GIS_myco, phy = tree_myco, method = , btol = , log.alpha.bound = , start.beta = , start.alpha = , boot = , full.matrix = )
```

### Conclusion: The best model

```{r}
# BM
mycorrhiza_BM <- mycorrhiza ~ 0.4401803 -0.1885960*Aspect + 0.1493415*ClayPercentage + 0.4445646*climaticMoistureIndex + 0.1417067*Slope -0.5940963*bio3 -0.5048234*bio14 + 0.2912411*bio19

# OUfixedRoot
mycorrhiza_OU <-
```