---
title: "GIS_data_analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# GIS data analysis

## Load raw GIS data
```{r}
#import raw_means data
project_root <- 'C:/Users/Esther/Documents/Master/Stage/'
raw_means_file <- paste(c(project_root, 'angiosperm-traits/raw_means.csv'), collapse = '')
raw_means <- read.table(raw_means_file, sep = ',', header = TRUE)
```

## Fix names and row names of raw_means
```{r}
#fix names
library(stringi)
raw_names <- as.character(unlist(c(raw_means[1])))
raw_names <- stri_replace_all_charclass(raw_names, "\\p{WHITE_SPACE}", "_")
raw_means$X <- raw_names
```

# Load regular GIS data
```{r import GIS data}
#import GIS data
GIS_file <- paste(c(project_root, 'niche_traits_merged.csv'), collapse = '')
GIS <- read.table(GIS_file, sep = ',', header = TRUE)

#change names of GIS data 
GIS_names <- as.character(unlist(c(GIS[1])))
GIS_names <- gsub('  ', '. ', GIS_names)
GIS_names <- stri_replace_all_charclass(GIS_names, "\\p{WHITE_SPACE}", "_")
#there is a discrepancy with several names that should have a - instead of a _ in some places. 

#change those specific GIS names:
GIS_names[GIS_names == "Acaena_novae_zelandiae"] <- "Acaena_novae-zelandiae"
GIS_names[GIS_names == "Impatiens_noli_tangere"] <- "Impatiens_noli-tangere"  
GIS_names[GIS_names == "Myrrhidendron_donnell_smithii" ] <- "Myrrhidendron_donnell-smithii" 
GIS_names[GIS_names == "Silene_flos_cuculi"] <- "Silene_flos-cuculi"
GIS_names[GIS_names == "Pithecellobium_unguis_cati"] <- "Pithecellobium_unguis-cati"
GIS_names[GIS_names == "Hyacinthoides_non_scripta" ] <- "Hyacinthoides_non-scripta" 
GIS_names[GIS_names == "Alisma_plantago_aquatica"] <- "Alisma_plantago-aquatica" 
GIS_names[GIS_names == "Apera_spica_venti"] <- "Apera_spica-venti"
GIS_names[GIS_names == "Arctostaphylos_uva_ursi"] <- "Arctostaphylos_uva-ursi"
GIS_names[GIS_names == "Blitum_bonus_henricus"] <- "Blitum_bonus-henricus"
GIS_names[GIS_names == "Begonia_meyeri_johannis"] <- "Begonia_meyeri-johannis"

#put the changed names back into the GIS table
GIS$allmb_name <- GIS_names
```

#### Load derived woodiness GIS data
```{r define data root and import tree}
woody_tsv_file <- paste(c(project_root, 'sdmdl-angiosperm-data/data/woody.tsv'), collapse = '')
woody_tsv <- read.table(woody_tsv_file, sep = '\t', header = TRUE)
woody_tsv_bi <- subset(woody_tsv, (woody_tsv$woody == 0 | woody_tsv$woody ==1))
library(stringi)
woody_tsv_bi$allmb_name <- stri_replace_all_charclass(woody_tsv_bi$allmb_name, "\\p{WHITE_SPACE}", "_")
keep_GIS_woody <- as.character(unlist(c(woody_tsv_bi[1])))
GIS_woody <- GIS[GIS$allmb_name %in% keep_GIS_woody,]
GIS_woody_0 <- woody_tsv_bi[woody_tsv_bi$woody == "0",]
GIS_woody_1 <- woody_tsv_bi[woody_tsv_bi$woody == "1",]
GIS_woody$woody[GIS_woody$allmb_name %in% GIS_woody_0$allmb_name] <- "0"  
GIS_woody$woody[GIS_woody$allmb_name %in% GIS_woody_1$allmb_name] <- "1"  
GIS_woody <- GIS_woody[,c(1,43, 2:42)]
```

## The range of woodiness-associated plants for monthCountByTemp10
```{r}
#select the woody taxa that should be kept in the raw_means file
keep_raw_woody <- as.character(unlist(c(woody_tsv_bi[1])))
#define the raw_means file for woodiness
raw_woody <- raw_means[raw_means$X %in% keep_raw_woody,]

#sort the table based on the monthCountByTemp10 variable
```
Highest value: 12
Lowest value: 0,48

## Set species names as row names
```{r}
#set species names as row names
row.names(raw_means) <- raw_means$X
raw_means[,1] <- NULL
```

## Determine highest and lowest value of each variable
```{r}
#function to return the highest value of each column
colMax <- function(data)
  sapply(data, max, na.rm = TRUE)
colMax(raw_means)

#function to return the lowest value of each column
colMin <- function(data)
  sapply(data, min, na.rm = TRUE)
colMin(raw_means)
colnames(raw_means)
```

### Results
Aspect                   1.297   - 4.656
BulkDensity              70.3    - 158.278
ClayPercentage           8.189   - 46.867
annualPET                173.379 - 1975.946
aridityIndexThornthwaite 0       - 99.042
climaticMoistureIndex    -0.989  - 0.718
continentality           0.767   - 40.646
embergerQ                2.379   - 1184.093
growingDegDays0          1073.07 - 117687.109
growingDegDays5          216.757 - 117687.109
maxTempColdest           -255.06 - 309.739
minTempWarmest           -6.68   - 252.095
monthCountByTemp10       0       - 12
PETColdestQuarter        0.119   - 145.757
PETDriestQuarter         3.21    - 203.433
PETseasonality           571.258 - 6616.934
PETWarmestQuarter        44.167  - 220.222
PETWettestQuarter        21.19   - 203.51
thermicityIndex          -428.485- 722.5
OrganicCarbon            0       - 1607.102
PhCaCL                   38.29   - 75.555
Slope                    0.001   - 0.142
bio1                     -138.55 - 272.191
bio2                     37.943  - 173.071
bio3                     15.289  - 88.636
bio4                     234.932 - 14490.064
bio5                     76.385  - 397.342
bio6                     -327.945- 229.162
bio7                     76.943  - 525.92
bio8                     -13.869 - 310.027
bio9                     -254.217- 302.941
bio10                    29.578  - 314.726
bio11                    -281.361- 266.533
bio12                    12.4    - 3503.571
bio13                    3.5     - 496.955
bio14                    0       - 189
bio15                    10.415  - 111.651
bio16                    9       - 1442.636
bio17                    0       - 650.727
bio18                    0       - 1160.182
bio19                    1.455   - 960.065

## Value check 
Are the values in compliance with with the type of plants you would normally see this in? E.g. are the plants with the driest values indeed cacti? To check whether rows and columns are not distorted. 
```{r}
#sort based on drought
raw_means <- raw_means[rev(order(raw_means$aridityIndexThornthwaite)),]
    #2nd plant grows in dry areas

#sort based on continentality
raw_means <- raw_means[rev(order(raw_means$continentality)),]
    #1st plant grows in dry, sandy, and stony places

#sort based on bio5 (Max Temperature of Warmest Month)
raw_means <- raw_means[rev(order(raw_means$bio5)),]
    #1st plant grows in the desert

#sort based on bio10 Mean Temperature of Warmest Quarter
raw_means <- raw_means[rev(order(raw_means$bio10)),]
    #1st grows in dry sandy areas of Australia
```

## Plot OMI-transformed variables against raw values
Plot the variables that are left in the final models, against their raw_means. This should result in a strong linear relationship, that is either negative or positive. This will give us a hint as to whether we can interpret the OMI-transformed variables directly, or we should reverse their coefficient. 

### Woodiness
The variables that are left are: monthCountByTemp10, PETWettestQuarter, OrganicCarbon, PhCaCL, Slope, and bio5.

```{r}
#make allmb_name column into the row names of this table
rownames(GIS_woody) <- GIS_woody[,1]
GIS_woody[,1] <- NULL

#make sure the values in the 'woody' column are seen as numeric values
GIS_woody$woody <- as.numeric(GIS_woody$woody)
#check whether it worked
is.numeric(GIS_woody$woody)
```

#### Trim raw_means for woody
```{r}
#trim data frame
raw_means_woody <- raw_means[row.names(raw_means) %in% keep_GIS_woody,]
combined_woody <- merge(raw_means_woody, GIS_woody, by = 0)
row.names(combined_woody) <- combined_woody$Row.names
combined_woody[,1] <- NULL
```

#### monthCountByTemp10
```{r}
#define two variables to plot
monthCountByTemp10_RAW <- combined_woody$monthCountByTemp10.x
monthCountByTemp10_OMI <- combined_woody$monthCountByTemp10.y

#plot against each other
plot(monthCountByTemp10_OMI, monthCountByTemp10_RAW)
abline(lm(monthCountByTemp10_RAW ~ monthCountByTemp10_OMI))
```

#### PETWettestQuarter
```{r}
#define two variables
PETWettestQuarter_RAW <- combined_woody$PETWettestQuarter.x
PETWettestQuarter_OMI <- combined_woody$PETWettestQuarter.y
#plot against each other
plot(PETWettestQuarter_OMI, PETWettestQuarter_RAW)
abline(lm(PETWettestQuarter_RAW ~ PETWettestQuarter_OMI))
```

#### OrganicCarbon
```{r}
#define two variables
OrganicCarbon_RAW <- combined_woody$OrganicCarbon.x
OrganicCarbon_OMI <- combined_woody$OrganicCarbon.y

#plot against each other
plot(OrganicCarbon_OMI, OrganicCarbon_RAW)
abline(lm(OrganicCarbon_RAW ~ OrganicCarbon_OMI))
```

#### PhCaCL
```{r}
#define two variables
PhCaCL_RAW <- combined_woody$PhCaCL.x
PhCaCL_OMI <- combined_woody$PhCaCL.y

#plot against each other
plot(PhCaCL_OMI, PhCaCL_RAW)
abline(lm(PhCaCL_RAW ~ PhCaCL_OMI))
```

#### Slope
```{r}
#define two variables
Slope_RAW <- combined_woody$Slope.x
Slope_OMI <- combined_woody$Slope.y

#plot against each other
plot(Slope_OMI, Slope_RAW)
abline(lm(Slope_RAW ~ Slope_OMI))
```

#### bio5
```{r}
#define two variables
bio5_RAW <- combined_woody$bio5.x
bio5_OMI <- combined_woody$bio5.y

#plot against each other
plot(bio5_OMI, bio5_RAW)
abline(lm(bio5_RAW ~ bio5_OMI))
```

### Mycorrhiza
The variables that are left are: ClayPercentage, bio2, and bio15.

#### Load mycorrhiza GIS data
```{r}
myco_tsv_file <- paste(c(project_root, 'sdmdl-angiosperm-data/data/myco_taxa.tsv'), collapse = '')
myco_tsv <- read.table(myco_tsv_file, sep = '\t', header = TRUE)
myco_tsv["myco_state_int"] <- NA
myco_tsv$myco_state_int[myco_tsv$myco_state == "arbuscular"] <- "0"
myco_tsv$myco_state_int[myco_tsv$myco_state == "ecto"] <- "1"
myco_tsv$myco_state_int[myco_tsv$myco_state == "both"] <- "1"
myco_tsv$myco_state_int[myco_tsv$myco_state == "neither"] <- "1"
myco_tsv$allmb_name <- stri_replace_all_charclass(myco_tsv$allmb_name, "\\p{WHITE_SPACE}", "_")
keep_GIS_myco <- as.character(unlist(c(myco_tsv[1]))) 
GIS_myco <- GIS[GIS$allmb_name %in% keep_GIS_myco,]
GIS_myco_0 <- myco_tsv[myco_tsv$myco_state_int == "0",]
GIS_myco_1 <- myco_tsv[myco_tsv$myco_state_int == "1",]
GIS_myco$mycorrhiza[GIS_myco$allmb_name %in% GIS_myco_0$allmb_name] <- "0"
GIS_myco$mycorrhiza[GIS_myco$allmb_name %in% GIS_myco_1$allmb_name] <- "1"
GIS_myco <- GIS_myco[,c(1,43, 2:42)]
```

```{r}
#make allmb_name column into the row names of this table
rownames(GIS_myco) <- GIS_myco[,1]
GIS_myco[,1] <- NULL

#make sure the values in the 'woody' column are seen as numeric values
GIS_myco$mycorrhiza <- as.numeric(GIS_myco$mycorrhiza)
#check whether it worked
is.numeric(GIS_myco$mycorrhiza)
```

#### Trim raw_means for myco
```{r}
#trim data frame
raw_means_myco <- raw_means[row.names(raw_means) %in% keep_GIS_myco,]
combined_myco <- merge(raw_means_myco, GIS_myco, by = 0)
row.names(combined_myco) <- combined_myco$Row.names
combined_myco[,1] <- NULL
```

#### ClayPercentage
```{r}
#define two variables
ClayPercentage_RAW_myco <- combined_myco$ClayPercentage.x
ClayPercentage_OMI_myco <- combined_myco$ClayPercentage.y

#plot against each other
plot(ClayPercentage_OMI_myco, ClayPercentage_RAW_myco)
abline(lm(ClayPercentage_RAW_myco ~ ClayPercentage_OMI_myco))
```

#### bio2
```{r}
#define two variables
bio2_RAW <- combined_myco$bio2.x
bio2_OMI <- combined_myco$bio2.y

#plot against each other
plot(bio2_OMI, bio2_RAW)
abline(lm(bio2_RAW ~ bio2_OMI))
```

#### bio15
```{r}
#define two variables
bio15_RAW <- combined_myco$bio15.x
bio15_OMI <- combined_myco$bio15.y

#plot against each other
plot(bio15_OMI, bio15_RAW)
abline(lm(bio15_RAW ~ bio15_OMI))
```

### Domestication
The variables that are left are: ClayPercentage and PETWarmestQuarter.

#### Load domestication GIS data
```{r}
tree_file <- paste(c(project_root, 'ALLMB.tre'), collapse = '')
tree <- read.tree(tree_file)
dom_tsv_file <- paste(c(project_root, 'sdmdl-angiosperm-data/data/crops.tsv'), collapse = '')
dom_tsv <- read.table(dom_tsv_file, sep = '\t', header = TRUE)
dom_1_3 <- subset(dom_tsv, (dom_tsv$domestication == 1 | dom_tsv$domestication == 3))
dom_1_3$domestication_binary[dom_1_3$domestication == "1"] <- "1"
tips_to_drop_GIS <- setdiff(tree$tip.label, GIS_names)
GIS_tree <- drop.tip(tree, tips_to_drop_GIS)
GIS_names[GIS_names == "Zea_mays_subsp._parviglumis"] <- "Zea_mays"
GIS$allmb_name <- GIS_names
dom_1_3$allmb_name[dom_1_3$allmb_name == "Zea_mays_subsp._parviglumis"] <- "Zea_mays"
keep_tips_dom_1_3 <- as.character(unlist(c(dom_1_3[1]))) 
tips_to_drop_dom <- setdiff(GIS_tree$tip.label, keep_tips_dom_1_3)
tree_dom_tsv <- drop.tip(GIS_tree, tips_to_drop_dom)
dom_names <- as.character(unlist(c(tree_dom_tsv$tip.label)))
allmb_name <- c("Euphorbia_usambarica", "Calycophysum_pedunculatum", "Ibervillea_sonorae", 
                "Psiguria_triphylla", "Gurania_eriantha", "Doyerea_emetocathartica", 
                "Ficus_pandurata", "Fissicalyx_fendleri", "Dendrosida_sharpiana_subsp._pubescens", 
                "Ipomoea_arborescens", "Heliconia_psittacorum", "Tradescantia_zebrina", 
                "Amischotolype_hispida", "Scleria_gaertneri", "Festuca_arundinacea_subsp._orientalis", 
                "Pogonarthria_squarrosa", "Digitaria_ischaemum", "Digitaria_violascens", 
                "Elymus_repen", "Aegilops_cylindrica", "Aegilops_crassa", "Elymus_elongatus", 
                "Poa_colensoi", "Molineria_capitulata", "Polygonatum_verticillatum", 
                "Epipactis_helleborine", "Evodianthus_funifer", "Schoenoplectus_pungens_var._badius", 
                "Helictotrichon_desertorum_subsp._basalticum","Eleocharis_acutangula_subsp._breviseta",
                "Raddiella_esenbeckii", "Dienia_ophrydis", "Arum_maculatum", "Calla_palustris")
dom_state0 <- data.frame(allmb_name)
dom_state0$gbif_species_key <- "NA"
dom_state0$domestication <- "0"
dom_state0$progenitor_of <- "NA"
dom_state0$domestication_binary <- "0"
dom_bi <- rbind(dom_1_3, dom_state0)  
GIS_dom_1 <- dom_1_3[dom_1_3$allmb_name %in% GIS$allmb_name,]
GIS_dom_0 <- dom_state0[dom_state0$allmb_name %in% GIS$allmb_name,]
keep_GIS_dom <- as.character(unlist(c(dom_bi[1]))) 
GIS_dom <- GIS[GIS$allmb_name %in% keep_GIS_dom,]
GIS_dom$domestication[GIS_dom$allmb_name %in% GIS_dom_1$allmb_name] <- "1"
GIS_dom$domestication[GIS_dom$allmb_name %in% GIS_dom_0$allmb_name] <- "0"
GIS_dom <- GIS_dom[,c(1,43, 2:42)]
```

```{r}
#make allmb_name column into the row names of this table
rownames(GIS_dom) <- GIS_dom[,1]
GIS_dom[,1] <- NULL

#make sure the values in the 'woody' column are seen as numeric values
GIS_dom$domestication <- as.numeric(GIS_dom$domestication)
#check whether it worked
is.numeric(GIS_dom$domestication)
```

#### Trim raw_means for domestication
```{r}
#trim data frame
raw_means_dom <- raw_means[row.names(raw_means) %in% keep_GIS_dom,]
combined_dom <- merge(raw_means_dom, GIS_dom, by = 0)
row.names(combined_dom) <- combined_dom$Row.names
combined_dom[,1] <- NULL
```

#### ClayPercentage
```{r}
#define two variables
ClayPercentage_RAW_dom <- combined_dom$ClayPercentage.x
ClayPercentage_OMI_dom <- combined_dom$ClayPercentage.y

#plot against each other
plot(ClayPercentage_OMI_dom, ClayPercentage_RAW_dom)
abline(lm(ClayPercentage_RAW_dom ~ ClayPercentage_OMI_dom))
```

#### PETWarmestQuarter
```{r}
#define two variables
PETWarmestQuarter_RAW <- combined_dom$PETWarmestQuarter.x
PETWarmestQuarter_OMI <- combined_dom$PETWarmestQuarter.y

#plot against each other
plot(PETWarmestQuarter_OMI, PETWarmestQuarter_RAW)
abline(lm(PETWarmestQuarter_RAW ~ PETWarmestQuarter_OMI))
```
