---
title: "Interaction_analysis_OMI"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load hypervolume package
```{r}
library(hypervolume)
library(dplyr)
library(alphahull)
library(stringi)
```

# Load GIS data (OMI)
```{r import GIS data}
#import GIS data
GIS_file <- paste(c(project_root, 'niche_traits_merged.csv'), collapse = '')
GIS <- read.table(GIS_file, sep = ',', header = TRUE)

#change names of GIS data 
GIS_names <- as.character(unlist(c(GIS[1])))
GIS_names <- gsub('  ', '. ', GIS_names)
GIS_names <- stri_replace_all_charclass(GIS_names, "\\p{WHITE_SPACE}", "_")
#there is a discrepancy with several names that should have a - instead of a _ in some places. 

#change those specific GIS names:
GIS_names[GIS_names == "Acaena_novae_zelandiae"] <- "Acaena_novae-zelandiae"
GIS_names[GIS_names == "Impatiens_noli_tangere"] <- "Impatiens_noli-tangere"  
GIS_names[GIS_names == "Myrrhidendron_donnell_smithii" ] <- "Myrrhidendron_donnell-smithii" 
GIS_names[GIS_names == "Silene_flos_cuculi"] <- "Silene_flos-cuculi"
GIS_names[GIS_names == "Pithecellobium_unguis_cati"] <- "Pithecellobium_unguis-cati"
GIS_names[GIS_names == "Hyacinthoides_non_scripta" ] <- "Hyacinthoides_non-scripta" 
GIS_names[GIS_names == "Alisma_plantago_aquatica"] <- "Alisma_plantago-aquatica" 
GIS_names[GIS_names == "Apera_spica_venti"] <- "Apera_spica-venti"
GIS_names[GIS_names == "Arctostaphylos_uva_ursi"] <- "Arctostaphylos_uva-ursi"
GIS_names[GIS_names == "Blitum_bonus_henricus"] <- "Blitum_bonus-henricus"
GIS_names[GIS_names == "Begonia_meyeri_johannis"] <- "Begonia_meyeri-johannis"

#put the changed names back into the GIS table
GIS$allmb_name <- GIS_names
```

## Load mycorrhiza GIS data
```{r}
myco_tsv_file <- paste(c(project_root, 'sdmdl-angiosperm-data/data/myco_taxa.tsv'), collapse = '')
myco_tsv <- read.table(myco_tsv_file, sep = '\t', header = TRUE)
myco_tsv["myco_state_int"] <- NA
myco_tsv$myco_state_int[myco_tsv$myco_state == "arbuscular"] <- "0"
myco_tsv$myco_state_int[myco_tsv$myco_state == "ecto"] <- "1"
myco_tsv$myco_state_int[myco_tsv$myco_state == "both"] <- "1"
myco_tsv$myco_state_int[myco_tsv$myco_state == "neither"] <- "1"
myco_tsv$allmb_name <- stri_replace_all_charclass(myco_tsv$allmb_name, "\\p{WHITE_SPACE}", "_")
keep_GIS_myco <- as.character(unlist(c(myco_tsv[1]))) 
GIS_myco <- GIS[GIS$allmb_name %in% keep_GIS_myco,]
GIS_myco_0 <- myco_tsv[myco_tsv$myco_state_int == "0",]
GIS_myco_1 <- myco_tsv[myco_tsv$myco_state_int == "1",]
GIS_myco$mycorrhiza[GIS_myco$allmb_name %in% GIS_myco_0$allmb_name] <- "0"
GIS_myco$mycorrhiza[GIS_myco$allmb_name %in% GIS_myco_1$allmb_name] <- "1"
GIS_myco <- GIS_myco[,c(1,43, 2:42)]
```

```{r}
#define two subsets
OMI_AM <- GIS_myco %>%
    subset(mycorrhiza == 0) %>%
    select(ClayPercentage, bio2, bio15) 
OMI_EM <- GIS_myco %>%
    subset(mycorrhiza == 1) %>%
    select(ClayPercentage, bio2, bio15) 
OMI_EM <- na.omit(OMI_EM)

# calculate hypervolume
hv.AM <- hypervolume(OMI_AM, method = 'svm', name = 'Arbuscular mycorrhiza', svm.nu = 0.001)
hv.EM <- hypervolume(OMI_EM, method = 'svm', name = 'Ectomycorrhiza, both, and neither',  svm.nu = 0.001)
hv.list.myco <- hypervolume_join(hv.AM,hv.EM)
plot(hv.list.myco, show.3d = F, cex.names = 1.5, cex.legend = 2)
```

## Load domestication GIS data
```{r}
dom_tsv_file <- paste(c(project_root, 'sdmdl-angiosperm-data/data/crops.tsv'), collapse = '')
dom_tsv <- read.table(dom_tsv_file, sep = '\t', header = TRUE)
dom_1_3 <- subset(dom_tsv, (dom_tsv$domestication == 1 | dom_tsv$domestication == 3))
dom_1_3$domestication_binary[dom_1_3$domestication == "1"] <- "1"
tips_to_drop_GIS <- setdiff(tree$tip.label, GIS_names)
GIS_tree <- drop.tip(tree, tips_to_drop_GIS)
GIS_names[GIS_names == "Zea_mays_subsp._parviglumis"] <- "Zea_mays"
GIS$allmb_name <- GIS_names
dom_1_3$allmb_name[dom_1_3$allmb_name == "Zea_mays_subsp._parviglumis"] <- "Zea_mays"
keep_tips_dom_1_3 <- as.character(unlist(c(dom_1_3[1]))) 
tips_to_drop_dom <- setdiff(GIS_tree$tip.label, keep_tips_dom_1_3)
tree_dom_tsv <- drop.tip(GIS_tree, tips_to_drop_dom)
dom_names <- as.character(unlist(c(tree_dom_tsv$tip.label)))
allmb_name <- c("Euphorbia_usambarica", "Calycophysum_pedunculatum", "Ibervillea_sonorae", 
                "Psiguria_triphylla", "Gurania_eriantha", "Doyerea_emetocathartica", 
                "Ficus_pandurata", "Fissicalyx_fendleri", "Dendrosida_sharpiana_subsp._pubescens", 
                "Ipomoea_arborescens", "Heliconia_psittacorum", "Tradescantia_zebrina", 
                "Amischotolype_hispida", "Scleria_gaertneri", "Festuca_arundinacea_subsp._orientalis", 
                "Pogonarthria_squarrosa", "Digitaria_ischaemum", "Digitaria_violascens", 
                "Elymus_repen", "Aegilops_cylindrica", "Aegilops_crassa", "Elymus_elongatus", 
                "Poa_colensoi", "Molineria_capitulata", "Polygonatum_verticillatum", 
                "Epipactis_helleborine", "Evodianthus_funifer", "Schoenoplectus_pungens_var._badius", 
                "Helictotrichon_desertorum_subsp._basalticum","Eleocharis_acutangula_subsp._breviseta",
                "Raddiella_esenbeckii", "Dienia_ophrydis", "Arum_maculatum", "Calla_palustris")
dom_state0 <- data.frame(allmb_name)
dom_state0$gbif_species_key <- "NA"
dom_state0$domestication <- "0"
dom_state0$progenitor_of <- "NA"
dom_state0$domestication_binary <- "0"
dom_bi <- rbind(dom_1_3, dom_state0)  
GIS_dom_1 <- dom_1_3[dom_1_3$allmb_name %in% GIS$allmb_name,]
GIS_dom_0 <- dom_state0[dom_state0$allmb_name %in% GIS$allmb_name,]
keep_GIS_dom <- as.character(unlist(c(dom_bi[1]))) 
GIS_dom <- GIS[GIS$allmb_name %in% keep_GIS_dom,]
GIS_dom$domestication[GIS_dom$allmb_name %in% GIS_dom_1$allmb_name] <- "1"
GIS_dom$domestication[GIS_dom$allmb_name %in% GIS_dom_0$allmb_name] <- "0"
GIS_dom <- GIS_dom[,c(1,43, 2:42)]
```

### Domestication OMI means GIS data boxplots
```{r boxplots}
#define two subsets
OMI_non_predom <- GIS_dom %>%
    subset(domestication == 0) %>%
    select(ClayPercentage, PETWarmestQuarter)
OMI_predom <- GIS_dom %>%
    subset(domestication == 1) %>%
    select(ClayPercentage, PETWarmestQuarter) 

# calculate hypervolume
hv.non.predom <- hypervolume(OMI_non_predom, method = 'svm', name = 'Non domesticated', svm.nu = 0.001)
hv.predom <- hypervolume(OMI_predom, method = 'svm', name = 'Pre-adapted for domesticated',  svm.nu = 0.001)
hv.list.dom <- hypervolume_join(hv.non.predom,hv.predom)
dom.2D <- plot(hv.list.dom, show.3d = F, cex.names = 1.5, cex.legend = 2)
```

## Load derived woodiness GIS data
```{r define data root and import tree}
woody_tsv_file <- paste(c(project_root, 'sdmdl-angiosperm-data/data/woody.tsv'), collapse = '')
woody_tsv <- read.table(woody_tsv_file, sep = '\t', header = TRUE)
woody_tsv_bi <- subset(woody_tsv, (woody_tsv$woody == 0 | woody_tsv$woody ==1))
library(stringi)
woody_tsv_bi$allmb_name <- stri_replace_all_charclass(woody_tsv_bi$allmb_name, "\\p{WHITE_SPACE}", "_")
keep_GIS_woody <- as.character(unlist(c(woody_tsv_bi[1])))
GIS_woody <- GIS[GIS$allmb_name %in% keep_GIS_woody,]
GIS_woody_0 <- woody_tsv_bi[woody_tsv_bi$woody == "0",]
GIS_woody_1 <- woody_tsv_bi[woody_tsv_bi$woody == "1",]
GIS_woody$woody[GIS_woody$allmb_name %in% GIS_woody_0$allmb_name] <- "0"  
GIS_woody$woody[GIS_woody$allmb_name %in% GIS_woody_1$allmb_name] <- "1"  
GIS_woody <- GIS_woody[,c(1,43, 2:42)]
```

### Set 1: the three most significant variables
```{r boxplots}
#define two subsets
OMI_non_woody1 <- GIS_woody %>%
    subset(woody == 0) %>% 
    select(monthCountByTemp10, PhCaCL, Slope)
OMI_non_woody1 <- na.omit(OMI_non_woody1)
OMI_woody1 <- GIS_woody %>%
    subset(woody == 1) %>%
    select(monthCountByTemp10, PhCaCL, Slope) 

# calculate hypervolume
hv.non.woody1 <- hypervolume(OMI_non_woody1, method = 'svm', name = 'Herbaceous', svm.nu = 0.001)
hv.woody1 <- hypervolume(OMI_woody1, method = 'svm', name = 'Derived woody',  svm.nu = 0.001)

hv.list.woody1 <- hypervolume_join(hv.non.woody1,hv.woody1)
plot(hv.list.woody1, show.3d = F, cex.names = 1.5, cex.legend = 2)
```

### Set 2: the other three variables
```{r boxplots}
#define two subsets
OMI_non_woody2 <- GIS_woody %>%
    subset(woody == 0) %>% 
    select(PETWettestQuarter, OrganicCarbon, bio5)
OMI_non_woody2 <- na.omit(OMI_non_woody2)
OMI_woody2 <- GIS_woody %>%
    subset(woody == 1) %>%
    select(PETWettestQuarter, OrganicCarbon, bio5) 

# calculate hypervolume
hv.non.woody2 <- hypervolume(OMI_non_woody2, method = 'svm', name = 'Herbaceous', svm.nu = 0.001)
hv.woody2 <- hypervolume(OMI_woody2, method = 'svm', name = 'Derived woody',  svm.nu = 0.001)

hv.list.woody2 <- hypervolume_join(hv.non.woody2,hv.woody2)
plot(hv.list.woody2, show.3d = F, cex.names = 1.5, cex.legend = 2)
```

### Set 3: monthCountByTemp10 vs OrganicCarbon/bio5
```{r boxplots}
#define two subsets
OMI_non_woody3 <- GIS_woody %>%
    subset(woody == 0) %>% 
    select(monthCountByTemp10, OrganicCarbon, bio5)
OMI_non_woody3 <- na.omit(OMI_non_woody3)
OMI_woody3 <- GIS_woody %>%
    subset(woody == 1) %>%
    select(monthCountByTemp10, OrganicCarbon, bio5) 

# calculate hypervolume
hv.non.woody3 <- hypervolume(OMI_non_woody3, method = 'svm', name = 'Herbaceous', svm.nu = 0.001)
hv.woody3 <- hypervolume(OMI_woody3, method = 'svm', name = 'Derived woody',  svm.nu = 0.001)

hv.list.woody3 <- hypervolume_join(hv.non.woody3,hv.woody3)
plot(hv.list.woody3, show.3d = F, cex.names = 1.5, cex.legend = 2)
```

### Set 4: PhCaCL vs OrganicCarbon/bio5
```{r boxplots}
#define two subsets
OMI_non_woody4 <- GIS_woody %>%
    subset(woody == 0) %>% 
    select(PhCaCL, OrganicCarbon, bio5)
OMI_non_woody4 <- na.omit(OMI_non_woody4)
OMI_woody4 <- GIS_woody %>%
    subset(woody == 1) %>%
    select(PhCaCL, OrganicCarbon, bio5) 

# calculate hypervolume
hv.non.woody4 <- hypervolume(OMI_non_woody4, method = 'svm', name = 'Herbaceous', svm.nu = 0.001)
hv.woody4 <- hypervolume(OMI_woody4, method = 'svm', name = 'Derived woody',  svm.nu = 0.001)

hv.list.woody4 <- hypervolume_join(hv.non.woody4,hv.woody4)
plot(hv.list.woody4, show.3d = F, cex.names = 1.5, cex.legend = 2)
```

### Set 5: Slope vs OrganicCarbon/bio5
```{r boxplots}
#define two subsets
OMI_non_woody5 <- GIS_woody %>%
    subset(woody == 0) %>% 
    select(Slope, OrganicCarbon, bio5)
OMI_non_woody5 <- na.omit(OMI_non_woody5)
OMI_woody5 <- GIS_woody %>%
    subset(woody == 1) %>%
    select(Slope, OrganicCarbon, bio5) 

# calculate hypervolume
hv.non.woody5 <- hypervolume(OMI_non_woody5, method = 'svm', name = 'Herbaceous', svm.nu = 0.001)
hv.woody5 <- hypervolume(OMI_woody5, method = 'svm', name = 'Derived woody',  svm.nu = 0.001)

hv.list.woody5 <- hypervolume_join(hv.non.woody5,hv.woody5)
plot(hv.list.woody5, show.3d = F, cex.names = 1.5, cex.legend = 2)
```

### Set 6: PETWettestQuarter vs PhCaCL/Slope
```{r boxplots}
#define two subsets
OMI_non_woody6 <- GIS_woody %>%
    subset(woody == 0) %>% 
    select(PETWettestQuarter, PhCaCL, Slope)
OMI_non_woody6 <- na.omit(OMI_non_woody6)
OMI_woody6 <- GIS_woody %>%
    subset(woody == 1) %>%
    select(PETWettestQuarter, PhCaCL, Slope) 

# calculate hypervolume
hv.non.woody6 <- hypervolume(OMI_non_woody6, method = 'svm', name = 'Herbaceous', svm.nu = 0.001)
hv.woody6 <- hypervolume(OMI_woody6, method = 'svm', name = 'Derived woody',  svm.nu = 0.001)

hv.list.woody6 <- hypervolume_join(hv.non.woody6,hv.woody6)
plot(hv.list.woody6, show.3d = F, cex.names = 1.5, cex.legend = 2)
```

### Set 7: PETWettestQuarter vs monthCountByTemp10
```{r boxplots}
#define two subsets
OMI_non_woody7 <- GIS_woody %>%
    subset(woody == 0) %>% 
    select(PETWettestQuarter, monthCountByTemp10)
OMI_non_woody7 <- na.omit(OMI_non_woody7)
OMI_woody7 <- GIS_woody %>%
    subset(woody == 1) %>%
    select(PETWettestQuarter, monthCountByTemp10) 

# calculate hypervolume
hv.non.woody7 <- hypervolume(OMI_non_woody7, method = 'svm', name = 'Herbaceous', svm.nu = 0.001)
hv.woody7 <- hypervolume(OMI_woody7, method = 'svm', name = 'Derived woody',  svm.nu = 0.001)

hv.list.woody7 <- hypervolume_join(hv.non.woody7,hv.woody7)
plot(hv.list.woody7, show.3d = F, cex.names = 1.5, cex.legend = 2)
```
