---
title: "Model_selection_crops_elaborated"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Model selection domestication

The linear model will have to be made, with the trait-presence-column as the dependent variable, and the 
GIS data frames as the predictor variables. The function phylostep of the phylolm package will be used. 

## Define formulas 

### First set of model formuals
Based on the hypothesis. 

```{r}
# 0. Starting formula
crops_starting_formula <- domestication ~ 1 

# 1. Lowest VIF value
crops_formula_1 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio14 + bio19

# 2. With PETseasonality
crops_formula_2 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio16 + bio19

# 3. With BIO2
crops_formula_3 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio13 + bio14 + bio19

# 4. With BIO4
crops_formula_4a <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio4 + bio13 + bio15 + bio19
crops_formula_4b <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio15 + bio16 + bio19
crops_formula_4c <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio4 + bio15 + bio18 + bio19

# 5. BIO15
crops_formula_5a <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio7 + bio13 + bio15
crops_formula_5b <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio15 + bio19

# 6. With PETseasonality and BIO2
crops_formula_6 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio16 + bio19

# 7. With PETseasonality and BIO4
crops_formula_7a <- domestication ~ Aspect + ClayPercentage + minTempWarmest + PETseasonality + PETWettestQuarter + OrganicCarbon + Slope + bio4 + bio14 + bio19
crops_formula_7b <- domestication ~ Aspect + ClayPercentage + PETseasonality + PETWettestQuarter + OrganicCarbon + Slope + bio4 + bio10 + bio19
crops_formula_7c <- domestication ~ Aspect + ClayPercentage + PETseasonality + PETWettestQuarter + OrganicCarbon + Slope + bio4 + bio17 + bio19

# 8. With Bio15 and BIO2
crops_formula_8a <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio15 + bio16
crops_formula_8b <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio15 + bio18
```

### Second set of model formulas 
The rest of the variables that were not included yet. 

```{r}
# 9. With annualPET (6)
crops_formula_9 <- domestication ~ Aspect + BulkDensity + ClayPercentage + annualPET + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio14 + bio19

# 10. With aridityIndexThornthwaite 
crops_formula_10a <- domestication ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + continentality + minTempWarmest + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio19
crops_formula_10b <- domestication ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + maxTempColdest + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio19
crops_formula_10c <- domestication ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio19
crops_formula_10d <- domestication ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio7 + bio13 + bio19
crops_formula_10e <- domestication ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio4 + bio13 + bio19
crops_formula_10f <- domestication ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio13 + bio19
crops_formula_10g <- domestication ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio9 + bio13 + bio19
crops_formula_10h <- domestication ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio11 + bio13 + bio19

# 11. With climaticMoistureIndex 
crops_formula_11 <- domestication ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + minTempWarmest + PETDriestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio19

# 12. With continentality
crops_formula_12 <- domestication ~ Aspect + BulkDensity + ClayPercentage + continentality + minTempWarmest + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio14 + bio19

# 13. With embergerQ 
crops_formula_13a <- domestication ~ Aspect + BulkDensity + ClayPercentage + embergerQ + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio18
crops_formula_13b <- domestication ~ Aspect + BulkDensity + ClayPercentage + continentality + embergerQ + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope
crops_formula_13c <- domestication ~ Aspect + BulkDensity + ClayPercentage + embergerQ + minTempWarmest + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope
crops_formula_13d <- domestication ~ Aspect + BulkDensity + ClayPercentage + embergerQ + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio4
crops_formula_13e <- domestication ~ Aspect + BulkDensity + ClayPercentage + embergerQ + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio7
crops_formula_13f <- domestication ~ Aspect + BulkDensity + ClayPercentage + embergerQ + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13
crops_formula_13g <- domestication ~ Aspect + BulkDensity + ClayPercentage + embergerQ + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15
crops_formula_13h <- domestication ~ Aspect + BulkDensity + ClayPercentage + embergerQ + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio16

# 14. With growingDegDays0 
crops_formula_14a <- domestication ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + growingDegDays0 + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio19
crops_formula_14b <- domestication ~ Aspect + BulkDensity + ClayPercentage + growingDegDays0 + minTempWarmest + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio19
crops_formula_14c <- domestication ~ Aspect + BulkDensity + ClayPercentage + growingDegDays0 + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio19

# 15. With maxTempColdestMonth 
crops_formula_15 <- domestication ~ Aspect + BulkDensity + ClayPercentage + maxTempColdest + minTempWarmest + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio14 + bio19

# 16. With PETColdestQuarter 
crops_formula_16 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETColdestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio14 + bio19

# 17. With PETWarmestQuarter 
crops_formula_17 <- domestication ~ Aspect + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14

# 18. With thermicityIndex
crops_formula_18a <- domestication ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + minTempWarmest + PETWettestQuarter + thermicityIndex + OrganicCarbon + PhCaCL + Slope + bio14 + bio19
crops_formula_18b <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWarmestQuarter + PETWettestQuarter + thermicityIndex + OrganicCarbon + PhCaCL + Slope + bio14 + bio19
crops_formula_18c <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWettestQuarter + thermicityIndex + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio19

# 19. With bio1
crops_formula_19a <- domestication ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio1 + bio14 + bio19
crops_formula_19b <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio1 + bio14 + bio19
crops_formula_19c <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio1 + bio2 + bio14 + bio19

# 20. With bio3 
crops_formula_20a <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio3 + bio13 + bio14 + bio19
crops_formula_20b <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio3 + bio13 + bio14 + bio19

# 21. With bio5 
crops_formula_21 <- domestication ~ Aspect + BulkDensity + ClayPercentage + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio5 + bio13 + bio14 + bio19

# 22. With bio6
crops_formula_22 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6  + bio13 + bio14 + bio19

# 23. With bio7 
crops_formula_23 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio7 + bio14 + bio19

# 24. With bio8 
crops_formula_24 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio8 + bio13 + bio14 + bio19

# 25. With bio9 
crops_formula_25 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio9 + bio13 + bio14 + bio19

# 26. With bio10 
crops_formula_26 <- domestication ~ Aspect + BulkDensity + ClayPercentage + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio10 + bio13  + bio14 + bio19

# 27. With bio11
crops_formula_27 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio11  + bio13 + bio14 + bio19

# 28. With bio12 
crops_formula_28 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio12 + bio19

# 29. With bio16 
crops_formula_29 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio16 + bio19

# 30. With bio17
crops_formula_30 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio13 + bio17

# 31. With bio18 
crops_formula_31a <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio18
crops_formula_31b <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio18

# 32. With things 
crops_formula_32 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio16 + bio19

# 33. With BIO2, BIO4 and BIO15
crops_formula_33 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio15 + bio16 + bio19

# 34. With growingDegDays5
crops_formula_34a <- domestication ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + growingDegDays5 + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio19
crops_formula_34b <- domestication ~ Aspect + BulkDensity + ClayPercentage + growingDegDays5 + minTempWarmest + PETWarmestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio19
crops_formula_34c <- domestication ~ Aspect + BulkDensity + ClayPercentage + growingDegDays5 + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio19

# 35. With monthCountByTemp10
crops_formula_35 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio14 + bio19

# 36. PETWettestQuarter
crops_formula_36 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio13 + bio14 + bio19
```

## Change data to match the wanted data format

There are some flaws in the GIS data frame that make it unfit for the phylostep function. First of all
the row names have to match the tip.lables of the tree, and secondly the trait value (0/1) has to be
seen as a numeric value. 
```{r}
#make allmb_name column into the row names of this table
rownames(GIS_crops) <- GIS_crops[,1]
GIS_crops[,1] <- NULL

#make sure the values in the 'woody' column are seen as numeric values
GIS_crops$domestication <- as.numeric(GIS_crops$domestication)
#check whether it worked
is.numeric(GIS_crops$domestication)
```

## Preform model selection with BM
BM stands for Brownian motion. 

```{r model selection BM}
#load required package
library(phylolm)

#define formula components and run function (forward)
phylostep(formula = crops_formula_8a, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "BM", direction = "forward", trace = 2)
  
#define formula components and run function (both)
phylostep(formula = crops_formula_8a, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "BM", direction = "both", trace = 2)
  
```

## Preform model selection with OUrandomRoot
OUrandomRoot stands for Ornstein-Uhlenbeck with a random root. 

```{r model selection OUrandomRoot}
#define formula components and run function (forward)
phylostep(formula = crops_formula_10h, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "OUrandomRoot", direction = "forward", trace = 2, lower.bound = 0.001, upper.bound = 1000)

#define formula components and run function (both)
phylostep(formula = crops_formula_31a, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "OUrandomRoot", direction = "both", trace = 2, lower.bound = 0.001, upper.bound = 1000)

```

## Preform model selection with OUfixedRoot
OUfixedRoot stands for Ornstein-Uhlenbeck with a fixed root. 

```{r model selection OUfixedRoot}
#define formula components and run function (forward)
phylostep(formula = crops_formula_8a, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "OUfixedRoot", direction = "forward", trace = 2, lower.bound = 0.001, upper.bound = 100)

#define formula components and run function (both)
phylostep(formula = crops_formula_20b, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "OUfixedRoot", direction = "both", trace = 2, lower.bound = 0.001, upper.bound = 1000)
```


