---
title: "Model_selection_woody"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Model selection derived woodiness

The linear model will have to be made, with the trait-presence-column as the dependent variable, and the 
GIS data frames as the predictor variables. The function phylostep of the phylolm package will be used. 

## Results VIF analysis

```{r VIF analysis}
# 1. Without ClimaticMoistureIndex: 
vif(GIS_woody[,c(3:5, 14, 15, 17, 18, 20, 22:24, 38, 39, 42, 43)])

# 2. With ClimaticMoistureIndex:
vif(GIS_woody[,c(3:5, 8, 14, 17, 18, 20, 22:24, 39, 43)])

# 3. With BIO6, without ClimaticMoistureIndex:
vif(GIS_woody[,c(3:5, 14, 15, 17, 20, 22:24, 30, 38, 39, 42, 43)]) 

# 4. With BIO6, with ClimaticMoistureIndex:
vif(GIS_woody[,c(3:5, 8, 14, 17, 20, 22:24, 30, 39, 43)]) 
```

## Change data to match the wanted data format

There are some flaws in the GIS data frame that make it unfit for the phylostep function. First of all
the row names have to match the tip.lables of the tree, and secondly the trait value (0/1) has to be
seen as a numeric value. 
```{r}
#make allmb_name column into the row names of this table
rownames(GIS_woody) <- GIS_woody[,1]
GIS_woody[,1] <- NULL

#make sure the values in the 'woody' column are seen as numeric values
GIS_woody$woody <- as.numeric(GIS_woody$woody)
#check whether it worked
is.numeric(GIS_woody$woody)
```

## Define formulas

```{r define formula}
# 0. Define starting formula
woody_starting_formula <- woody ~ 1 

# 1. Without ClimaticMoistureIndex: 
woody_formula_1 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19

# 2. With ClimaticMoistureIndex:
woody_formula_2 <- woody ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + minTempWarmest + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio19

# 3. With BIO6, without ClimaticMoistureIndex:
woody_formula_3 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio14 + bio15 + bio18 + bio19

# 4. With BIO6, with ClimaticMoistureIndex:
woody_formula_4 <- woody ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio15 + bio19
```

## Model selection with BM 

BM stands for Brownian motion. 

```{r model selection BM}
#load required package
library(phylolm)

#define formula components and run function (forward)
##1
phylostep(formula = woody_formula_1, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "BM", direction = "forward", trace = 2)
  # woody ~ 1 + BulkDensity + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio19
  # AIC(k=2): 1635.28034518215

##2
phylostep(formula = woody_formula_2, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "BM", direction = "forward", trace = 2)
  # woody ~ 1 + BulkDensity + ClayPercentage + climaticMoistureIndex + PETseasonality + PETWettestQuarter + OrganicCarbon + Slope + bio19
  # AIC(k=2): 1666.58927644216
  
##3
phylostep(formula = woody_formula_3, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "BM", direction = "forward", trace = 2)
  # woody ~ 1 + BulkDensity + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio19
  # AIC(k=2): 1635.52026051005

##4
phylostep(formula = woody_formula_4, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "BM", direction = "forward", trace = 2)
  # woody ~ 1 + ClayPercentage + climaticMoistureIndex + PETDriestQuarter + PETWettestQuarter + Slope + bio6 + bio19
  # AIC(k=2): 1652.66052221238

#define formula components and run function (both)
##1
phylostep(formula = woody_formula_1, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "BM", direction = "both", trace = 2)
  # woody ~ 1 + BulkDensity + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio19
  # AIC(k=2): 1635.28034518215

##2
phylostep(formula = woody_formula_2, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "BM", direction = "both", trace = 2)
  # woody ~ 1 + BulkDensity + ClayPercentage + climaticMoistureIndex + PETseasonality + PETWettestQuarter + OrganicCarbon + Slope + bio19
  # AIC(k=2): 1666.58927644216

##3
phylostep(formula = woody_formula_3, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "BM", direction = "both", trace = 2)
  # woody ~ 1 + BulkDensity + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio19
  # AIC(k=2): 1635.28034518215

##4
phylostep(formula = woody_formula_4, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "BM", direction = "both", trace = 2)
  # woody ~ 1 + ClayPercentage + climaticMoistureIndex + PETDriestQuarter + PETWettestQuarter + Slope + bio6 + bio19
  # AIC(k=2): 1652.66052221238

```

The forward model with BM has the best model with formula 1, with an AIC of 1635.28034518215
  woody ~ 1 + BulkDensity + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio19
The both model with BM has the best  model with formula 1 and 3, with an AIC of 1635.28034518215
  woody ~ 1 + BulkDensity + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio19
  woody ~ 1 + BulkDensity + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio19

## Model selection with OUrandomRoot

OUrandomRoot stands for Ornstein-Uhlenbeck with a random root. 

```{r model selection OUrandomRoot}
#define formula components and run function (forward)
## 1
phylostep(formula = woody_formula_1, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "OUrandomRoot", direction = "forward", trace = 2)
  # woody ~ 1 + BulkDensity + minTempWarmest + monthCountByTemp10 + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio19
  # AIC(k=2): 863.408359010372

## 2
phylostep(formula = woody_formula_2, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "OUrandomRoot", direction = "forward", trace = 2)
  # woody ~ 1 + BulkDensity + ClayPercentage + climaticMoistureIndex + minTempWarmest + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + Slope
  # AIC(k=2): 893.479538952285

## 3
phylostep(formula = woody_formula_3, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "OUrandomRoot", direction = "forward", trace = 2)
  # woody ~ 1 + BulkDensity + minTempWarmest + monthCountByTemp10 + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio14 + bio19
  # AIC(k=2): 863.968744050452

## 4
phylostep(formula = woody_formula_4, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "OUrandomRoot", direction = "forward", trace = 2)
  # woody ~ 1 + climaticMoistureIndex + PETDriestQuarter + PETWettestQuarter + Slope + bio6
  # AIC(k=2): 882.506556683052

#define formula components and run function (both)
## 1
phylostep(formula = woody_formula_1, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "OUrandomRoot", direction = "both", trace = 2)
  # woody ~ 1 + BulkDensity + minTempWarmest + monthCountByTemp10 + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio19
  # AIC(k=2): 863.408359010372

## 2
phylostep(formula = woody_formula_2, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "OUrandomRoot", direction = "both", trace = 2)
  # woody ~ 1 + BulkDensity + ClayPercentage + climaticMoistureIndex + minTempWarmest + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + Slope
  # AIC(k=2): 893.479538952285

## 3
phylostep(formula = woody_formula_3, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "OUrandomRoot", direction = "both", trace = 2)
  # woody ~ 1 + BulkDensity + minTempWarmest + monthCountByTemp10 + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio19
  # AIC(k=2): 863.408359010372

## 4
phylostep(formula = woody_formula_4, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "OUrandomRoot", direction = "both", trace = 2)
  # woody ~ 1 + climaticMoistureIndex + PETDriestQuarter + PETWettestQuarter + Slope + bio6
  # AIC(k=2): 882.506556683052

```

The forward model with OUrandomRoot has the best model with formula 1, with an AIC of 863.408359010372
  woody ~ 1 + BulkDensity + minTempWarmest + monthCountByTemp10 + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio19
The both model with OUrandomRoot has the best model with formula 1 and 3, with an AIC of 863.408359010372
  woody ~ 1 + BulkDensity + minTempWarmest + monthCountByTemp10 + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio19
  woody ~ 1 + BulkDensity + minTempWarmest + monthCountByTemp10 + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio19

## Model selection with OUfixedRoot

OUfixedRoot stands for Ornstein-Uhlenbeck with a fixed root. 

```{r model selection OUfixedRoot}
#define formula components and run function (forward)
## 1
phylostep(formula = woody_formula_1, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "OUfixedRoot", direction = "forward", trace = 2)
  # woody ~ 1 + BulkDensity + minTempWarmest + monthCountByTemp10 + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio19
  # AIC(k=2): 863.408359010372
  
## 2
phylostep(formula = woody_formula_2, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "OUfixedRoot", direction = "forward", trace = 2)
  # woody ~ 1 + BulkDensity + ClayPercentage + climaticMoistureIndex + minTempWarmest + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + Slope
  # AIC(k=2): 893.479538952285

## 3
phylostep(formula = woody_formula_3, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "OUfixedRoot", direction = "forward", trace = 2)
  # woody ~ 1 + BulkDensity + minTempWarmest + monthCountByTemp10 + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio14 + bio19
  # AIC(k=2): 863.968744050452

## 4
phylostep(formula = woody_formula_4, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "OUfixedRoot", direction = "forward", trace = 2)
  # woody ~ 1 + climaticMoistureIndex + PETDriestQuarter + PETWettestQuarter + Slope + bio6
  # AIC(k=2): 882.506556683052

#define formula components and run function (both)
## 1
phylostep(formula = woody_formula_1, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "OUfixedRoot", direction = "both", trace = 2)
  # woody ~ 1 + BulkDensity + minTempWarmest + monthCountByTemp10 + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio19
  # AIC(k=2): 863.408359010372

## 2
phylostep(formula = woody_formula_2, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "OUfixedRoot", direction = "both", trace = 2)
  # woody ~ 1 + BulkDensity + ClayPercentage + climaticMoistureIndex + minTempWarmest + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + Slope
  # AIC(k=2): 893.479538952285

## 3
phylostep(formula = woody_formula_3, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "OUfixedRoot", direction = "both", trace = 2)
  # woody ~ 1 + BulkDensity + minTempWarmest + monthCountByTemp10 + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio19
  # AIC(k=2): 863.408359010372

## 4
phylostep(formula = woody_formula_4, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "OUfixedRoot", direction = "both", trace = 2)
  # woody ~ 1 + climaticMoistureIndex + PETDriestQuarter + PETWettestQuarter + Slope + bio6
  # AIC(k=2): 882.506556683052
```

The forward model with OUfixedRoot has the best model with formula 1, with an AIC of 863.408359010372
  woody ~ 1 + BulkDensity + minTempWarmest + monthCountByTemp10 + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio19
The both model with OUfixedRoot has the best model with formula 1 and 3, with an AIC of 863.408359010372
  woody ~ 1 + BulkDensity + minTempWarmest + monthCountByTemp10 + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio19
  woody ~ 1 + BulkDensity + minTempWarmest + monthCountByTemp10 + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio19

## Conclusion
The best model can be made with the following combination:
Formula: 
Model: 
Direction: 
