---
title: "Model_selection_myco"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Model selection mycorrhiza 

The linear model will have to be made, with the trait-presence-column as the dependent variable, and the 
GIS data frames as the predictor variables. The function phylostep of the phylolm package will be used. 

## Results of VIF-analysis
```{r VIF analysis}
#load required package
library(usdm)

#fill in the arguments of the function and execute it 
vif(GIS_myco[,c(3:43)])

#choose the variables to delete that have a higher VIF value than 10, but make sure it makes sense in terms of biology, results in:

# 1. Solely based on highest VIF value
vif(GIS_myco[,c(3:5, 7, 14, 20, 22:24, 26, 28, 39, 42, 43)])

# 2. With embergerQ
vif(GIS_myco[,c(3:5, 7, 10, 14, 20, 22:24, 26, 28, 39, 42)])

# 3. With climaticMoistureIndex
vif(GIS_myco[,c(3:5, 8, 14, 20, 22:24, 26, 28, 39, 43)])

# 4. With continentality
vif(GIS_myco[,c(3:5, 7, 9, 14, 20, 22:24, 26, 39, 42, 43)])

# 5. With thermicityIndex
vif(GIS_myco[,c(3:5, 20, 21, 22:24, 26, 39, 42, 43)])

# 6. With BIO8 and BIO9 
vif(GIS_myco[,c(3:5, 7, 22:24, 26, 32, 33, 39, 42, 43)])

# 7. With BIO8, BIO9, and EmbergerQ
vif(GIS_myco[,c(3:5, 10, 22:24, 26, 32, 33, 39, 42)])

# 8. With BIO8, BIO9, and climaticMoistureIndex
vif(GIS_myco[,c(3:5, 8, 22:24, 32, 33, 39, 43)])

# 9. With embergerQ and thermicityIndex
vif(GIS_myco[,c(3, 4, 10, 21, 23, 24, 26, 39, 42)])

# 10. With climaticMoistureIndex and thermicityIndex
vif(GIS_myco[,c(3:5, 8, 20, 21, 22:24, 26, 39, 43)])

```

## Change data to match the wanted data format
There are some flaws in the GIS data frame that make it unfit for the phylostep function. First of all
the row names have to match the tip.lables of the tree, and secondly the trait value (0/1) has to be
seen as a numeric value. 
```{r}
#make allmb_name column into the row names of this table
rownames(GIS_myco) <- GIS_myco[,1]
GIS_myco[,1] <- NULL

#make sure the values in the 'woody' column are seen as numeric values
GIS_myco$mycorrhiza <- as.numeric(GIS_myco$mycorrhiza)
#check whether it worked
is.numeric(GIS_myco$mycorrhiza)
```

## Define formulas
```{r define formula}
# 0. Starting formula
myco_starting_formula <- mycorrhiza ~ 1
  
# 1. Solely based on highest VIF value
myco_formula_1 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio15 + bio18 + bio19

# 2. With embergerQ
myco_formula_2 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + embergerQ + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio15 + bio18 

# 3. With climaticMoistureIndex
myco_formula_3 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio15 + bio19

# 4. With continentality
myco_formula_4 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + continentality + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio15 + bio18 + bio19

# 5. With thermicityIndex
myco_formula_5 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + PETWettestQuarter + thermicityIndex + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio15 + bio18 + bio19

# 6. With BIO8 and BIO9 
myco_formula_6 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + OrganicCarbon + PhCaCL + Slope + bio2 + bio8 + bio9 + bio15 + bio18 + bio19

# 7. With BIO8, BIO9, and EmbergerQ
myco_formula_7a <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + embergerQ + OrganicCarbon + PhCaCL + Slope + bio2 + bio8 + bio9 + bio15 + bio17 + bio18
myco_formula_7b <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + embergerQ + OrganicCarbon + PhCaCL + Slope + bio2 + bio8 + bio9 + bio14 + bio15 + bio18

# 8. With BIO8, BIO9, and climaticMoistureIndex
myco_formula_8a <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + OrganicCarbon + PhCaCL + Slope + bio8 + bio9 + bio14 + bio15 + bio19
myco_formula_8b <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + OrganicCarbon + PhCaCL + Slope + bio3 + bio8 + bio9 + bio15 + bio17 + bio19

# 9. With embergerQ and thermicityIndex
myco_formula_9a <- mycorrhiza ~ Aspect + BulkDensity + embergerQ + thermicityIndex + PhCaCL + Slope + bio2 + bio14 + bio15 + bio18
myco_formula_9b <- mycorrhiza ~ Aspect + BulkDensity + embergerQ + PETWettestQuarter + thermicityIndex + OrganicCarbon + PhCaCL + Slope + bio2 + bio15 + bio17 + bio18

# 10. With climaticMoistureIndex and thermicityIndex
myco_formula_10 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + PETWettestQuarter + thermicityIndex + OrganicCarbon + PhCaCL + Slope + bio2 + bio15 + bio19
```

## Extra formulas
```{r}
# 11. With annualPET
myco_formula_11 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + annualPET + minTempWarmest + PETseasonality + OrganicCarbon + PhCaCL + Slope + bio7 + bio14 + bio15 + bio18 + bio19

# 12. With growingDegDays0
myco_formula_12 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + growingDegDays0 + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio15 + bio18 + bio19

# 13. With growingDegDays5
myco_formula_13 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + growingDegDays5 + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio15 + bio18 + bio19

# 14. With maxTempColdest
myco_formula_14 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + maxTempColdest + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio15 + bio18 + bio19

# 15. With monthCountByTemp10
myco_formula_15 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + monthCountByTemp10 + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio15 +  bio18 + bio19

# 16. With PETColdestQuarter
myco_formula_16 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETColdestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio15 +  bio18 + bio19

# 17. With PETDriestQuarter
myco_formula_17 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio14 + bio15 +  bio18 + bio19

# 18. With PETseasonality
myco_formula_18 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETseasonality + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio15 +  bio18 + bio19

# 19. With PETWarmestQuarter
myco_formula_19 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio4 + bio15 + bio18 + bio19

# 20. With bio1
myco_formula_20 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio1 + bio2 + bio15 + bio18 + bio19

# 21. With bio3
myco_formula_21 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio3 + bio15 + bio18 + bio19

# 22. With bio5 
myco_formula_22 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio15 + bio18 + bio19

# 23. With bio6
myco_formula_23 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio6 + bio15 + bio18 + bio19

# 24. With bio7
myco_formula_24 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio7 + bio15 + bio18 + bio19

# 25. With bio10
myco_formula_25 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio10 + bio15 + bio18 + bio19

# 26. With bio11
myco_formula_26 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio11 + bio15 + bio18 + bio19

# 27. With bio12 
myco_formula_27 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio12 + bio15 + bio19

# 28. With bio13
myco_formula_28 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio13 + bio15 + bio19

# 29. With bio14 
myco_formula_29 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio14 + bio15 +  bio18 + bio19

# 30. With bio16
myco_formula_30 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio15 + bio16 + bio19

# 31. With bio17
myco_formula_31 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio13 + bio15 + bio17 + bio19

# 32. With things
myco_formula_32 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + embergerQ + OrganicCarbon + PhCaCL + Slope + bio2 + bio6 + bio8 + bio15 + bio19

```


## Model selection with BM
```{r model selection BM}
#load required package
library(phylolm)

#define formula components and run function (forward)
## 1
phylostep(formula = myco_formula_1, starting.formula = myco_starting_formula, data = GIS_myco, phy = tree_myco, model = "BM", direction = "forward", trace = 2)
  # mycorrhiza ~ 1 + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + PhCaCL + Slope + bio2 + bio4 + bio15 + bio18 + bio19

## other formulas
phylostep(formula = myco_formula_9b, starting.formula = myco_starting_formula, data = GIS_myco, phy = tree_myco, model = "BM", direction = "forward", trace = 2)

#define formula components and run function (both)
## 1 
phylostep(formula = myco_formula_1, starting.formula = myco_starting_formula, data = GIS_myco, phy = tree_myco, model = "BM", direction = "both", trace = 2)
  # mycorrhiza ~ 1 + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + PhCaCL + Slope + bio2 + bio4 + bio15 + bio18 + bio19
  # AIC(k=2): 2472.33792047599

## other formulas
phylostep(formula = myco_formula_9b, starting.formula = myco_starting_formula, data = GIS_myco, phy = tree_myco, model = "BM", direction = "both", trace = 2)
  
```

## Model selection OUrandomRoot

```{r model selection OUrandomRoot}
#define formula components and run function (forward)
## 1
phylostep(formula = myco_formula_1, starting.formula = myco_starting_formula, data = GIS_myco, phy = tree_myco, model = "OUrandomRoot", direction = "forward", trace = 2)
  #  mycorrhiza ~ 1 + BulkDensity + minTempWarmest + PETWettestQuarter + PhCaCL + bio2 + bio4 + bio15 + bio19
  # AIC(k=2): 1497.98125594055

## other formulas
phylostep(formula = myco_formula_9b, starting.formula = myco_starting_formula, data = GIS_myco, phy = tree_myco, model = "OUrandomRoot", direction = "forward", trace = 2)

#define formula components and run function (both)
## 1
phylostep(formula = myco_formula_1, starting.formula = myco_starting_formula, data = GIS_myco, phy = tree_myco, model = "OUrandomRoot", direction = "both", trace = 2)
  # mycorrhiza ~ 1 + BulkDensity + minTempWarmest + PETWettestQuarter + PhCaCL + bio2 + bio4 + bio15 + bio19
  # AIC(k=2): 1497.98125594055

## other formulas
phylostep(formula = myco_formula_9a, starting.formula = myco_starting_formula, data = GIS_myco, phy = tree_myco, model = "OUrandomRoot", direction = "both", trace = 2)
  
```

## Model selection with OUfixedRoot

```{r model selection OUfixedRoot}
#define formula components and run function (forward)
## 1
phylostep(formula = myco_formula_1, starting.formula = myco_starting_formula, data = GIS_myco, phy = tree_myco, model = "OUfixedRoot", direction = "forward", trace = 2)
  # mycorrhiza ~ 1 + BulkDensity + minTempWarmest + PETWettestQuarter + PhCaCL + bio2 + bio4 + bio15 + bio19
  # AIC(k=2): 1497.98125594055

## other formulas
phylostep(formula = myco_formula_9b, starting.formula = myco_starting_formula, data = GIS_myco, phy = tree_myco, model = "OUfixedRoot", direction = "forward", trace = 2)

#define formula components and run function (both)
## 1
phylostep(formula = myco_formula_1, starting.formula = myco_starting_formula, data = GIS_myco, phy = tree_myco, model = "OUfixedRoot", direction = "both", trace = 2)
  # mycorrhiza ~ 1 + BulkDensity + minTempWarmest + PETWettestQuarter + PhCaCL + bio2 + bio4 + bio15 + bio19
  # AIC(k=2): 1497.98125594055

## other formulas
phylostep(formula = myco_formula_9a, starting.formula = myco_starting_formula, data = GIS_myco, phy = tree_myco, model = "OUfixedRoot", direction = "both", trace = 2)
 
```

## Conclusion
The best model can be made with the following combination:
Formula: 
Model: 
Direction: 

