---
title: "Model_selection_myco"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Model selection mycorrhiza

The linear model will have to be made, with the trait-presence-column as the dependent variable, and the 
GIS data frames as the predictor variables. The function phylostep of the phylolm package will be used. 

## Define formulas 

### First set of model formuals
Based on the hypothesis. 

```{r}
# 0. Starting formula
myco_starting_formula <- mycorrhiza ~ 1
  
# 1. Solely based on highest VIF value
myco_formula_1 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio15 + bio18 + bio19

# 2. With embergerQ
myco_formula_2 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + embergerQ + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio15 + bio18 

# 3. With climaticMoistureIndex
myco_formula_3 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio15 + bio19

# 4. With continentality
myco_formula_4 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + continentality + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio15 + bio18 + bio19

# 5. With thermicityIndex
myco_formula_5 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + PETWettestQuarter + thermicityIndex + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio15 + bio18 + bio19

# 6. bio8
myco_formula_6 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio8 + bio18 + bio19

# 7. bio9
myco_formula_7 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio9 + bio18 + bio19

# 8. With BIO8 and BIO9 
myco_formula_8 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + OrganicCarbon + PhCaCL + Slope + bio2 + bio8 + bio9 + bio15 + bio18 + bio19

# 9. With BIO8, BIO9, and EmbergerQ
myco_formula_9a <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + embergerQ + OrganicCarbon + PhCaCL + Slope + bio2 + bio8 + bio9 + bio15 + bio17 + bio18
myco_formula_9b <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + embergerQ + OrganicCarbon + PhCaCL + Slope + bio2 + bio8 + bio9 + bio14 + bio15 + bio18

# 10. With BIO8, BIO9, and climaticMoistureIndex
myco_formula_10a <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + OrganicCarbon + PhCaCL + Slope + bio8 + bio9 + bio14 + bio15 + bio19
myco_formula_10b <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + OrganicCarbon + PhCaCL + Slope + bio3 + bio8 + bio9 + bio15 + bio17 + bio19
myco_formula_10c <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + OrganicCarbon + PhCaCL + Slope + bio3 + bio8 + bio9 + bio14 + bio15 + bio19

# 11. With embergerQ and thermicityIndex
myco_formula_11a <- mycorrhiza ~ Aspect + BulkDensity + embergerQ + thermicityIndex + PhCaCL + Slope + bio2 + bio14 + bio15 + bio18
myco_formula_11b <- mycorrhiza ~ Aspect + BulkDensity + embergerQ + PETWettestQuarter + thermicityIndex + OrganicCarbon + PhCaCL + Slope + bio2 + bio15 + bio17 + bio18
myco_formula_11c <- mycorrhiza ~ Aspect + BulkDensity + embergerQ + PETWettestQuarter + thermicityIndex + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio15 + bio18

# 12. With climaticMoistureIndex and thermicityIndex
myco_formula_12 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + PETWettestQuarter + thermicityIndex + OrganicCarbon + PhCaCL + Slope + bio2 + bio15 + bio19
```

### Second set of model formulas 
The rest of the variables that were not included yet. 

```{r}
# 13. With annualPET
myco_formula_13 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + annualPET + minTempWarmest + PETseasonality + OrganicCarbon + PhCaCL + Slope + bio7 + bio14 + bio15 + bio18 + bio19

# 14. With growingDegDays0
myco_formula_14 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + growingDegDays0 + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio15 + bio18 + bio19

# 15. With growingDegDays5
myco_formula_15 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + growingDegDays5 + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio15 + bio18 + bio19

# 16. With maxTempColdest
myco_formula_16 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + maxTempColdest + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio15 + bio18 + bio19

# 17. With monthCountByTemp10
myco_formula_17 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + monthCountByTemp10 + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio15 +  bio18 + bio19

# 18. With PETColdestQuarter
myco_formula_18 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETColdestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio15 +  bio18 + bio19

# 19. With PETDriestQuarter
myco_formula_19 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio14 + bio15 +  bio18 + bio19

# 20. With PETseasonality
myco_formula_20 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETseasonality + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio15 +  bio18 + bio19

# 21. With PETWarmestQuarter
myco_formula_21 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio4 + bio15 + bio18 + bio19

# 22. With bio1
myco_formula_22 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio1 + bio2 + bio15 + bio18 + bio19

# 23. With bio3
myco_formula_23 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio3 + bio15 + bio18 + bio19

# 25. With bio5 
myco_formula_24 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio15 + bio18 + bio19

# 25. With bio6
myco_formula_25 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio6 + bio15 + bio18 + bio19

# 26. With bio7
myco_formula_26 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio7 + bio15 + bio18 + bio19

# 27. With bio10
myco_formula_27 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio10 + bio15 + bio18 + bio19

# 28. With bio11
myco_formula_28 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio11 + bio15 + bio18 + bio19

# 29. With bio12 
myco_formula_29 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio12 + bio15 + bio19

# 30. With bio13
myco_formula_30 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio13 + bio15 + bio19

# 31. With bio14 
myco_formula_31 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio14 + bio15 +  bio18 + bio19

# 32. With bio16
myco_formula_32 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio15 + bio16 + bio19

# 33. With bio17
myco_formula_33 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio13 + bio15 + bio17 + bio19
```

## Change data to match the wanted data format
There are some flaws in the GIS data frame that make it unfit for the phylostep function. First of all
the row names have to match the tip.lables of the tree, and secondly the trait value (0/1) has to be
seen as a numeric value. 
```{r}
#make allmb_name column into the row names of this table
rownames(GIS_myco) <- GIS_myco[,1]
GIS_myco[,1] <- NULL

#make sure the values in the 'woody' column are seen as numeric values
GIS_myco$mycorrhiza <- as.numeric(GIS_myco$mycorrhiza)
#check whether it worked
is.numeric(GIS_myco$mycorrhiza)
```

## Preform model selection 
With the phyloglmstep() function by Rutger Vos.

```{r}
#source function and load package
library(phylolm)
source('phyloglmstep.R')

#fill in arguments and execute function
phyloglmstep(myco_formula_4, starting.formula = NULL, data=GIS_myco, phy=tree_myco, 
                         method="logistic_MPLE",
                         direction = "both", trace = 0,
                         btol = 100, log.alpha.bound = 8, start.beta=NULL, 
                         start.alpha=NULL, boot = 0, full.matrix = TRUE,
                         k=2)
```