---
title: "Model_selection_myco"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Model selection mycorrhiza 

The linear model will have to be made, with the trait-presence-column as the dependent variable, and the 
GIS data frames as the predictor variables. The function phylostep of the phylolm package will be used. 

## Results of VIF-analysis
```{r VIF analysis}
#load required package
library(usdm)

#fill in the arguments of the function and execute it 
vif(GIS_myco[,c(3:43)])

#choose the variables to delete that have a higher VIF value than 10, but make sure it makes sense in terms of biology, results in:
# 1. Without embergerQ
vif(GIS_myco[,c(3:5, 7, 14, 20, 22:24, 26, 28, 39, 42, 43)])

# 2. With embergerQ
vif(GIS_myco[,c(3:5, 7, 10, 14, 20, 22:24, 26, 28, 39, 42)])
```

## Change data to match the wanted data format
There are some flaws in the GIS data frame that make it unfit for the phylostep function. First of all
the row names have to match the tip.lables of the tree, and secondly the trait value (0/1) has to be
seen as a numeric value. 
```{r}
#make allmb_name column into the row names of this table
rownames(GIS_myco) <- GIS_myco[,1]
GIS_myco[,1] <- NULL

#make sure the values in the 'woody' column are seen as numeric values
GIS_myco$mycorrhiza <- as.numeric(GIS_myco$mycorrhiza)
#check whether it worked
is.numeric(GIS_myco$mycorrhiza)
```

## Define formulas
```{r define formula}
# 0. Starting formula
myco_starting_formula <- mycorrhiza ~ 1
  
# 1. Without embergerQ
myco_formula_1 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio15 + bio18 + bio19

# 2. With embergerQ
myco_formula_2 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + embergerQ + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio15 + bio18 
```

## Model selection with BM
```{r model selection BM}
#load required package
library(phylolm)

#define formula components and run function (forward)
## 1
phylostep(formula = myco_formula_1, starting.formula = myco_starting_formula, data = GIS_myco, phy = tree_myco, model = "BM", direction = "forward", trace = 2)
  # mycorrhiza ~ 1 + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + PhCaCL + Slope + bio2 + bio4 + bio15 + bio18 + bio19

## 2
phylostep(formula = myco_formula_2, starting.formula = myco_starting_formula, data = GIS_myco, phy = tree_myco, model = "BM", direction = "forward", trace = 2)
  # mycorrhiza ~ 1 + BulkDensity + ClayPercentage + aridityIndexThornthwaite + embergerQ + PETWettestQuarter + Slope + bio2 + bio4 + bio15

#define formula components and run function (both)
## 1 
phylostep(formula = myco_formula_1, starting.formula = myco_starting_formula, data = GIS_myco, phy = tree_myco, model = "BM", direction = "both", trace = 2)
  # mycorrhiza ~ 1 + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + PhCaCL + Slope + bio2 + bio4 + bio15 + bio18 + bio19

## 2
phylostep(formula = myco_formula_2, starting.formula = myco_starting_formula, data = GIS_myco, phy = tree_myco, model = "BM", direction = "both", trace = 2)
  # mycorrhiza ~ 1 + ClayPercentage + aridityIndexThornthwaite + embergerQ + PETWettestQuarter + Slope + bio2 + bio4 + bio15
```

The forward model with BM has the following model as the best model, with an AIC of...
  mycorrhiza ~
The both model with BM has the following model as the best model, with an AIC of ...
  mycorrhiza ~

## Model selection OUrandomRoot

```{r model selection OUrandomRoot}
#define formula components and run function (forward)
## 1
phylostep(formula = myco_formula_1, starting.formula = myco_starting_formula, data = GIS_myco, phy = tree_myco, model = "OUrandomRoot", direction = "forward", trace = 2)
  #  mycorrhiza ~ 1 + BulkDensity + minTempWarmest + PETWettestQuarter + PhCaCL + bio2 + bio4 + bio15 + bio19

## 2
phylostep(formula = myco_formula_2, starting.formula = myco_starting_formula, data = GIS_myco, phy = tree_myco, model = "OUrandomRoot", direction = "forward", trace = 2)
  # mycorrhiza ~ 1 + BulkDensity + embergerQ + minTempWarmest + PhCaCL + Slope + bio2 + bio15

#define formula components and run function (both)
## 1
phylostep(formula = myco_formula_1, starting.formula = myco_starting_formula, data = GIS_myco, phy = tree_myco, model = "OUrandomRoot", direction = "both", trace = 2)
  # mycorrhiza ~ 1 + BulkDensity + minTempWarmest + PETWettestQuarter + PhCaCL + bio2 + bio4 + bio15 + bio19

## 2
phylostep(formula = myco_formula_2, starting.formula = myco_starting_formula, data = GIS_myco, phy = tree_myco, model = "OUrandomRoot", direction = "both", trace = 2)
  # mycorrhiza ~ 1 + BulkDensity + embergerQ + minTempWarmest + PhCaCL + Slope + bio2 + bio15
```

The forward model with OUrandomRoot has the following model as the best model, with an AIC of...
  mycorrhiza ~
The both model with OUrandomRoot has the following model as the best model, with an AIC of ...
  mycorrhiza ~

## Model selection with OUfixedRoot

```{r model selection OUfixedRoot}
#define formula components and run function (forward)
## 1
phylostep(formula = myco_formula_1, starting.formula = myco_starting_formula, data = GIS_myco, phy = tree_myco, model = "OUfixedRoot", direction = "forward", trace = 2)
  # mycorrhiza ~ 1 + BulkDensity + minTempWarmest + PETWettestQuarter + PhCaCL + bio2 + bio4 + bio15 + bio19

## 2
phylostep(formula = myco_formula_2, starting.formula = myco_starting_formula, data = GIS_myco, phy = tree_myco, model = "OUfixedRoot", direction = "forward", trace = 2)
  # mycorrhiza ~ 1 + BulkDensity + embergerQ + minTempWarmest + PhCaCL + Slope + bio2 + bio15

#define formula components and run function (both)
## 1
phylostep(formula = myco_formula_1, starting.formula = myco_starting_formula, data = GIS_myco, phy = tree_myco, model = "OUfixedRoot", direction = "both", trace = 2)
  # mycorrhiza ~ 1 + BulkDensity + minTempWarmest + PETWettestQuarter + PhCaCL + bio2 + bio4 + bio15 + bio19

## 2
phylostep(formula = myco_formula_2, starting.formula = myco_starting_formula, data = GIS_myco, phy = tree_myco, model = "OUfixedRoot", direction = "both", trace = 2)
  # mycorrhiza ~ 1 + BulkDensity + embergerQ + minTempWarmest + PhCaCL + Slope + bio2 + bio15
```

The forward model with OUfixedRoot has the following model as the best model, with an AIC of...
  mycorrhiza ~
The both model with OUfixedRoot has the following model as the best model, with an AIC of ...
  mycorrhiza ~

## Conclusion
The best model can be made with the following combination:
Formula: 
Model: 
Direction: 

