---
title: "Model_selection_woody"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Model selection derived woodiness

The linear model will have to be made, with the trait-presence-column as the dependent variable, and the 
GIS data frames as the predictor variables. The function phylostep of the phylolm package will be used. 

## Define formulas 

### First set of model formuals
Based on the hypothesis. 

```{r}
# 0. Starting formula
woody_starting_formula <- woody ~ 1 

# 1. Solely based on lowest VIF value 
woody_formula_1 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19

# 2. With aridityIndexThornthwaite
woody_formula_2a <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio15 + bio19
woody_formula_2b <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio16 + bio19
woody_formula_2c <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio19

# 3. With ClimaticMoistureIndex
woody_formula_3 <- woody ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + minTempWarmest + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio19

# 4. With embergerQ
woody_formula_4a <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + continentality + embergerQ + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio19
woody_formula_4b <- woody ~ Aspect + BulkDensity + ClayPercentage +embergerQ + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio4 + bio14 + bio15 + bio19
woody_formula_4c <- woody ~ Aspect + BulkDensity + ClayPercentage +embergerQ + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio15 + bio17 + bio19
woody_formula_4d <- woody ~ Aspect + BulkDensity + ClayPercentage +embergerQ + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio7 + bio15 + bio19

# 5. With growingDegDays0
woody_formula_6 <- woody ~ Aspect + BulkDensity + ClayPercentage + growingDegDays0 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio18 + bio19

# 6. With PETColdestQuarter
woody_formula_7 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETColdestQuarter + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19

# 7. With PETWarmestQuarter
woody_formula_8 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWarmestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19

# 8. With BIO6 (30) 
woody_formula_9 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio14 + bio15 + bio18 + bio19

# 9. With bio12
woody_formula_9 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio12 + bio15 + bio19

# 10. With bio13
woody_formula_10 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio14 + bio15 + bio19

# 11. With bio16
woody_formula_11 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio16 + bio19

# 12. With bio17
woody_formula_12a <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio15 + bio17 + bio19
woody_formula_12b <- woody ~ Aspect + BulkDensity + ClayPercentage+ minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio16 + bio17 + bio19
```

### Second set of model formulas 
The rest of the variables that were not included yet. 

```{r}
# 13. With annualPET
woody_formula_13 <- woody ~ Aspect + BulkDensity + ClayPercentage + annualPET + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19

# 14. With continentality
woody_formula_14a <- woody ~ Aspect + BulkDensity + ClayPercentage + annualPET + continentality + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19
woody_formula_14b <- woody ~ Aspect + BulkDensity + ClayPercentage + continentality + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19
woody_formula_14c <- woody ~ Aspect + BulkDensity + ClayPercentage + continentality + minTempWarmest + monthCountByTemp10 + PETDriestQuarter +  PETWarmestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19
woody_formula_14d <- woody ~ Aspect + BulkDensity + ClayPercentage + continentality + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19
woody_formula_14e <- woody ~ Aspect + BulkDensity + ClayPercentage + continentality + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio15 + bio18 + bio19
woody_formula_14f <- woody ~ Aspect + BulkDensity + ClayPercentage + continentality + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + OrganicCarbon + PhCaCL + Slope + bio3 + bio14 + bio15 + bio18 + bio19
woody_formula_14g <- woody ~ Aspect + BulkDensity + ClayPercentage + continentality + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + OrganicCarbon + PhCaCL + Slope + bio8 + bio14 + bio15 + bio18 + bio19

# 15. With growingDegDays5
woody_formula_15 <- woody ~ Aspect + BulkDensity + ClayPercentage + growingDegDays5 + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio18 + bio19

# 16. With maxTempColdest
woody_formula_16 <- woody ~ Aspect + BulkDensity + ClayPercentage + maxTempColdest + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19

# 17. With thermicityIndex
woody_formula_17 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + thermicityIndex + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19

# 18. With bio1
woody_formula_18 <- woody ~ Aspect + BulkDensity + ClayPercentage + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio1 + bio14 + bio15 + bio18 + bio19

# 19. With bio2
woody_formula_19 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio15 + bio18 + bio19

# 20. With bio3
woody_formula_20 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio3 + bio14 + bio15 + bio18 + bio19

# 21. With bio4
woody_formula_21 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio4 + bio14 + bio15 + bio18 + bio19

# 22. With bio5
woody_formula_22a <- woody ~ Aspect + BulkDensity + ClayPercentage + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio5 + bio14 + bio15 + bio18 + bio19
woody_formula_22b <- woody ~ Aspect + BulkDensity + ClayPercentage + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio5 + bio6 + bio15 + bio18 + bio19

# 23. With bio7
woody_formula_23 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio7 + bio14 + bio15 +  bio18 + bio19

# 24. With bio8 
woody_formula_24 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + OrganicCarbon + PhCaCL + Slope + bio8 + bio14 + bio15 + bio18 + bio19

# 25. With bio9
woody_formula_25 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio9 + bio14 + bio15 + bio18 + bio19

# 26. With bio10
woody_formula_26 <- woody ~ Aspect + BulkDensity + ClayPercentage + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio10 + bio14 + bio15 + bio18 + bio19

# 27. With bio11
woody_formula_27 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio11 + bio14 + bio15 + bio18 + bio19
```

## Preform model selection
With the phyloglmstep() function by Rutger Vos.

```{r}
#source function and load required package
source('phyloglmstep.R')
library(phylolm)

#fill in arguments and execute function
phyloglmstep(woody_formula_22a, starting.formula = NULL, data=GIS_woody, phy=tree_woody, 
                         method="logistic_MPLE",
                         direction = "forward", trace = 0,
                         btol = 100, log.alpha.bound = 8, start.beta=NULL, 
                         start.alpha=NULL, boot = 0, full.matrix = TRUE,
                         k=2)
```