---
title: "Model_selection_crops"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Model selection domestication

The linear model will have to be made, with the trait-presence-column as the dependent variable, and the 
GIS data frames as the predictor variables. The function phylostep of the phylolm package will be used. 

## Results of VIF-analysis
```{r}
# 1. 


# 2. 

```

## Define formulas
```{r}
# 0. Starting formula
crops_starting_formula <- 

# 1. 
crops_formula_1 <- 

# 2. 
crops_formula_2 <- 
```


## Model selection with BM

```{r model selection BM}
#load required package
library(phylolm)

#define formula components and run function (forward)
## 1
phylostep(formula = crops_formula_1, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "BM", direction = "forward", trace = 2)
## 2
phylostep(formula = crops_formula_2, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "BM", direction = "forward", trace = 2)

#define formula components and run function (both)
## 1
phylostep(formula = crops_formula_1, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "BM", direction = "both", trace = 2)
## 2
phylostep(formula = crops_formula_1, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "BM", direction = "both", trace = 2)
```

The forward model with BM has the following model as the best model, with an AIC of...
  domestication ~
The both model with BM has the following model as the best model, with an AIC of ...
  domestication ~

## Model selection with OUrandomRoot

```{r model selection OUrandomRoot}
#define formula components and run function (forward)
## 1
phylostep(formula = crops_formula_1, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "OUrandomRoot", direction = "forward", trace = 2)
## 2
phylostep(formula = crops_formula_2, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "OUrandomRoot", direction = "forward", trace = 2)

#define formula components and run function (both)
## 1
phylostep(formula = crops_formula_1, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "OUrandomRoot", direction = "both", trace = 2)
## 2
phylostep(formula = crops_formula_2, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "OUrandomRoot", direction = "both", trace = 2)
```

The forward model with OUrandomRoot has the following model as the best model, with an AIC of...
  domestication ~
The both model with OUrandomRoot has the following model as the best model, with an AIC of ...
  domestication ~

## Model selection with OUfixedRoot

```{r model selection OUfixedRoot}
#define formula components and run function (forward)
## 1
phylostep(formula = crops_formula_1, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "OUfixedRoot", direction = "forward", trace = 2)
## 2
phylostep(formula = crops_formula_2, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "OUfixedRoot", direction = "forward", trace = 2)

#define formula components and run function (both)
## 1
phylostep(formula = crops_formula_1, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "OUfixedRoot", direction = "both", trace = 2)
## 2
phylostep(formula = crops_formula_2, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "OUfixedRoot", direction = "both", trace = 2)
```

The forward model with OUfixedRoot has the following model as the best model, with an AIC of...
  domestication ~
The both model with OUfixedRoot has the following model as the best model, with an AIC of ...
  domestication ~
