---
title: "Model_selection_woody_elaborated"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Model selection derived woodiness

## Define formulas 

### First set of model formuals
Based on the hypothesis. 

```{r}
# 0. Starting formula
woody_starting_formula <- woody ~ 1 

# 1. Solely based on lowest VIF value 
woody_formula_1 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19

# 2. With ClimaticMoistureIndex
woody_formula_2 <- woody ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + minTempWarmest + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio19

# 3. With BIO6 (30) 
woody_formula_3 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio14 + bio15 + bio18 + bio19

# 4. With growingDegDays0
woody_formula_4 <- woody ~ Aspect + BulkDensity + ClayPercentage + growingDegDays0 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio18 + bio19

# 5. With aridityIndexThornthwaite
woody_formula_5a <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio15 + bio19
woody_formula_5b <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio16 + bio19
woody_formula_5c <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio19

# 6. With embergerQ
woody_formula_6a <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + continentality + embergerQ + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio19
woody_formula_6b <- woody ~ Aspect + BulkDensity + ClayPercentage +embergerQ + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio4 + bio14 + bio15 + bio19
woody_formula_6c <- woody ~ Aspect + BulkDensity + ClayPercentage +embergerQ + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio15 + bio17 + bio19
woody_formula_6d <- woody ~ Aspect + BulkDensity + ClayPercentage +embergerQ + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio7 + bio15 + bio19

# 7. With BIO6 and ClimaticMoistureIndex
woody_formula_7 <- woody ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio15 + bio19

# 8. With BIO6 and aridityIndexThornthwaite
woody_formula_8a <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + embergerQ + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio15 + bio19
woody_formula_8b <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio13 + bio15 + bio19
woody_formula_8c <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio15 + bio16 + bio19

# 9. With BIO6 and embergerQ
woody_formula_9a <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + embergerQ + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio15 + bio19
woody_formula_9b <- woody ~ Aspect + BulkDensity + ClayPercentage + embergerQ + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio14 + bio15 + bio19
woody_formula_9c <- woody ~ Aspect + BulkDensity + ClayPercentage + embergerQ + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio15 + bio17 + bio19

# 10. With growingDegDays0 and ClimaticMoistureIndex
woody_formula_10 <- woody ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + growingDegDays0 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio19

# 11. With growingDegDays0 and aridityIndexThornthwaite
woody_formula_11a <- woody ~ Aspect + BulkDensity + ClayPercentage + embergerQ + aridityIndexThornthwaite + growingDegDays0 + minTempWarmest + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio19
woody_formula_11b <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + growingDegDays0 + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio13 + bio15 + bio19
woody_formula_11c <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + growingDegDays0 + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio16 + bio19
woody_formula_11d <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + growingDegDays0 + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio18 + bio19

# 12. With growingDegDays0 and embergerQ
woody_formula_12a <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + continentality + embergerQ + growingDegDays0 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio19
woody_formula_12b <- woody ~ Aspect + BulkDensity + ClayPercentage + embergerQ + growingDegDays0 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio4 + bio15 + bio19
woody_formula_12c <- woody ~ Aspect + BulkDensity + ClayPercentage + embergerQ + growingDegDays0 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio3 + bio14 + bio15 + bio19
woody_formula_12d <- woody ~ Aspect + BulkDensity + ClayPercentage + embergerQ + growingDegDays0 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio4 + bio15 + bio17 + bio19
woody_formula_12e <- woody ~ Aspect + BulkDensity + ClayPercentage + embergerQ + growingDegDays0 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio5 + bio15 + bio19

```

### Second set of model formulas 
The rest of the variables that were not included yet. 

```{r}
# 13. With things
woody_formula_13 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19

# 14. With annualPET
woody_formula_14 <- woody ~ Aspect + BulkDensity + ClayPercentage + annualPET + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19

# 15. With continentality
woody_formula_15a <- woody ~ Aspect + BulkDensity + ClayPercentage + annualPET + continentality + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19
woody_formula_15b <- woody ~ Aspect + BulkDensity + ClayPercentage + continentality + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19
woody_formula_15c <- woody ~ Aspect + BulkDensity + ClayPercentage + continentality + minTempWarmest + monthCountByTemp10 + PETDriestQuarter +  PETWarmestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19
woody_formula_15d <- woody ~ Aspect + BulkDensity + ClayPercentage + continentality + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19
woody_formula_15e <- woody ~ Aspect + BulkDensity + ClayPercentage + continentality + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio15 + bio18 + bio19
woody_formula_15f <- woody ~ Aspect + BulkDensity + ClayPercentage + continentality + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + OrganicCarbon + PhCaCL + Slope + bio3 + bio14 + bio15 + bio18 + bio19
woody_formula_15g <- woody ~ Aspect + BulkDensity + ClayPercentage + continentality + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + OrganicCarbon + PhCaCL + Slope + bio8 + bio14 + bio15 + bio18 + bio19

# 16. With growingDegDays0
woody_formula_16 <- woody ~ Aspect + BulkDensity + ClayPercentage + growingDegDays5 + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio18 + bio19

# 17. With maxTempColdest
woody_formula_17 <- woody ~ Aspect + BulkDensity + ClayPercentage + maxTempColdest + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19
 
# 18. With PETColdestQuarter
woody_formula_18 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETColdestQuarter + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19

# 19. With PETWarmestQuarter
woody_formula_19 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWarmestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19

# 20. With thermicityIndex
woody_formula_20 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + thermicityIndex + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19

# 21. With bio1
woody_formula_21 <- woody ~ Aspect + BulkDensity + ClayPercentage + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio1 + bio14 + bio15 + bio18 + bio19

# 22. With bio2
woody_formula_22 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio15 + bio18 + bio19

# 23. With bio3
woody_formula_23 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio3 + bio14 + bio15 + bio18 + bio19

# 24. With bio4
woody_formula_24 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio4 + bio14 + bio15 + bio18 + bio19

# 25. With bio5
woody_formula_25a <- woody ~ Aspect + BulkDensity + ClayPercentage + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio5 + bio14 + bio15 + bio18 + bio19
woody_formula_25b <- woody ~ Aspect + BulkDensity + ClayPercentage + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio5 + bio6 + bio15 + bio18 + bio19

# 26. With bio7
woody_formula_26 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio7 + bio14 + bio15 +  bio18 + bio19

# 27. With bio8 
woody_formula_27 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + OrganicCarbon + PhCaCL + Slope + bio8 + bio14 + bio15 + bio18 + bio19

# 28. With bio9
woody_formula_28 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio9 + bio14 + bio15 + bio18 + bio19

# 29. With bio10
woody_formula_29 <- woody ~ Aspect + BulkDensity + ClayPercentage + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio10 + bio14 + bio15 + bio18 + bio19

# 30. With bio11
woody_formula_30 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio11 + bio14 + bio15 + bio18 + bio19

# 31. With bio12
woody_formula_31 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio12 + bio15 + bio19

# 32. With bio13
woody_formula_32 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio14 + bio15 + bio19

# 33. With bio16
woody_formula_33 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio16 + bio19

# 34. With bio17
woody_formula_34a <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio15 + bio17 + bio19
woody_formula_34b <- woody ~ Aspect + BulkDensity + ClayPercentage+ minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio16 + bio17 + bio19
```

## Preform model selection

### Model selection with BM 
BM stands for Brownian motion. 

```{r model selection BM}
#load required package
library(phylolm)

#define formula components and run function (forward)
phylostep(formula = woody_formula_9a, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "BM", direction = "forward", trace = 2)
 
#define formula components and run function (both)
phylostep(formula = woody_formula_13, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "BM", direction = "both", trace = 2)

```

### Model selection with OUrandomRoot
OUrandomRoot stands for Ornstein-Uhlenbeck with a random root. 

```{r model selection OUrandomRoot}
#define formula components and run function (forward)
phylostep(formula = woody_formula_13, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "OUrandomRoot", direction = "forward", trace = 2, lower.bound = 0, upper.bound = 1000)

#define formula components and run function (both)
phylostep(formula = woody_formula_14, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "OUrandomRoot", direction = "both", trace = 2, lower.bound = 0, upper.bound = 1000)
  
```

### Model selection with OUfixedRoot
OUfixedRoot stands for Ornstein-Uhlenbeck with a fixed root. 

```{r model selection OUfixedRoot}
#define formula components and run function (forward)
phylostep(formula = woody_formula_11, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "OUfixedRoot", direction = "forward", trace = 2, lower.bound = 0, upper.bound = 1000)
 
#define formula components and run function (both)
phylostep(formula = woody_formula_25b, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "OUfixedRoot", direction = "both", trace = 2, lower.bound = 0, upper.bound = 1000)

```
