---
title: "Model_selection_crops"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Model selection domestication

The linear model will have to be made, with the trait-presence-column as the dependent variable, and the 
GIS data frames as the predictor variables. The function phylostep of the phylolm package will be used. 

## Results of VIF-analysis
```{r}
# 1. Solely based on lowest VIF value
vif(GIS_crops[,c(3:5, 14, 17, 20, 22:24, 37, 38, 43)]) 

# 2. With PETseasonality
vif(GIS_crops[,c(3:5, 14, 18, 20, 22:24, 38, 39, 43)]) 

# 3. With BIO2 
vif(GIS_crops[,c(3:5, 14, 17, 20, 22:24, 26, 37, 38, 43)]) 

# 4. With BIO4 
vif(GIS_crops[,c(3:5, 14, 20, 22:24, 28, 39, 43)]) 

# 5. With BIO15
vif(GIS_crops[,c(3:5, 14, 17, 20, 22:24, 37, 39)]) 

# 6. With PETseasonality and BIO2
vif(GIS_crops[,c(3:5, 14, 18, 20, 22:24, 26, 38, 43)]) 

# 7. With PETseasonality and BIO4
vif(GIS_crops[,c(3, 5, 18, 20, 22, 24, 28, 43)]) 

# 8. With BIO15 and BIO2 
vif(GIS_crops[,c(3:5, 14, 17, 20, 22:24, 26, 39)]) 

# 9. With BIO15 and BIO4 = same as with just BIO4
vif(GIS_crops[,c(3:5, 14, 20, 22:24, 28, 39, 43)]) 
```

## Change data to match the wanted data format

There are some flaws in the GIS data frame that make it unfit for the phylostep function. First of all
the row names have to match the tip.lables of the tree, and secondly the trait value (0/1) has to be
seen as a numeric value. 
```{r}
#make allmb_name column into the row names of this table
rownames(GIS_crops) <- GIS_crops[,1]
GIS_crops[,1] <- NULL

#make sure the values in the 'woody' column are seen as numeric values
GIS_crops$domestication <- as.numeric(GIS_crops$domestication)
#check whether it worked
is.numeric(GIS_crops$domestication)
```

## Define formulas
```{r}
# 0. Starting formula
crops_starting_formula <- domestication ~ 1 

# 1. Lowest VIF value
crops_formula_1 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio14 + bio19

# 2. With PETseasonality
crops_formula_2 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio19

# 3. With BIO2
crops_formula_3 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio13 + bio14 + bio19

# 4. With BIO4
crops_formula_4 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio4 + bio15 + bio19

# 5. BIO15
crops_formula_5 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio15

# 6. With PETseasonality and BIO2
crops_formula_6 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio19

# 7. With PETseasonality and BIO4
crops_formula_7 <- domestication ~ Aspect + ClayPercentage + PETseasonality + PETWettestQuarter + OrganicCarbon + Slope + bio4 + bio19

# 8. With Bio15 and BIO2
crops_formula_8 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio15 

```

## Model selection with BM

```{r model selection BM}
#load required package
library(phylolm)

#define formula components and run function (forward)
## 1
phylostep(formula = crops_formula_1, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "BM", direction = "forward", trace = 2)
  # domestication ~ 1 + Aspect + ClayPercentage + PhCaCL + Slope
  # AIC(k=2): 94.383384041981

## other formulas
phylostep(formula = crops_formula_8, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "BM", direction = "forward", trace = 2)

#define formula components and run function (both)
## 1
phylostep(formula = crops_formula_1, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "BM", direction = "both", trace = 2)
  # domestication ~ 1 + Aspect + ClayPercentage + PhCaCL + Slope
  # AIC(k=2): 94.383384041981

## other formulas
phylostep(formula = crops_formula_8, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "BM", direction = "both", trace = 2)
```

The forward model with BM has the following model as the best model, with an AIC of... and a LogLik of ...
  domestication ~
The both model with BM has the following model as the best model, with an AIC of ... and a LogLik of ...
  domestication ~

## Model selection with OUrandomRoot

```{r model selection OUrandomRoot}
#define formula components and run function (forward)
## 1
phylostep(formula = crops_formula_1, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "OUrandomRoot", direction = "forward", trace = 2)

## other formulas
phylostep(formula = crops_formula_8, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "OUrandomRoot", direction = "forward", trace = 2)

#define formula components and run function (both)
## 1
phylostep(formula = crops_formula_1, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "OUrandomRoot", direction = "both", trace = 2)

## other formulas
phylostep(formula = crops_formula_8, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "OUrandomRoot", direction = "both", trace = 2)
```

The forward model with OUrandomRoot has the following model as the best model, with an AIC of... and a LogLik of ...
  domestication ~
The both model with OUrandomRoot has the following model as the best model, with an AIC of ... and a LogLik of ...
  domestication ~

## Model selection with OUfixedRoot

```{r model selection OUfixedRoot}
#define formula components and run function (forward)
## 1
phylostep(formula = crops_formula_1, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "OUfixedRoot", direction = "forward", trace = 2)

## other formulas
phylostep(formula = crops_formula_8, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "OUfixedRoot", direction = "forward", trace = 2)

#define formula components and run function (both)
## 1
phylostep(formula = crops_formula_1, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "OUfixedRoot", direction = "both", trace = 2)

## other formulas
phylostep(formula = crops_formula_8, starting.formula = crops_starting_formula, data = GIS_crops, phy = tree_crops, model = "OUfixedRoot", direction = "both", trace = 2)
```

The forward model with OUfixedRoot has the following model as the best model, with an AIC of... and a LogLik of ...
  domestication ~
The both model with OUfixedRoot has the following model as the best model, with an AIC of ... and a LogLik of ...
  domestication ~

## Conclusion
The best model can be made with the following combination:
Formula: 
Model: 
Direction: 