---
title: "Model_selection_crops"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Model selection domestication

The linear model will have to be made, with the trait-presence-column as the dependent variable, and the 
GIS data frames as the predictor variables. The function phylostep of the phylolm package will be used. 

## Define formulas 

### First set of model formuals
Based on the hypothesis. 

```{r}
# 0. Starting formula
crops_starting_formula <- domestication ~ 1 

# 1. Lowest VIF value
crops_formula_1 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio14 + bio19

# 2. With PETseasonality
crops_formula_2 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio16 + bio19

# 3. With BIO2
crops_formula_3 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio13 + bio14 + bio19

# 4. With BIO4
crops_formula_4a <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio4 + bio13 + bio15 + bio19
crops_formula_4b <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio15 + bio16 + bio19
crops_formula_4c <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio4 + bio15 + bio18 + bio19

# 5. With bio5 
crops_formula_5 <- domestication ~ Aspect + BulkDensity + ClayPercentage + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio5 + bio13 + bio14 + bio19

# 6. With bio6
crops_formula_6 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6  + bio13 + bio14 + bio19

# 7. With bio7 
crops_formula_7 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio7 + bio14 + bio19

# 8. BIO15
crops_formula_8a <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio7 + bio13 + bio15
crops_formula_8b <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio15 + bio19
```

### Second set of model formulas 
The rest of the variables that were not included yet. 

```{r}
# 9. With annualPET (6)
crops_formula_9 <- domestication ~ Aspect + BulkDensity + ClayPercentage + annualPET + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio14 + bio19

# 10. With aridityIndexThornthwaite 
crops_formula_10a <- domestication ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + continentality + minTempWarmest + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio19
crops_formula_10b <- domestication ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + maxTempColdest + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio19
crops_formula_10c <- domestication ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio19
crops_formula_10d <- domestication ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio7 + bio13 + bio19
crops_formula_10e <- domestication ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio4 + bio13 + bio19
crops_formula_10f <- domestication ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio13 + bio19
crops_formula_10g <- domestication ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio9 + bio13 + bio19
crops_formula_10h <- domestication ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio11 + bio13 + bio19

# 11. With climaticMoistureIndex 
crops_formula_11 <- domestication ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + minTempWarmest + PETDriestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio19

# 12. With continentality
crops_formula_12 <- domestication ~ Aspect + BulkDensity + ClayPercentage + continentality + minTempWarmest + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio14 + bio19

# 13. With embergerQ 
crops_formula_13a <- domestication ~ Aspect + BulkDensity + ClayPercentage + embergerQ + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio18
crops_formula_13b <- domestication ~ Aspect + BulkDensity + ClayPercentage + continentality + embergerQ + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope
crops_formula_13c <- domestication ~ Aspect + BulkDensity + ClayPercentage + embergerQ + minTempWarmest + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope
crops_formula_13d <- domestication ~ Aspect + BulkDensity + ClayPercentage + embergerQ + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio4
crops_formula_13e <- domestication ~ Aspect + BulkDensity + ClayPercentage + embergerQ + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio7
crops_formula_13f <- domestication ~ Aspect + BulkDensity + ClayPercentage + embergerQ + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13
crops_formula_13g <- domestication ~ Aspect + BulkDensity + ClayPercentage + embergerQ + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15
crops_formula_13h <- domestication ~ Aspect + BulkDensity + ClayPercentage + embergerQ + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio16

# 14. With growingDegDays0 
crops_formula_14a <- domestication ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + growingDegDays0 + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio19
crops_formula_14b <- domestication ~ Aspect + BulkDensity + ClayPercentage + growingDegDays0 + minTempWarmest + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio19
crops_formula_14c <- domestication ~ Aspect + BulkDensity + ClayPercentage + growingDegDays0 + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio19

# 15. With growingDegDays5
crops_formula_15a <- domestication ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + growingDegDays5 + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio19
crops_formula_15b <- domestication ~ Aspect + BulkDensity + ClayPercentage + growingDegDays5 + minTempWarmest + PETWarmestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio19
crops_formula_15c <- domestication ~ Aspect + BulkDensity + ClayPercentage + growingDegDays5 + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio19

# 16. With maxTempColdestMonth 
crops_formula_16 <- domestication ~ Aspect + BulkDensity + ClayPercentage + maxTempColdest + minTempWarmest + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio14 + bio19

# 17. With monthCountByTemp10
crops_formula_17 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio14 + bio19

# 18. With PETColdestQuarter 
crops_formula_18 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETColdestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio14 + bio19

# 19. With PETWarmestQuarter 
crops_formula_19 <- domestication ~ Aspect + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14

# 20. PETWettestQuarter
crops_formula_20 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio13 + bio14 + bio19

# 21. With thermicityIndex
crops_formula_21a <- domestication ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + minTempWarmest + PETWettestQuarter + thermicityIndex + OrganicCarbon + PhCaCL + Slope + bio14 + bio19
crops_formula_21b <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWarmestQuarter + PETWettestQuarter + thermicityIndex + OrganicCarbon + PhCaCL + Slope + bio14 + bio19
crops_formula_21c <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWettestQuarter + thermicityIndex + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio19

# 22. With bio1
crops_formula_22a <- domestication ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio1 + bio14 + bio19
crops_formula_22b <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio1 + bio14 + bio19
crops_formula_22c <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio1 + bio2 + bio14 + bio19

# 23. With bio3 
crops_formula_23a <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio3 + bio13 + bio14 + bio19
crops_formula_23b <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio3 + bio13 + bio14 + bio19

# 24. With bio8 
crops_formula_24 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio8 + bio13 + bio14 + bio19

# 25. With bio9 
crops_formula_25 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio9 + bio13 + bio14 + bio19

# 26. With bio10 
crops_formula_26 <- domestication ~ Aspect + BulkDensity + ClayPercentage + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio10 + bio13  + bio14 + bio19

# 27. With bio11
crops_formula_27 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio11  + bio13 + bio14 + bio19

# 28. With bio12 
crops_formula_28 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio12 + bio19

# 29. With bio16 
crops_formula_29 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio16 + bio19

# 30. With bio17
crops_formula_30 <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio13 + bio17

# 31. With bio18 
crops_formula_31a <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio18
crops_formula_31b <- domestication ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio18
```

## Change data to match the wanted data format

There are some flaws in the GIS data frame that make it unfit for the phylostep function. First of all
the row names have to match the tip.lables of the tree, and secondly the trait value (0/1) has to be
seen as a numeric value. 
```{r}
#make allmb_name column into the row names of this table
rownames(GIS_crops) <- GIS_crops[,1]
GIS_crops[,1] <- NULL

#make sure the values in the 'woody' column are seen as numeric values
GIS_crops$domestication <- as.numeric(GIS_crops$domestication)
#check whether it worked
is.numeric(GIS_crops$domestication)
```

## Perform model selection
With the phyloglmstep() function by Rutger Vos.

```{r}
#source function and load required package
library(phylolm)
source('phyloglmstep.R')

#fill in arguments and execute function
phyloglmstep(crops_formula_4a, starting.formula = NULL, data=GIS_crops, phy=tree_crops, 
                         method="logistic_MPLE",
                         direction = "forward", trace = 0,
                         btol = 100, log.alpha.bound = 8, start.beta=NULL, 
                         start.alpha=NULL, boot = 0, full.matrix = TRUE,
                         k=2)
```