---
title: "Model_selection_woody"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Model selection derived woodiness

The linear model will have to be made, with the trait-presence-column as the dependent variable, and the 
GIS data frames as the predictor variables. The function phylostep of the phylolm package will be used. 

## Results VIF analysis

```{r VIF analysis}
# 1. Without ClimaticMoistureIndex: 
vif(GIS_woody[,c(3:5, 14, 15, 17, 18, 20, 22:24, 38, 39, 42, 43)])

# 2. With ClimaticMoistureIndex:
vif(GIS_woody[,c(3:5, 8, 14, 17, 18, 20, 22:24, 39, 43)])

# 3. With BIO6
vif(GIS_woody[,c(3:5, 14, 15, 17, 20, 22:24, 30, 38, 39, 42, 43)]) 

# 4. With BIO6 and ClimaticMoistureIndex:
vif(GIS_woody[,c(3:5, 8, 14, 17, 20, 22:24, 30, 39, 43)]) 


################### New, more elaborate selection:

# 1. With ClimaticMoistureIndex
vif(GIS_woody[,c(3:5, 8, 14, 17, 18, 20, 22:24, 39, 43)])

# 2. With BIO6
vif(GIS_woody[,c(3:5, 14, 15, 17, 20, 22:24, 30, 38, 39, 42, 43)]) 

# 3. With growingDegDays0
vif(GIS_woody[,c(3:5, 11, 17, 18, 20, 22:24, 39, 42, 43)])

# 4. With aridityIndexThornthwaite
vif(GIS_woody[,c(3:5, 7, 14, 15, 17, 18, 20, 22:24, 39, 43)])

# 5. With embergerQ
vif(GIS_woody[,c(3:5, 10, 14, 15, 17, 20, 22:24, 39, 43)])

# 6. With BIO6 and ClimaticMoistureIndex
vif(GIS_woody[,c(3:5, 8, 14, 17, 20, 22:24, 30, 39, 43)]) 

# 7. With BIO6 and aridityIndexThornthwaite
vif(GIS_woody[,c(3:5, 7, 14, 17, 20, 22:24, 30, 39, 43)]) 

# 8. With BIO6 and embergerQ
vif(GIS_woody[,c(3:5, 10, 14, 15, 17, 20, 22:24, 30, 39, 43)])

# 9. With growingDegDays0 and ClimaticMoistureIndex
vif(GIS_woody[,c(3:5, 8, 11, 17, 18, 20, 22:24, 39, 43)])

# 10. With growingDegDays0 and aridityIndexThornthwaite
vif(GIS_woody[,c(3:5, 7, 11, 14, 20, 22:24, 39, 43)])

# 11. With growingDegDays0 and embergerQ
vif(GIS_woody[,c(3:5, 10, 11, 17, 20, 22:24, 39, 43)])

# 12. Solely based on highest VIF value 
vif(GIS_woody[,c(3:5, 14, 15, 17, 18, 20, 22:24, 38, 39, 42, 43)])

```

## Change data to match the wanted data format

There are some flaws in the GIS data frame that make it unfit for the phylostep function. First of all
the row names have to match the tip.lables of the tree, and secondly the trait value (0/1) has to be
seen as a numeric value. 
```{r}
#make allmb_name column into the row names of this table
rownames(GIS_woody) <- GIS_woody[,1]
GIS_woody[,1] <- NULL

#make sure the values in the 'woody' column are seen as numeric values
GIS_woody$woody <- as.numeric(GIS_woody$woody)
#check whether it worked
is.numeric(GIS_woody$woody)
```

## Define formulas

```{r define formula}
# 0. Starting formula
woody_starting_formula <- woody ~ 1 

# 1. Solely based on lowest VIF value 
woody_formula_1 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19

# 2. With ClimaticMoistureIndex:
woody_formula_2 <- woody ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + minTempWarmest + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio19

# 3. With BIO6
woody_formula_3 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio14 + bio15 + bio18 + bio19

# 4. With growingDegDays0
woody_formula_4 <- woody ~ Aspect + BulkDensity + ClayPercentage + growingDegDays0 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio18 + bio19

# 5. With aridityIndexThornthwaite
woody_formula_5 <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio19

# 6. With embergerQ
woody_formula_6 <- woody ~ Aspect + BulkDensity + ClayPercentage +embergerQ + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio19

# 7. With BIO6, with ClimaticMoistureIndex:
woody_formula_7 <- woody ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio15 + bio19

# 8. With BIO6 and aridityIndexThornthwaite
woody_formula_8 <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio15 + bio19

# 9. With BIO6 and embergerQ
woody_formula_9 <- woody ~ Aspect + BulkDensity + ClayPercentage + embergerQ + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio15 + bio19

# 10. With growingDegDays0 and ClimaticMoistureIndex
woody_formula_10 <- woody ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + growingDegDays0 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio19

# 11. With growingDegDays0 and aridityIndexThornthwaite
woody_formula_11 <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + growingDegDays0 + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio19

# 12. With growingDegDays0 and embergerQ
woody_formula_12 <- woody ~ Aspect + BulkDensity + ClayPercentage + embergerQ + growingDegDays0 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio19

```

## Model selection with BM 

BM stands for Brownian motion. 

```{r model selection BM}
#load required package
library(phylolm)

#define formula components and run function (forward)
phylostep(formula = woody_formula_1, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "BM", direction = "forward", trace = 2)
 
#define formula components and run function (both)
phylostep(formula = woody_formula_1, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "BM", direction = "both", trace = 2)

```

## Model selection with OUrandomRoot

OUrandomRoot stands for Ornstein-Uhlenbeck with a random root. 

```{r model selection OUrandomRoot}
#define formula components and run function (forward)
phylostep(formula = woody_formula_1, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "OUrandomRoot", direction = "forward", trace = 2)

#define formula components and run function (both)
phylostep(formula = woody_formula_1, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "OUrandomRoot", direction = "both", trace = 2)
  
```

## Model selection with OUfixedRoot

OUfixedRoot stands for Ornstein-Uhlenbeck with a fixed root. 

```{r model selection OUfixedRoot}
#define formula components and run function (forward)
phylostep(formula = woody_formula_1, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "OUfixedRoot", direction = "forward", trace = 2)
 
#define formula components and run function (both)
phylostep(formula = woody_formula_1, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "OUfixedRoot", direction = "both", trace = 2)

```

## Conclusion
The best model can be made with the following combination:
Formula: 
Model: 
Direction: 
