---
title: "Model_selection_woody"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Model selection derived woodiness

The linear model will have to be made, with the trait-presence-column as the dependent variable, and the 
GIS data frames as the predictor variables. The function phylostep of the phylolm package will be used. 

## Define formulas 

### First set of model formuals
Based on the hypothesis. 

```{r}
# 0. Starting formula
woody_starting_formula <- woody ~ 1 

# 1. Solely based on lowest VIF value 
woody_formula_1 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19

# 2. With ClimaticMoistureIndex
woody_formula_2 <- woody ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + minTempWarmest + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio19

# 3. With BIO6 (30) 
woody_formula_3 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio14 + bio15 + bio18 + bio19

# 4. With growingDegDays0
woody_formula_4 <- woody ~ Aspect + BulkDensity + ClayPercentage + growingDegDays0 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio18 + bio19

# 5. With aridityIndexThornthwaite
woody_formula_5a <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio15 + bio19
woody_formula_5b <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio16 + bio19
woody_formula_5c <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio19

# 6. With embergerQ
woody_formula_6a <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + continentality + embergerQ + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio19
woody_formula_6b <- woody ~ Aspect + BulkDensity + ClayPercentage +embergerQ + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio4 + bio14 + bio15 + bio19
woody_formula_6c <- woody ~ Aspect + BulkDensity + ClayPercentage +embergerQ + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio15 + bio17 + bio19
woody_formula_6d <- woody ~ Aspect + BulkDensity + ClayPercentage +embergerQ + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio7 + bio15 + bio19

# 7. With BIO6 and ClimaticMoistureIndex
woody_formula_7 <- woody ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio15 + bio19

# 8. With BIO6 and aridityIndexThornthwaite
woody_formula_8a <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + embergerQ + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio15 + bio19
woody_formula_8b <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio13 + bio15 + bio19
woody_formula_8c <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio15 + bio16 + bio19

# 9. With BIO6 and embergerQ
woody_formula_9a <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + embergerQ + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio15 + bio19
woody_formula_9b <- woody ~ Aspect + BulkDensity + ClayPercentage + embergerQ + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio14 + bio15 + bio19
woody_formula_9c <- woody ~ Aspect + BulkDensity + ClayPercentage + embergerQ + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio6 + bio15 + bio17 + bio19

# 10. With growingDegDays0 and ClimaticMoistureIndex
woody_formula_10 <- woody ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + growingDegDays0 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio19

# 11. With growingDegDays0 and aridityIndexThornthwaite
woody_formula_11a <- woody ~ Aspect + BulkDensity + ClayPercentage + embergerQ + aridityIndexThornthwaite + growingDegDays0 + minTempWarmest + PETWarmestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio19
woody_formula_11b <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + growingDegDays0 + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio13 + bio15 + bio19
woody_formula_11c <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + growingDegDays0 + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio16 + bio19
woody_formula_11d <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + growingDegDays0 + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio18 + bio19

# 12. With growingDegDays0 and embergerQ
woody_formula_12a <- woody ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + continentality + embergerQ + growingDegDays0 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio19
woody_formula_12b <- woody ~ Aspect + BulkDensity + ClayPercentage + embergerQ + growingDegDays0 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio4 + bio15 + bio19
woody_formula_12c <- woody ~ Aspect + BulkDensity + ClayPercentage + embergerQ + growingDegDays0 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio3 + bio14 + bio15 + bio19
woody_formula_12d <- woody ~ Aspect + BulkDensity + ClayPercentage + embergerQ + growingDegDays0 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio4 + bio15 + bio17 + bio19
woody_formula_12e <- woody ~ Aspect + BulkDensity + ClayPercentage + embergerQ + growingDegDays0 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio5 + bio15 + bio19
```

### Second set of model formulas 
The rest of the variables that were not included yet. 

```{r}
# 13. With annualPET
woody_formula_13 <- woody ~ Aspect + BulkDensity + ClayPercentage + annualPET + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19

# 14. With continentality
woody_formula_14a <- woody ~ Aspect + BulkDensity + ClayPercentage + annualPET + continentality + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19
woody_formula_14b <- woody ~ Aspect + BulkDensity + ClayPercentage + continentality + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19
woody_formula_14c <- woody ~ Aspect + BulkDensity + ClayPercentage + continentality + minTempWarmest + monthCountByTemp10 + PETDriestQuarter +  PETWarmestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19
woody_formula_14d <- woody ~ Aspect + BulkDensity + ClayPercentage + continentality + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19
woody_formula_14e <- woody ~ Aspect + BulkDensity + ClayPercentage + continentality + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio15 + bio18 + bio19
woody_formula_14f <- woody ~ Aspect + BulkDensity + ClayPercentage + continentality + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + OrganicCarbon + PhCaCL + Slope + bio3 + bio14 + bio15 + bio18 + bio19
woody_formula_14g <- woody ~ Aspect + BulkDensity + ClayPercentage + continentality + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + OrganicCarbon + PhCaCL + Slope + bio8 + bio14 + bio15 + bio18 + bio19

# 15. With growingDegDays0
woody_formula_15 <- woody ~ Aspect + BulkDensity + ClayPercentage + growingDegDays5 + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio18 + bio19

# 16. With maxTempColdest
woody_formula_16 <- woody ~ Aspect + BulkDensity + ClayPercentage + maxTempColdest + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19
 
# 17. With PETColdestQuarter
woody_formula_17 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETColdestQuarter + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19

# 18. With PETWarmestQuarter
woody_formula_18 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWarmestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19

# 19. With thermicityIndex
woody_formula_19 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + thermicityIndex + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio18 + bio19

# 20. With bio1
woody_formula_20 <- woody ~ Aspect + BulkDensity + ClayPercentage + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio1 + bio14 + bio15 + bio18 + bio19

# 21. With bio2
woody_formula_21 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio15 + bio18 + bio19

# 22. With bio3
woody_formula_22 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio3 + bio14 + bio15 + bio18 + bio19

# 23. With bio4
woody_formula_23 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio4 + bio14 + bio15 + bio18 + bio19

# 24. With bio5
woody_formula_24a <- woody ~ Aspect + BulkDensity + ClayPercentage + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio5 + bio14 + bio15 + bio18 + bio19
woody_formula_24b <- woody ~ Aspect + BulkDensity + ClayPercentage + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio5 + bio6 + bio15 + bio18 + bio19

# 25. With bio7
woody_formula_25 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio7 + bio14 + bio15 +  bio18 + bio19

# 26. With bio8 
woody_formula_26 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + OrganicCarbon + PhCaCL + Slope + bio8 + bio14 + bio15 + bio18 + bio19

# 27. With bio9
woody_formula_27 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio9 + bio14 + bio15 + bio18 + bio19

# 28. With bio10
woody_formula_28 <- woody ~ Aspect + BulkDensity + ClayPercentage + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio10 + bio14 + bio15 + bio18 + bio19

# 29. With bio11
woody_formula_29 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio11 + bio14 + bio15 + bio18 + bio19

# 30. With bio12
woody_formula_30 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio12 + bio15 + bio19

# 31. With bio13
woody_formula_31 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio14 + bio15 + bio19

# 32. With bio16
woody_formula_32 <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio14 + bio15 + bio16 + bio19

# 33. With bio17
woody_formula_33a <- woody ~ Aspect + BulkDensity + ClayPercentage + minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio13 + bio15 + bio17 + bio19
woody_formula_33b <- woody ~ Aspect + BulkDensity + ClayPercentage+ minTempWarmest + monthCountByTemp10 + PETDriestQuarter + PETseasonality + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio15 + bio16 + bio17 + bio19
```

## Change data to match the wanted data format

There are some flaws in the GIS data frame that make it unfit for the phylostep function. First of all
the row names have to match the tip.lables of the tree, and secondly the trait value (0/1) has to be
seen as a numeric value. 
```{r}
#make allmb_name column into the row names of this table
rownames(GIS_woody) <- GIS_woody[,1]
GIS_woody[,1] <- NULL

#make sure the values in the 'woody' column are seen as numeric values
GIS_woody$woody <- as.numeric(GIS_woody$woody)
#check whether it worked
is.numeric(GIS_woody$woody)
```

## Preform model selection
With the phyloglmstep() function by Rutger Vos.

```{r}
#source function and load required package
source('phyloglmstep.R')
library(phylolm)

#fill in arguments and execute function
phyloglmstep(woody_formula_33b, starting.formula = NULL, data=GIS_woody, phy=tree_woody, 
                         method="logistic_MPLE",
                         direction = "forward", trace = 2,
                         btol = 100, log.alpha.bound = 8, start.beta=NULL, 
                         start.alpha=NULL, boot = 0, full.matrix = TRUE,
                         k=2)
z```