---
title: "Pipeline_myco"
output: html_document
---

```{r setup, i---
title: "Pipeline"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Pipeline of the necessary steps in R and outside of R

## 1 Import data

First the traits table and the original phylogeny (tree) have to be read into R in the right format 
to be able to use them. 

### The project root and the main tree
```{r define data root and import tree}
#load required package
library(ape)

#import and define data
project_root <- 'C:/Users/Esther/Documents/Master/Stage/'
tree_file <- paste(c(project_root, 'ALLMB.tre'), collapse = '')
tree <- read.tree(tree_file)
```

### Import mycorrhiza data
```{r import mycorrhiza data}
#import and define trait table 
myco_tsv_file <- paste(c(project_root, 'sdmdl-angiosperm-data/data/myco_taxa.tsv'), collapse = '')
myco_tsv <- read.table(myco_tsv_file, sep = '\t', header = TRUE)
```

## 2 Recode dependent variables to binary form

The data of the dependent characters (the plant traits) have to be recoded. This is because a binary 
input is prefered for this analysis due to technical reasons. BUt at the moment, the three plant 
traits all have 3 or 4 states.
For mycorrhiza, there are 4 trait-states: EM, AM, both, neither. This will be turned into: AM (0) and 
the rest together (1). So in this case nothing will be eliminated from the analysis, just (1)(2)(3) 
have to be collapsed together. 

```{r collapse state 1, 2 and 3}
#create new column
myco_tsv["myco_state_int"] <- NA

#add binary values to column based on names 
myco_tsv$myco_state_int[myco_tsv$myco_state == "arbuscular"] <- "0"
myco_tsv$myco_state_int[myco_tsv$myco_state == "ecto"] <- "1"
myco_tsv$myco_state_int[myco_tsv$myco_state == "both"] <- "1"
myco_tsv$myco_state_int[myco_tsv$myco_state == "neither"] <- "1"

#now myco_tsv contains an added column with binary state values
```

## 3 Tidy data

### Match the names 
Make the taxa names the same in the table as in the tree; the difference is that the tsv file has 
white spaces where the tree file has underscores (_). 
```{r match names}
#load required package
library(stringi)

#replace the whitespaces in the taxa names with '_' 
myco_tsv$allmb_name <- stri_replace_all_charclass(myco_tsv$allmb_name, "\\p{WHITE_SPACE}", "_")
```

## 4 Combine the data

The trait states, the selected subset of taxa names, and the GIS information all have to be 
combined into one file. 

### Import GIS data
Import the data and make sure the taxa names are the same in this table as in the tsv file and 
in the tree file, so with _ between the words instead of whitespaces. 
```{r import GIS data}
#import GIS data 
GIS_file <- paste(c(project_root, 'niche_traits_merged.csv'), collapse = '')
GIS <- read.table(GIS_file, sep = ',', header = TRUE)

#change names of GIS data 
GIS_names <- as.character(unlist(c(GIS[1])))
GIS_names <- gsub('  ', '. ', GIS_names)
GIS_names <- stri_replace_all_charclass(GIS_names, "\\p{WHITE_SPACE}", "_")
#there is a discrepancy with several names that should have a - instead of a _ in some places. 

#change those specific GIS names:
GIS_names[GIS_names == "Acaena_novae_zelandiae"] <- "Acaena_novae-zelandiae"
GIS_names[GIS_names == "Impatiens_noli_tangere"] <- "Impatiens_noli-tangere"  
GIS_names[GIS_names == "Myrrhidendron_donnell_smithii" ] <- "Myrrhidendron_donnell-smithii" 
GIS_names[GIS_names == "Silene_flos_cuculi"] <- "Silene_flos-cuculi"
GIS_names[GIS_names == "Pithecellobium_unguis_cati"] <- "Pithecellobium_unguis-cati"
GIS_names[GIS_names == "Hyacinthoides_non_scripta" ] <- "Hyacinthoides_non-scripta" 
GIS_names[GIS_names == "Alisma_plantago_aquatica"] <- "Alisma_plantago-aquatica" 
GIS_names[GIS_names == "Apera_spica_venti"] <- "Apera_spica-venti"
GIS_names[GIS_names == "Arctostaphylos_uva_ursi"] <- "Arctostaphylos_uva-ursi"
GIS_names[GIS_names == "Blitum_bonus_henricus"] <- "Blitum_bonus-henricus"
GIS_names[GIS_names == "Begonia_meyeri_johannis"] <- "Begonia_meyeri-johannis"

#put the changed names back into the GIS table
GIS$allmb_name <- GIS_names
```

### Tidy data
Make the right selection of taxa out of the GIS variables table and the tree. 
```{r trim GIS table}
#select the taxa that should be kept in the GIS file
keep_GIS_myco <- as.character(unlist(c(myco_tsv[1]))) 
#define the final GIS file for mycorrhiza
GIS_myco <- GIS[GIS$allmb_name %in% keep_GIS_myco,]

#select the names that have to be kept in the tree
keep_tips_myco <- as.character(unlist(c(GIS_myco[1]))) 
#define the final tree for mycorrhiza
tips_to_drop_myco <- setdiff(tree$tip.label, keep_tips_myco)
tree_myco <- drop.tip(tree, tips_to_drop_myco)
```

### Add the binary trait state to the GIS table
```{r add binary state to GIS table}
#make new tables for both mycorrhiza states separately 
GIS_myco_0 <- myco_tsv[myco_tsv$myco_state_int == "0",]
GIS_myco_1 <- myco_tsv[myco_tsv$myco_state_int == "1",]

#add a column to GIS_myco that contains the binary trait values
GIS_myco$mycorrhiza[GIS_myco$allmb_name %in% GIS_myco_0$allmb_name] <- "0"
GIS_myco$mycorrhiza[GIS_myco$allmb_name %in% GIS_myco_1$allmb_name] <- "1"

#change order of the columns for esthetics
GIS_myco <- GIS_myco[,c(1,43, 2:42)]

#there are less plants in this GIS file than in the tree or the tsv file, this makes sense as it 
#could be the case that there are not enough occurrences for each plant in the ALLMB tree, and 
#thus not enough GIS data. 
```

## 5 Correct for colinearity in the predictor variables

There will probably be some colinearity in the GIS-data. This could lead to instability and bias 
in the model, so a correction will have to be done. This can be done using the variance inflation 
factor (of the usdm package) or a correlation matrix. 
Only the combinations that result in the best models are shown, see VIF_analysis_myco" for all 
the used combinations.

### VIF-analysis
```{r VIF analysis}
#load required package
library(usdm)

#fill in the arguments of the function and execute it 
vif(GIS_myco[,c(3:43)])

#choose the variables to delete that have a higher VIF value than 10, but make sure it makes 
#sense in terms of biology, results in:
# 3. With continentality (no alpha warning)
vif(GIS_myco[,c(3:5, 7, 9, 14, 20, 22:24, 26, 39, 42, 43)])

# 4. With embergerQ (alpha warning)
vif(GIS_myco[,c(3:5, 7, 10, 14, 20, 22:24, 26, 28, 39, 42)])
```

## 6 Model selection: Define formula and select model 

The linear model will have to be made, with the trait-presence-column as the dependent variable, 
and the GIS data frames as the predictor variables. The function phylostep of the phylolm 
package will be used. 

### Define input formulas
Define the formulas that will be used as input formula in the phylostep function, based on the 
results of the VIF analysis. Only the formulas that result in the best models are shown, see 
"Model_selection_crops_elaborated" for all the used formulas. 

```{r define formula}
# Formula 3. With continentality (no alpha warning AIC 1037,3)
myco_formula_3 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + continentality + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio15 + bio18 + bio19

# Formula 4. With embergerQ (alpha warning, AIC 1026,5) 
myco_formula_4 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + embergerQ + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio15 + bio18
```

### Change data to match the wanted data format
There are some flaws in the GIS data frame that make it unfit for the phylostep function. First of all
the row names have to match the tip.lables of the tree, and secondly the trait value (0/1) has to be
seen as a numeric value. 
```{r}
#make allmb_name column into the row names of this table
rownames(GIS_myco) <- GIS_myco[,1]
GIS_myco[,1] <- NULL

#make sure the values in the 'woody' column are seen as numeric values
GIS_myco$mycorrhiza <- as.numeric(GIS_myco$mycorrhiza)
#check whether it worked
is.numeric(GIS_myco$mycorrhiza)
```

### Perform model selection
What formula gives us the best result (in terms of the AIC value)?
```{r}
#source function and load package
library(phylolm)
source('phyloglmstep.R')

#fill in arguments and execute function
## no alpha warning
phyloglmstep(myco_formula_3, starting.formula = NULL, data=GIS_myco, phy=tree_myco, 
             method="logistic_MPLE", direction = "both", trace = 0, btol = 100, 
             log.alpha.bound = 8, start.beta=NULL, start.alpha=NULL, boot = 0, 
             full.matrix = TRUE, k=2)

## with alpha warning
phyloglmstep(myco_formula_4, starting.formula = NULL, data=GIS_myco, phy=tree_myco, 
             method="logistic_MPLE", direction = "both", trace = 0, btol = 100, 
             log.alpha.bound = 8, start.beta=NULL, start.alpha=NULL, boot = 0, 
             full.matrix = TRUE, k=2)
```

## 7 Phylogenetic generalized linear modelling analysis

After this, the phylogenetic generalized linear modeling analysis will be done, which 
optimizes the model. There is a function in the phylolm package that allows us to 
do this: phyloglm. 

```{r phylogenetic generalized linear modelling}
#construct model
myco_best_model_no_alpha <- mycorrhiza ~ 1 + BulkDensity + ClayPercentage + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + bio2 + bio15 + bio19

myco_best_model_alpha <- mycorrhiza ~ 1 + ClayPercentage + bio2 + bio15

#fill in the arguments for the phyloglm function and execute it
model_myco <- phyloglm(formula = myco_best_model_no_alpha, data = GIS_myco, phy = tree_myco, 
                       method = "logistic_MPLE", btol = 100, log.alpha.bound = 8)
summary(model_myco)

model_myco <- phyloglm(formula = myco_best_model_alpha, data = GIS_myco, phy = tree_myco, 
                       method = "logistic_MPLE", btol = 100, log.alpha.bound = 8)
summary(model_myco)
```

Coefficients:
                   Estimate    StdErr z.value   p.value    
(Intercept)       -0.545280  0.124625 -4.3754 1.212e-05 ***
BulkDensity       -0.083022  0.328782 -0.2525  0.800644    
ClayPercentage    -0.644497  0.248637 -2.5921  0.009539 ** 
minTempWarmest    -0.011444  0.237225 -0.0482  0.961525    
PETWettestQuarter  0.140902  0.199277  0.7071  0.479526    
OrganicCarbon      0.198053  0.428093  0.4626  0.643621    
PhCaCL             0.105180  0.217995  0.4825  0.629458    
bio2              -0.665559  0.205055 -3.2458  0.001171 ** 
bio15              0.503568  0.183735  2.7407  0.006130 ** 
bio19              0.048930  0.127545  0.3836  0.701252    

### Second round
As BulkDensity, minTempWarmest, PETWettestQuarter, OrganicCarbon, PhCaCL, and bio19 were all non 
significant, they were left out, which leads to the same model as myco_best_model_alpha. So, both 
these formulas can be used, they result in the same final model, with the following significance 
values:
Coefficients:
               Estimate   StdErr z.value   p.value    
(Intercept)    -0.58705  0.11215 -5.2346 1.654e-07 ***
ClayPercentage -0.57211  0.16058 -3.5628 0.0003670 ***
bio2           -0.64992  0.14236 -4.5655 4.984e-06 ***
bio15           0.55083  0.15636  3.5228 0.0004271 ***

### Define final model
```{r}
final_model_myco <- mycorrhiza ~ - 0.58705 - 0.57211*ClayPercentage -0.64992*bio2 + 0.55083*bio15
```

## 8 Model evaluation

This is needed to see whether the best models we constructed, are still in compliance with the 
assumptions that have to be met when performing a GLM. There is a package that can do this for you, 
called 'performance'. We'll do a comprehensive model check, with function 'check_model()'. However, 
this function is not appropriate for models that include a phylogeny, so in this case, a phyloGLM is 
not seen as a subclass of GLMs. We will still do it while keeping out the phylogeny, but keep in mind 
this might result in skewed interpretation.

```{r}
#load required packages
library(performance)
library(see)
library(gridExtra)

#construct model without phylogeny
model_check_myco <- glm(mycorrhiza ~ 1 + ClayPercentage + bio2 + bio15, data = GIS_myco, 
                        family = binomial())

#execute function
par(mar=c(5,5,5,5))
check_model(model_check_myco)
```

1. Check for multicolinearity
All variables have low correlation. 
2. Non-normality of residuals and outliers
...
3. Homogeneity of variance (scale-location)
...
4. Check for influential observations
...

## 9 Plot results

```{r boxplots}
#load required package 
library(dplyr)

#define two subsets
GIS_filtered_AM <- GIS_myco %>%
    select(mycorrhiza, ClayPercentage, bio2, bio15) %>%
    subset(mycorrhiza == 0)
GIS_filtered_EM <- GIS_myco %>%
    select(mycorrhiza, ClayPercentage, bio2, bio15) %>%
    subset(mycorrhiza == 1)

#change margins
par(mar=c(3,9,3,3))

#make boxplots
boxplot(GIS_filtered_EM$ClayPercentage,GIS_filtered_AM$ClayPercentage,
    main = "Clay percentage",
    at = c(1,2),
    names = c("ectomycorrhiza", "arbuscular mycorrhiza"),
    las = 2,
    col = c("cornflowerblue","lightcoral"),
    border = "gray29",
    horizontal = TRUE,
    notch = TRUE
)

boxplot(GIS_filtered_EM$bio2,GIS_filtered_AM$bio2,
    main = "Mean diurnal range (mean of monthly (max temp - min temp))",
    at = c(1,2),
    names = c("ectomycorrhiza", "arbuscular mycorrhiza"),
    las = 2,
    col = c("cornflowerblue","lightcoral"),
    border = "gray29",
    horizontal = TRUE,
    notch = TRUE
)

boxplot(GIS_filtered_EM$bio15,GIS_filtered_AM$bio15, 
    main = "Precipitation seasonality (coefficient of variation)",
    at = c(1,2),
    names = c("ectomycorrhiza", "arbuscular mycorrhiza"),
    las = 2,
    col = c("cornflowerblue","lightcoral"),
    border = "gray29",
    horizontal = TRUE,
    notch = TRUE
)
```

## 10 Biological interpretation

In the end, the final models and the predictor variables that are included will have to be 
interpreted biologically. With this knowledge, the thesis can be completed. 