---
title: "Pipeline_woody"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Pipeline of the necessary steps in R and outside of R

## 1 Import data

First the traits table and the original phylogeny (tree) have to be read into R in the right format to be able to use them. 

### The project root and the main tree
```{r define data root and import tree}
#load required package
library(ape)

#import and define data
project_root <- 'C:/Users/Esther/Documents/Master/Stage/'
tree_file <- paste(c(project_root, 'ALLMB.tre'), collapse = '')
tree <- read.tree(tree_file)
```

### Import derived woodiness data
```{r import derived woodiness data}
#import and define trait tables
woody_tsv_file <- paste(c(project_root, 'sdmdl-angiosperm-data/data/woody.tsv'), collapse = '')
woody_tsv <- read.table(woody_tsv_file, sep = '\t', header = TRUE)
```

## 2 Recode dependent variables to binary form

The data of the dependent characters (the plant traits) have to be recoded. This is because a binary 
input is prefered for this analysis due to technical reasons. BUt at the moment, the three plant traits 
all have 3 or 4 states For secondary woodiness, there are 3 trait-states: derived woody (0), non-woody (1), 
ancestral woody (2). 
This will be turned into: derived woody (0) and non-woody (1). Because we want to know more about this 
precise shift between non-woody and derived woodiness. So all the plants with ancestral woodiness (3) as 
trait, have to be eliminated from the analysis. 

```{r cut out state 2}
#define new sub-woody, which contains binary trait values (tsv)
woody_tsv_bi <- subset(woody_tsv, (woody_tsv$woody == 0 | woody_tsv$woody ==1))
```

## 3 Tidy data

### Match the names
Make the taxa names the same in the table as in the tree; the difference is that the tsv file has white spaces where the tree file has underscores (_). 
```{r match names}
#load required package 
library(stringi)

#replace the spaces in the taxa names with '_' 
woody_tsv_bi$allmb_name <- stri_replace_all_charclass(woody_tsv_bi$allmb_name, "\\p{WHITE_SPACE}", "_")
```

## 4 Combine the data

The trait states, the selected subset of taxa names, and the GIS information all have to be combined into one file. 

### Import GIS data
Import the data and make sure the taxa names are the same in this table as in the tsv file and in the tree file,
so with _ between the words instead of whitespaces. 
```{r import GIS data}
#import GIS data
GIS_file <- paste(c(project_root, 'niche_traits_merged.csv'), collapse = '')
GIS <- read.table(GIS_file, sep = ',', header = TRUE)

#change names of GIS data 
GIS_names <- as.character(unlist(c(GIS[1])))
GIS_names <- gsub('  ', '. ', GIS_names)
GIS_names <- stri_replace_all_charclass(GIS_names, "\\p{WHITE_SPACE}", "_")
#there is a discrepancy with several names that should have a - instead of a _ in some places. 

#change those specific GIS names:
GIS_names[GIS_names == "Acaena_novae_zelandiae"] <- "Acaena_novae-zelandiae"
GIS_names[GIS_names == "Impatiens_noli_tangere"] <- "Impatiens_noli-tangere"  
GIS_names[GIS_names == "Myrrhidendron_donnell_smithii" ] <- "Myrrhidendron_donnell-smithii" 
GIS_names[GIS_names == "Silene_flos_cuculi"] <- "Silene_flos-cuculi"
GIS_names[GIS_names == "Pithecellobium_unguis_cati"] <- "Pithecellobium_unguis-cati"
GIS_names[GIS_names == "Hyacinthoides_non_scripta" ] <- "Hyacinthoides_non-scripta" 
GIS_names[GIS_names == "Alisma_plantago_aquatica"] <- "Alisma_plantago-aquatica" 
GIS_names[GIS_names == "Apera_spica_venti"] <- "Apera_spica-venti"
GIS_names[GIS_names == "Arctostaphylos_uva_ursi"] <- "Arctostaphylos_uva-ursi"
GIS_names[GIS_names == "Blitum_bonus_henricus"] <- "Blitum_bonus-henricus"
GIS_names[GIS_names == "Begonia_meyeri_johannis"] <- "Begonia_meyeri-johannis"

#put the changed names back into the GIS table
GIS$allmb_name <- GIS_names
```

### Tidy data
Make the right selection of taxa out of the GIS variables table and the tree. 
```{r trim GIS table}
#select the taxa that should be kept in the GIS file
keep_GIS_woody <- as.character(unlist(c(woody_tsv_bi[1])))
#define the final GIS file for woodiness
GIS_woody <- GIS[GIS$allmb_name %in% keep_GIS_woody,]

#select the names that have to be kept in the tree based on the GIS file
keep_tips_woody <- as.character(unlist(c(GIS_woody[1]))) 
#define the final tree for mycorrhiza
tips_to_drop_woody <- setdiff(tree$tip.label, keep_tips_woody)
tree_woody <- drop.tip(tree, tips_to_drop_woody)
```

### Add the binary trait state to the GIS table
```{r add binary state to GIS table}
#make new tables for both woodiness states separately 
GIS_woody_0 <- woody_tsv_bi[woody_tsv_bi$woody == "0",]
GIS_woody_1 <- woody_tsv_bi[woody_tsv_bi$woody == "1",]

#add a column to GIS_woody that contains the binary trait values
GIS_woody$woody[GIS_woody$allmb_name %in% GIS_woody_0$allmb_name] <- "0"  
GIS_woody$woody[GIS_woody$allmb_name %in% GIS_woody_1$allmb_name] <- "1"  

#change order of the columns for esthetics
GIS_woody <- GIS_woody[,c(1,43, 2:42)]

#there are less plants in this GIS file than in the tree or the tsv file, this makes sense as it could be the case that there are not enough occurrences for each plant in the ALLMB tree, and thus not enough GIS data. 
```

## 5 Correct for colinearity in the predictor variables

There will probably be some colinearity in the GIS-data. This could lead to instability and bias in the model, 
so a correction will have to be done. This can be done using the variance inflation factor (of the usdm package)
or a correlation matrix. 
Only the combination that results in the best model is shown, see VIF_analysis_myco" for all the used combinations.

### VIF-analysis
```{r VIF analysis}
#load required package
install.packages("usdm")
library(usdm)

#fill in the arguments of the function and execute it 
vif(GIS_woody[,c(3:43)])

#choose variables to delete that have a higher VIF value than 10, but make sure it makes sense in terms of biology, resulting in:
# 25a. With bio5
vif(GIS_woody[,c(3:5, 15, 17, 20, 22:24, 26, 29, 38, 39, 42, 43)])
```

## 6 Model selection: Define formula and select model 

The linear model will have to be made, with the trait-presence-column as the dependent variable, and the 
GIS data frames as the predictor variables. The function phylostep of the phylolm package will be used. 

### Define input formulas
Define the formulas that will be used as input formula in the phylostep function, based on the results of the VIF analysis. 
Only the formula that results in the best model is shown, see "Model_selection_crops_elaborated" for all the used formulas. 
```{r define formula}
# 0. Define starting formula
woody_starting_formula <- woody ~ 1 

# 25a. With bio5
woody_formula_25a <- woody ~ Aspect + BulkDensity + ClayPercentage + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio5 + bio14 + bio15 + bio18 + bio19
```

### Change data to match the wanted data format
There are some flaws in the GIS data frame that make it unfit for the phylostep function. First of all
the row names have to match the tip.lables of the tree, and secondly the trait value (0/1) has to be
seen as a numeric value. 
```{r}
#make allmb_name column into the row names of this table
rownames(GIS_woody) <- GIS_woody[,1]
GIS_woody[,1] <- NULL

#make sure the values in the 'woody' column are seen as numeric values
GIS_woody$woody <- as.numeric(GIS_woody$woody)
#check whether it worked
is.numeric(GIS_woody$woody)
```

### Perform model selection
What formula gives us the best result (in terms of the AIC value)?
```{r model selection BM}
#load required package
install.packages("phylolm")
library(phylolm)

#define function components and run function, while changing:
  # - formula: all input formulas 
  # - model: BM, OUrandomRoot, or OUfixedRoot
  # - direction: forward or both

#BM, both: define formula components and run function
phylostep(formula = woody_formula_25a, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "BM", direction = "both", trace = 2)

#OUfixedRoot both: define formula components and run function
phylostep(formula = woody_formula_25a, starting.formula = woody_starting_formula, data = GIS_woody, phy = tree_woody, model = "OUfixedRoot", direction = "both", trace = 2, lower.bound = 0, upper.bound = 1000)
```

What gives us the best result (in terms of the AIC value)?  
- Formula: which input formula 
- Model: BM or OUfixedRoot
- Direction: forward or both 
- Model: what is the best model you get 

### Define best models
```{r}
# 25a. With bio5 for BM
woodiness_BM <- woody ~ 0.83475898 + 0.15198885*BulkDensity -0.48410553*monthCountByTemp10 + 0.08058603*PETWettestQuarter + 0.43981669*OrganicCarbon -0.30052534*PhCaCL -0.19695033*Slope + 0.48659487*bio5 + 0.20864413*bio19

# 25a. With bio5 for OUfixedRoot
woodiness_OU <- woody ~ 0.79869004 -0.32765328*monthCountByTemp10 -0.16620515*PETDriestQuarter + 0.18423198*OrganicCarbon -0.17396725*PhCaCL -0.12842927*Slope + 0.34782999*bio5 -0.08450621*bio14 + 0.07455151*bio19 
```

## 7 Model evaluation

To see whether the best model we constructed, is still in compliance with the assumptions that have to be met when performing a GLM. There is a package that can do this for you, called 'performance'. We'll do a comprehensive model check, with function 'check_model()'
```{r}
#install and load required package 
install.packages("performance")
install.packages("see")
library(performance)
library(see)

#fill in arguments and execute functions
# BM
model <- glm(model = woodiness_BM, data = GIS_woody)
check_model(model)

# OU
model <- glm(model = woodiness_OU, data = GIS_woody)
check_model(model)
```

1. Check for multicolinearity
All correlation is low. 
2. Non-normality of residuals and outliers.
All dots are plotted along the line. 
3. Non-normality of residuals
----- not for glm
4. Homoscedasticity (linear relationship)
----- not for glm
5. Homogeneity of variance (scale-location)
The dots are mirrored... 
6. Check for influential observations
All values are below 0,5; which means that there are no outliers in this dataset. 

## 8 Phylogenetic generalized linear modelling analysis

After this, the phylogenetic generalized linear modeling analysis will be done, which optimizes the model. 
There is a function in the phylolm package that allows us to do this: phyloglm. 

```{r phylogenetic generalized linear modelling}
#fill in the arguments for the phyloglm function and execute it
phyloglm(formula = , data = GIS_woody, phy = tree_woody, method = , btol = , log.alpha.bound = , start.beta = , start.alpha = , boot = , full.matrix = )
```

## 9 Biological interpretation

In the end, the final models and the predictor variables that are included will have to be interpreted biologically. 
With this knowledge, the thesis can be completed. 
