---
title: "Pipeline_myco"
output: html_document
---

```{r setup, i---
title: "Pipeline"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Pipeline of the necessary steps in R and outside of R

## 1 Import data

First the traits table and the original phylogeny (tree) have to be read into R in the right format to be able to use them. 

### The project root and the main tree
```{r define data root and import tree}
#load required package
library(ape)

#import and define data
project_root <- 'C:/Users/Esther/Documents/Master/Stage/'
tree_file <- paste(c(project_root, 'ALLMB.tre'), collapse = '')
tree <- read.tree(tree_file)
```

### Import mycorrhiza data
```{r import mycorrhiza data}
#import and define trait table 
myco_tsv_file <- paste(c(project_root, 'sdmdl-angiosperm-data/data/myco_taxa.tsv'), collapse = '')
myco_tsv <- read.table(myco_tsv_file, sep = '\t', header = TRUE)
```

## 2 Recode dependent variables to binary form

The data of the dependent characters (the plant traits) have to be recoded. This is because a binary 
input is prefered for this analysis due to technical reasons. BUt at the moment, the three plant traits 
all have 3 or 4 states.
For mycorrhiza, there are 4 trait-states: EM, AM, both, neither. This will be turned into: AM (0) and 
the rest together (1). So in this case nothing will be eliminated from the analysis, just (1)(2)(3) have 
to be collapsed together. 
```{r collapse state 1, 2 and 3}
#create new column
myco_tsv["myco_state_int"] <- NA

#add binary values to column based on names 
myco_tsv$myco_state_int[myco_tsv$myco_state == "arbuscular"] <- "0"
myco_tsv$myco_state_int[myco_tsv$myco_state == "ecto"] <- "1"
myco_tsv$myco_state_int[myco_tsv$myco_state == "both"] <- "1"
myco_tsv$myco_state_int[myco_tsv$myco_state == "neither"] <- "1"

#now myco_tsv contains an added column with binary state values
```

## 3 Tidy data

### Match the names 
Make the taxa names the same in the table as in the tree; the difference is that the tsv file has white spaces where the tree file has underscores (_). 
```{r match names}
#load required package
library(stringi)

#replace the whitespaces in the taxa names with '_' 
myco_tsv$allmb_name <- stri_replace_all_charclass(myco_tsv$allmb_name, "\\p{WHITE_SPACE}", "_")
```

## 4 Combine the data

The trait states, the selected subset of taxa names, and the GIS information all have to be combined into one file. 

### Import GIS data
Import the data and make sure the taxa names are the same in this table as in the tsv file and in the tree file,
so with _ between the words instead of whitespaces. 
```{r import GIS data}
#import GIS data 
GIS_file <- paste(c(project_root, 'niche_traits_merged.csv'), collapse = '')
GIS <- read.table(GIS_file, sep = ',', header = TRUE)

#change names of GIS data 
GIS_names <- as.character(unlist(c(GIS[1])))
GIS_names <- gsub('  ', '. ', GIS_names)
GIS_names <- stri_replace_all_charclass(GIS_names, "\\p{WHITE_SPACE}", "_")
#there is a discrepancy with several names that should have a - instead of a _ in some places. 

#change those specific GIS names:
GIS_names[GIS_names == "Acaena_novae_zelandiae"] <- "Acaena_novae-zelandiae"
GIS_names[GIS_names == "Impatiens_noli_tangere"] <- "Impatiens_noli-tangere"  
GIS_names[GIS_names == "Myrrhidendron_donnell_smithii" ] <- "Myrrhidendron_donnell-smithii" 
GIS_names[GIS_names == "Silene_flos_cuculi"] <- "Silene_flos-cuculi"
GIS_names[GIS_names == "Pithecellobium_unguis_cati"] <- "Pithecellobium_unguis-cati"
GIS_names[GIS_names == "Hyacinthoides_non_scripta" ] <- "Hyacinthoides_non-scripta" 
GIS_names[GIS_names == "Alisma_plantago_aquatica"] <- "Alisma_plantago-aquatica" 
GIS_names[GIS_names == "Apera_spica_venti"] <- "Apera_spica-venti"
GIS_names[GIS_names == "Arctostaphylos_uva_ursi"] <- "Arctostaphylos_uva-ursi"
GIS_names[GIS_names == "Blitum_bonus_henricus"] <- "Blitum_bonus-henricus"
GIS_names[GIS_names == "Begonia_meyeri_johannis"] <- "Begonia_meyeri-johannis"

#put the changed names back into the GIS table
GIS$allmb_name <- GIS_names
```

### Tidy data
Make the right selection of taxa out of the GIS variables table and the tree. 
```{r trim GIS table}
#select the taxa that should be kept in the GIS file
keep_GIS_myco <- as.character(unlist(c(myco_tsv[1]))) 
#define the final GIS file for mycorrhiza
GIS_myco <- GIS[GIS$allmb_name %in% keep_GIS_myco,]

#select the names that have to be kept in the tree
keep_tips_myco <- as.character(unlist(c(GIS_myco[1]))) 
#define the final tree for mycorrhiza
tips_to_drop_myco <- setdiff(tree$tip.label, keep_tips_myco)
tree_myco <- drop.tip(tree, tips_to_drop_myco)
```

### Add the binary trait state to the GIS table
```{r add binary state to GIS table}
#make new tables for both mycorrhiza states separately 
GIS_myco_0 <- myco_tsv[myco_tsv$myco_state_int == "0",]
GIS_myco_1 <- myco_tsv[myco_tsv$myco_state_int == "1",]

#add a column to GIS_myco that contains the binary trait values
GIS_myco$mycorrhiza[GIS_myco$allmb_name %in% GIS_myco_0$allmb_name] <- "0"
GIS_myco$mycorrhiza[GIS_myco$allmb_name %in% GIS_myco_1$allmb_name] <- "1"

#change order of the columns for esthetics
GIS_myco <- GIS_myco[,c(1,43, 2:42)]

#there are less plants in this GIS file than in the tree or the tsv file, this makes sense as it could be the case that there are not enough occurrences for each plant in the ALLMB tree, and thus not enough GIS data. 
```

## 5 Correct for colinearity in the predictor variables

There will probably be some colinearity in the GIS-data. This could lead to instability and bias in the model, 
so a correction will have to be done. This can be done using the variance inflation factor (of the usdm package)
or a correlation matrix. See the VIF_analysis_myco file for details of the decision making process.

### VIF-analysis
```{r VIF analysis}
#load required package
library(usdm)

#fill in the arguments of the function and execute it 
vif(GIS_myco[,c(3:43)])

#choose the variables to delete that have a higher VIF value than 10, but make sure it makes sense in terms of biology, results in:
# 1. Solely based on highest VIF value
vif(GIS_myco[,c(3:5, 7, 14, 20, 22:24, 26, 28, 39, 42, 43)])

# 2. With embergerQ
vif(GIS_myco[,c(3:5, 7, 10, 14, 20, 22:24, 26, 28, 39, 42)])

# 3. With climaticMoistureIndex
vif(GIS_myco[,c(3:5, 8, 14, 20, 22:24, 26, 28, 39, 43)])

# 4. With continentality
vif(GIS_myco[,c(3:5, 7, 9, 14, 20, 22:24, 26, 39, 42, 43)])

# 5. With thermicityIndex
vif(GIS_myco[,c(3:5, 20:24, 26, 39, 42, 43)])

# 6. With BIO8 and BIO9 
vif(GIS_myco[,c(3:5, 7, 22:24, 26, 32, 33, 39, 42, 43)])

# 7. With BIO8, BIO9, and EmbergerQ
vif(GIS_myco[,c(3:5, 10, 22:24, 26, 32, 33, 39, 42)])

# 8. With BIO8, BIO9, and climaticMoistureIndex
vif(GIS_myco[,c(3:5, 8, 22:24, 32, 33, 39, 43)])

# 9. With embergerQ and thermicityIndex
vif(GIS_myco[,c(3, 4, 10, 21, 23, 24, 26, 39, 42)])

# 10. With climaticMoistureIndex and thermicityIndex
vif(GIS_myco[,c(3:5, 8, 20, 21, 22:24, 26, 39, 43)])
```

## 6 Model selection: Define formula and select model 

The linear model will have to be made, with the trait-presence-column as the dependent variable, and the 
GIS data frames as the predictor variables. The function phylostep of the phylolm package will be used. 

### Define input formulas
Define the formulas that will be used as input formula in the phylostep function, based on the results of the VIF analysis. 
```{r define formula}
# 0. Define starting formula
myco_starting_formula <- mycorrhiza ~ 1 
  
# 1. Solely based on highest VIF value
myco_formula_1 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio15 + bio18 + bio19

# 2. With embergerQ
myco_formula_2 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + embergerQ + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio15 + bio18 

# 3. With climaticMoistureIndex
myco_formula_3 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio4 + bio15 + bio19

# 4. With continentality
myco_formula_4 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + continentality + minTempWarmest + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio15 + bio18 + bio19

# 5. With thermicityIndex
myco_formula_5 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + PETWettestQuarter + thermicityIndex + OrganicCarbon + PhCaCL + Slope + bio2 + bio14 + bio15 + bio18 + bio19

# 6. With BIO8 and BIO9 
myco_formula_6 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + aridityIndexThornthwaite + OrganicCarbon + PhCaCL + Slope + bio2 + bio8 + bio9 + bio15 + bio18 + bio19

# 7. With BIO8, BIO9, and EmbergerQ
myco_formula_7a <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + embergerQ + OrganicCarbon + PhCaCL + Slope + bio2 + bio8 + bio9 + bio15 + bio17 + bio18
myco_formula_7b <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + embergerQ + OrganicCarbon + PhCaCL + Slope + bio2 + bio8 + bio9 + bio14 + bio15 + bio18

# 8. With BIO8, BIO9, and climaticMoistureIndex
myco_formula_8a <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + OrganicCarbon + PhCaCL + Slope + bio8 + bio9 + bio14 + bio15 + bio19
myco_formula_8b <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + OrganicCarbon + PhCaCL + Slope + bio3 + bio8 + bio9 + bio15 + bio17 + bio19

# 9. With embergerQ and thermicityIndex
myco_formula_9a <- mycorrhiza ~ Aspect + BulkDensity + embergerQ + thermicityIndex + PhCaCL + Slope + bio2 + bio14 + bio15 + bio18
myco_formula_9b <- mycorrhiza ~ Aspect + BulkDensity + embergerQ + PETWettestQuarter + thermicityIndex + OrganicCarbon + PhCaCL + Slope + bio2 + bio15 + bio17 + bio18

# 10. With climaticMoistureIndex and thermicityIndex
myco_formula_10 <- mycorrhiza ~ Aspect + BulkDensity + ClayPercentage + climaticMoistureIndex + PETWettestQuarter + thermicityIndex + OrganicCarbon + PhCaCL + Slope + bio2 + bio15 + bio19
```

### Change data to match the wanted data format
There are some flaws in the GIS data frame that make it unfit for the phylostep function. First of all
the row names have to match the tip.lables of the tree, and secondly the trait value (0/1) has to be
seen as a numeric value. 
```{r}
#make allmb_name column into the row names of this table
rownames(GIS_myco) <- GIS_myco[,1]
GIS_myco[,1] <- NULL

#make sure the values in the 'woody' column are seen as numeric values
GIS_myco$mycorrhiza <- as.numeric(GIS_myco$mycorrhiza)
#check whether it worked
is.numeric(GIS_myco$mycorrhiza)
```

### Perform model selection
```{r model selection BM}
#load required package
library(phylolm)

#define function components and run function, while changing:
  # - formula: all input formulas 
  # - model: BM, OUrandomRoot, or OUfixedRoot
  # - direction: forward or both
phylostep(formula = ??? , starting.formula = myco_starting_formula, data = GIS_myco, phy = tree_myco, model = "???", direction = "???", trace = 2)
```

What gives us the best result (in terms of the AIC value)?  
- Formula: which input formula 
- Model: BM, OUrandomRoot, or OUfixedRoot
- Direction: forward or both 
- Model: what is the best model you get 
For full script and outcomes of the model selection, see "Model_selection_myco" and "Model selection - Combined results". 

## 7 Model evaluation

To see whether the best model we constructed, is still in compliance with the assumptions that have to be met when performing a GLM. There is a package that can do this for you, called 'performance'. We'll do a comprehensive model check, with function 'check_model()'
```{r}
#install and load required package 
install.packages("performance")
library("performance")

model <- glm(model = , data = )
check_model(model)
```

1. Check for multicolinearity
2. Non-normality of residuals and outliers
3. Non-normality of residuals
4. Homoscedasticity (linear relationship)
5. Homogeneity of variance (scale-location)
6. Check for influential observations

## 8 Phylogenetic generalized linear modelling analysis

After this, the phylogenetic generalized linear modeling analysis will be done, which optimizes the model. 
There is a function in the phylolm package that allows us to do this: phyloglm. 

```{r phylogenetic generalized linear modelling}
#fill in the arguments for the phyloglm function and execute it
phyloglm(formula = , data = GIS_myco, phy = tree_myco, method = , btol = , log.alpha.bound = , start.beta = , start.alpha = , boot = , full.matrix = )
```

## 9 Biological interpretation

In the end, the final models and the predictor variables that are included will have to be interpreted biologically. 
With this knowledge, the thesis can be completed. 

