---
title: "Pipeline"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Pipeline of the necessary steps in R and outside of R

## 1 Import data

First the traits table and the original tree have to be read into R in the right format to be able to use them. 

### The project root and the main tree
```{r import tree}
#load required package
library(ape)

#import and define data
project_root <- 'C:/Users/Esther/Documents/Master/Stage/'
tree_file <- paste(c(project_root, 'ALLMB.tre'), collapse = '')
tree <- read.tree(tree_file)
```


### Secondary woodiness
```{r import secondary woodiness}
#load required package
library(ape)

#import and define trait tables
woody_nex_file <- paste(c(project_root, 'sdmdl-angiosperm-data/data/woody.nex'), collapse = '')
woody_nex <- read.nexus(woody_nex_file)

woody_tsv_file <- paste(c(project_root, 'sdmdl-angiosperm-data/data/woody.tsv'), collapse = '')
woody_tsv <- read.table(woody_tsv_file, sep = '\t', header = TRUE)
  
```

### Mycorrhiza
```{r import mycorrhiza}
#load required package
library(ape)

#import and define trait table 
myco_nex_file <- paste(c(project_root, 'sdmdl-angiosperm-data/data/myco.nex'), collapse = '')
myco_nex <- read.nexus(myco_nex_file)

myco_tsv_file <- paste(c(project_root, 'sdmdl-angiosperm-data/data/myco_taxa.tsv'), collapse = '')
myco_tsv <- read.table(myco_tsv_file, sep = '\t', header = TRUE)

myco_idk_file <- paste(c(project_root, 'sdmdl-angiosperm-data/data/myco.tsv'), collapse = '')
myco_idk <- read.table(myco_idk_file, sep = '\t', header = TRUE)

```

### Domestication 
```{r import domestication}
#load required package
library(ape)
#define trait table 
crops_tsv_file <- paste(c(project_root, 'sdmdl-angiosperm-data/data/crops.tsv'), collapse = '')
crops_tsv <- read.table(crops_tsv_file, sep = '\t', header = TRUE)
```

## 2 Match the names of the nexus files and the tsv files

While doing the recoding (step 3), it turned out that the taxa names in the nexus file and in the tsv 
file of woody did not match, this has to be solved and checked for the other traits as well.

### Secondary woodiness

```{r}
#load package and required data
package(ape)
woody_tsv
woody_nex

#make vector of the names inside the two files 
taxa_woody_tsv <- as.character(unlist(c(woody_tsv[1]))) 
taxa_woody_tsv <- sub(" ", "_", taxa_woody_tsv)

taxa_woody_nex <- 

  
#check whether there are differences 
  
#print the different names for both files
  
#correct these names in the tsv file (as nexus file has more specific rules it should stick to)

```


### Mycorrhiza

```{r}
#load package and required data
library(ape)
myco_tsv
tree

#make vector of the names inside the two files 
taxa_myco_tsv <- as.character(unlist(c(myco_tsv[1]))) 
taxa_myco_tsv <- sub(" ", "_", taxa_myco_tsv)

taxa_tree <- as.character(unlist(c(tree$tip.label))) #does not work!! 

#check whether there are differences 


#print the different names for both files that differ 


#correct these names in the tsv file (as the tree file has more specific rules it should stick to)

```


### Domestication
For domestication there is no nexus file. 

## 3 Recode dependent variables to binary form

The data of the dependent characters (the plant traits) have to be recoded. This is because a binary 
input is prefered for this analysis due to technical reasons, while all the data has 3 or 4 states at the
moment. 

### Secondary woodiness
For secondary woodiness, there are 3 trait-states: derived woody (0), non-woody (1), ancestral woody (2). 
This will be turned into: derived woody (0) and non-woody (1). As this is the actual shift between 
non-woody and derived woodiness. So all the plants with ancestral woodiness (3) as trait, have to be 
eliminated from the analysis. 
```{r}
#load required package 
library(ape)

#import the trait tables for woody
woody_tsv

#define new sub-woody, which contains binary trait values (tsv)
woody_tsv_bi <- subset(woody_tsv, (woody_tsv$woody == 0 | woody_tsv$woody ==1))

```

How to do this in the nexus file (not necessary):
```{r}
#load required package 
library(ape)

#import the trait tables for woody
woody_nex
woody_tsv

#define the taxa that should not be included (with value 2)
woody_tsv_2 <- subset(woody_tsv, (woody_tsv$woody == 2)) 
unwanted_tips_woody <- as.character(unlist(c(woody_tsv_2[1]))) 
unwanted_tips_woody <- sub(" ", "_", unwanted_tips_woody) #to make the names the same as in the swo tree. 

#make the subset in the original woody.nex file 
woody_bi_wrong <- drop.tip(woody_nex, unwanted_tips_woody)
#the previous command yielded a wrong dataset (as the name already tells us), it seems like the names still 
#do not match completely, we'll have to fix that. For now this is a different solution that also works: 
unwanted_numbers_woody <- as.numeric(row.names(woody_tsv_2))

#so define new sub-woody, which contains binary trait values (both nex and tsv)
woody_nex_bi <- drop.tip(woody_nex, unwanted_numbers_woody)

#the root edge is not present anymore in swo_bin, but that is okay. 
```


### Mycorrhiza
For mycorrhiza, there are 4 trait-states: EM, AM, both, neither. This will be turned into: AM (0) and 
the rest together (1). So in this case nothing will be eliminated from the analysis, just (1)(2)(3) have 
to be collapsed together. 
```{r}
#load required package
library(ape)

#load required trait table 
myco_tsv

#create new column
myco_tsv["myco_state_int"] <- NA

#add binary values to column based on names 
myco_tsv$myco_state_int[myco_tsv$myco_state == "arbuscular"] <- "0"
myco_tsv$myco_state_int[myco_tsv$myco_state == "ecto"] <- "1"
myco_tsv$myco_state_int[myco_tsv$myco_state == "both"] <- "1"
myco_tsv$myco_state_int[myco_tsv$myco_state == "neither"] <- "1"

#now myco_tsv contains an added column with binary state values
```

### Domestication
The domestication database is not complete yet, some non-domesticated plants have to be added to 
serve as the 0 state. These plants will come from the main tree, based on their distance on the
pylogeny to the plants already included in the domestication traits table. 

```{r}
#load domestication trait table and the tree
crops_tsv

#make a sub-selection of the tree based on earlier selected plants (based on location in phylogeny)


#add this sub-selection of the tree to the domestication trait table 
crops_tsv_complete <- 

#add a 0 to the domesticated column for these new plants

```

Now, for domestication, there are 4 trait-states: non-domesticated (0), progenitor of domesticated (1), 
only domesticated (2) and semi-domesticated (3). 
This will be turned into: non-domesticated (0) and progenitor of domesticated and semi-domesticated 
together (1). This eliminates the species that just exist in their domesticated form, which have 
already been altered by human selection, so they can't tell us anything about pre-adaptations. 
```{r}
#load necessary packages and data 
library(ape)
crops_tsv_complete

#cut out all the plants that have trait state 2 from the tsv file 
crops_tsv_013 <- subset(crops_tsv_complete, (crops_tsv_complete$domestication == 0 | crops_tsv_complete$domestication == 1))

#combine states 1 and 3


#define new sub-dom, which contains binary trait values
crops_tsv_bi <- 

```

## 4 Tidy data

The tree and the traits-tables have to match, but there are plants in the phylogeny (tree) that are 
not in the traits table -> cut these out of the tree. 

### Secondary woodiness
```{r match tips sec. wood.}
#load required packages and data
library(ape)
tree 
woody_tsv_bi

#define the final tree for woodiness
keep_tips_woody_tsv_bi <- as.character(unlist(c(woody_tsv_bi[1]))) 
keep_tips_woody_tsv_bi <- sub(" ", "_", keep_tips_woody_tsv_bi)
tips_to_drop_woody <- setdiff(tree$tip.label, keep_tips_woody_tsv_bi)
woody_tree <- drop.tip(tree, tips_to_drop_woody)

```

### Mycorrhiza 
```{r match tips myc.}
#load required packages and data
library(ape)
tree
myco_tsv_bi 

#define the final tree for myc
keep_tips_myco_tsv_bi <- as.character(unlist(c(myco_tsv_bi[1]))) 
keep_tips_myco_tsv_bi <- sub(" ", "_", keep_tips_myco_tsv_bi)
tips_to_drop_myco <- setdiff(tree$tip.label, keep_tips_myco_tsv_bi)
myco_tree <- drop.tip(tree, tips_to_drop_myco)

```

### Domestication 
```{r match tips dom.}
#load required packages and data
library(ape)
tree 
crops_tsv_bi 

#define the final tree for dom
keep_tips_crops_tsv_bi <- as.character(unlist(c(crops_tsv_bi[1]))) 
keep_tips_crops_tsv_bi <- sub(" ", "_", keep_tips_crops_tsv_bi)
tips_to_drop_crops <- setdiff(tree$tip.label, keep_tips_crops_tsv_bi)
crops_tree <- drop.tip(tree, tips_to_drop_crops)
```

## 5 Correct for colinearity in the predictor variables

There will probably be some colinearity in the GIS-data. This could lead to instability and bias in the model, 
so a correction will have to be done. This can be done using the variance inflation factor (of the usdm package)
or a correlation matrix. 

### VIF-analysis
```{r}
# import GIS data 

#load sdm package
library(usdm)

#execute VIF function 

```

### Biological interpretation
For each of the separate research question, check whether the predictor variables that are still present and
not present anymore, still make sense in terms of biological interpretation. 


## 6 Model selection: Define formula and select model 

The linear model will have to be made, with the trait-presence-column as the dependent variable, and the 
GIS data frames as the predictor variables. The function phylostep of the phylolm package will be used. 

### Secondary woodiness
```{r}
#define formula components

#construct formula

```


### Mycorrhiza
```{r}
#define formula components

#construct formula

```


### Domestication
```{r}
#define formula components

#construct formula

```


## 7 Analysis

After this, the phylogenetic generalized linear modeling analysis will be done, which optimizes the model. 
There is a function in the phylolm package that allows us to do this: phyloglm. 

### Secondary woodiness
```{r}
#fill in the arguments for the phyloglm function and execute it

```


### Mycorrhiza
```{r}
#fill in the arguments for the phyloglm function and execute it

```


### Domestication
```{r}
#fill in the arguments for the phyloglm function and execute it

```


## 8 Biological interpretation

In the end, the final models and the predictor variables that are included will have to be interpreted biologically. 
With this knowledge, the thesis can be completed. 



