---
title: "Pipeline_crops"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Pipeline of the necessary steps in R and outside of R

## 1 Import data

First the traits table and the original tree have to be read into R in the right format to be able to use them. 

### The project root and the main tree
```{r define data root and import tree}
#load required package
library(ape)

#import and define data
project_root <- 'C:/Users/Esther/Documents/Master/Stage/'
tree_file <- paste(c(project_root, 'ALLMB.tre'), collapse = '')
tree <- read.tree(tree_file)
```

### Domestication 
```{r import domestication}
#define trait table 
crops_tsv_file <- paste(c(project_root, 'sdmdl-angiosperm-data/data/crops.tsv'), collapse = '')
crops_tsv <- read.table(crops_tsv_file, sep = '\t', header = TRUE)
```

## 3 Recode dependent variables to binary form

The data of the dependent characters (the plant traits) have to be recoded. This is because a binary 
input is prefered for this analysis due to technical reasons, while all the data has 3 or 4 states at the
moment. 

### Domestication
For domestication, there are 3 trait-states: progenitor of domesticated (1), only domesticated (2) 
and semi-domesticated (3). 
This will be turned into: progenitor of domesticated and semi-domesticated together (1). This 
eliminates the species that just exist in their domesticated form, which have already been altered 
by human selection, so they can't tell us anything about pre-adaptations. 
```{r}
#load necessary packages and data 
library(ape)
crops_tsv

#cut out all the plants that have trait state 2 from the tsv file 
crops_tsv_1_3 <- subset(crops_tsv, (crops_tsv$domestication == 1 | crops_tsv$domestication == 3))

#combine states 1 and 3
crops_tsv_1_3$domestication_binary[crops_tsv_1_3$domestication == "1"] <- "1"
crops_tsv_1_3$domestication_binary[crops_tsv_1_3$domestication == "3"] <- "1"

```

The domestication database is not complete yet, some non-domesticated plants have to be added to 
serve as the 0 state. These plants will come from the main tree, based on their distance on the
pylogeny to the plants already included in the domestication traits table. 

```{r}
#load domestication trait table and the tree
crops_tsv_1_3
tree

#make a sub-selection of the tree based on earlier selected plants (based on location in phylogeny)
#Maybe: show node label for that specific tip label (taxa), then show taxa for that node, pick a taxa from that node (check whether it is in the tsv document on github), add if it's not on github


#add this sub-selection of the tree to the domestication trait table 
crops_tsv_bi <- 

#add a 0 to the domestication_binary and the domestication column for these new plants

  
```


## 4 Tidy data

4a. 
The tree and the traits-tables have to match, but there are plants in the phylogeny (tree) that are 
not in the traits table ->cut these out of the tree. It could be true that the taxa names don't 
match between the tree file and the trait-table, in that case not the same amount of taxa will still
be in the tree as there are in the trait table. If that is the case, this has to be fixed!

### Domestication 
```{r match tips dom.}
#load required packages and data
library(ape)
tree 
crops_tsv_bi 

#define the final tree for dom
keep_tips_crops_tsv_bi <- as.character(unlist(c(crops_tsv_bi[1]))) 
keep_tips_crops_tsv_bi <- stri_replace_all_charclass(keep_tips_crops_tsv_bi, "\\p{WHITE_SPACE}", "_")

tips_to_drop_crops <- setdiff(tree$tip.label, keep_tips_crops_tsv_bi)
crops_tree <- drop.tip(tree, tips_to_drop_crops)
```

## 5 Correct for colinearity in the predictor variables

There will probably be some colinearity in the GIS-data. This could lead to instability and bias in the model, 
so a correction will have to be done. This can be done using the variance inflation factor (of the usdm package)
or a correlation matrix. 

### VIF-analysis
```{r}
# import GIS data 


#load sdm package
library(usdm)

#execute VIF function 

```

### Biological interpretation
For each of the separate research question, check whether the predictor variables that are still present and
not present anymore, still make sense in terms of biological interpretation. 


## 6 Model selection: Define formula and select model 

The linear model will have to be made, with the trait-presence-column as the dependent variable, and the 
GIS data frames as the predictor variables. The function phylostep of the phylolm package will be used. 

### Domestication
```{r}
#define formula components

#construct formula

```


## 7 Analysis

After this, the phylogenetic generalized linear modeling analysis will be done, which optimizes the model. 
There is a function in the phylolm package that allows us to do this: phyloglm. 

### Domestication
```{r}
#fill in the arguments for the phyloglm function and execute it

```


## 8 Biological interpretation

In the end, the final models and the predictor variables that are included will have to be interpreted biologically. 
With this knowledge, the thesis can be completed. 



