---
title: "Pipeline_woody"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Pipeline of the necessary steps in R and outside of R

## 1 Import data

First the traits table and the original tree have to be read into R in the right format to be able to use them. 

### The project root and the main tree
```{r define data root and import tree}
#load required package
library(ape)

#import and define data
project_root <- 'C:/Users/Esther/Documents/Master/Stage/'
tree_file <- paste(c(project_root, 'ALLMB.tre'), collapse = '')
tree <- read.tree(tree_file)

```

### Secondary woodiness
```{r import secondary woodiness}
#import and define trait tables
woody_tsv_file <- paste(c(project_root, 'sdmdl-angiosperm-data/data/woody.tsv'), collapse = '')
woody_tsv <- read.table(woody_tsv_file, sep = '\t', header = TRUE)
  
```

## 3 Recode dependent variables to binary form

The data of the dependent characters (the plant traits) have to be recoded. This is because a binary 
input is prefered for this analysis due to technical reasons, while all the data has 3 or 4 states at the
moment. For secondary woodiness, there are 3 trait-states: derived woody (0), non-woody (1), ancestral woody (2). 
This will be turned into: derived woody (0) and non-woody (1). As this is the actual shift between 
non-woody and derived woodiness. So all the plants with ancestral woodiness (3) as trait, have to be 
eliminated from the analysis. 

```{r binary woodiness tsv}
#define new sub-woody, which contains binary trait values (tsv)
woody_tsv_bi <- subset(woody_tsv, (woody_tsv$woody == 0 | woody_tsv$woody ==1))
```

## 4 Tidy data

4a. 
The tree and the traits-tables have to match, but there are plants in the phylogeny (tree) that are 
not in the traits table -> cut these out of the tree. It could be true that the taxa names don't 
match between the tree file and the trait-table, in that case not the same amount of taxa will still
be in the tree as there are in the trait table. If that is the case, this has to be fixed!

```{r match tips sec. wood.}
#load required packages 
library(stringi)

#define the taxa names for the final tree for woodiness
keep_tips_woody_tsv_bi <- as.character(unlist(c(woody_tsv_bi[1]))) 
keep_tips_woody_tsv_bi <- stri_replace_all_charclass(keep_tips_woody_tsv_bi, "\\p{WHITE_SPACE}", "_")

tips_to_drop_woody <- setdiff(tree$tip.label, keep_tips_woody_tsv_bi)
woody_tree <- drop.tip(tree, tips_to_drop_woody)

```

4b. 
Make the taxa names the same in the table as in the tree. 
```{r}
#replace the spaces in the taxa names for _ 
woody_tsv_bi$allmb_name <- stri_replace_all_charclass(woody_tsv_bi$allmb_name, "\\p{WHITE_SPACE}", "_")

```


## 5 Combine the data

The trait states, the selected subset of taxa names, and the GIS information all have to be combined into one file. 

### Import data
Import the data and make sure the taxa names are the same in this table as in the tsv file and in the tree file,
so with _ instead of spaces 
```{r}
#import GIS data
GIS_file <- paste(c(project_root, 'niche_traits_merged.csv'), collapse = '')
GIS <- read.table(GIS_file, sep = ',', header = TRUE)

#change names of GIS data 
GIS_names <- as.character(unlist(c(GIS[1])))
GIS_names <- stri_replace_all_charclass(GIS_names, "\\p{WHITE_SPACE}", "_")
GIS$allmb_name <- GIS_names
```

### Tidy data
Make the right selection of taxa out of the GIS variables table. 
```{r}
#define the taxa that should be selected from the GIS file
keep_GIS_woody <- as.character(unlist(c(woody_tsv_bi[1])))

#make the selection in the GIS file
GIS_woody <- GIS[GIS$allmb_name %in% keep_GIS_woody,]

```

### Add the binary trait state to the GIS table
```{r}
#select the taxa of woody_tsv_bi that are in GIS_woody
woody_tsv_bi_GIS <- woody_tsv_bi[woody_tsv_bi$allmb_name %in% GIS_woody$allmb_name,]

#make sure that both files are displaying the names alphabetically, so that the trait values are not added
#randomly but are really connected to the names. 
GIS_woody <- GIS_woody[order(GIS_woody$allmb_name),]
woody_tsv_bi_GIS <- woody_tsv_bi_GIS[order(woody_tsv_bi_GIS$allmb_name),]

#add a column to GIS_woody that contains the binary trait values
GIS_woody$woody <-  

```

## 6 Correct for colinearity in the predictor variables

There will probably be some colinearity in the GIS-data. This could lead to instability and bias in the model, 
so a correction will have to be done. This can be done using the variance inflation factor (of the usdm package)
or a correlation matrix. 

### VIF-analysis
```{r}
# import GIS data 

#load sdm package
library(usdm)

#execute VIF function 

```

### Biological interpretation
For each of the separate research question, check whether the predictor variables that are still present and
not present anymore, still make sense in terms of biological interpretation. 

## 7 Model selection: Define formula and select model 

The linear model will have to be made, with the trait-presence-column as the dependent variable, and the 
GIS data frames as the predictor variables. The function phylostep of the phylolm package will be used. 

```{r}
#define formula components

#construct formula

```

## 8 Analysis

After this, the phylogenetic generalized linear modeling analysis will be done, which optimizes the model. 
There is a function in the phylolm package that allows us to do this: phyloglm. 

```{r}
#fill in the arguments for the phyloglm function and execute it

```

## 9 Biological interpretation

In the end, the final models and the predictor variables that are included will have to be interpreted biologically. 
With this knowledge, the thesis can be completed. 



