---
title: "Pipeline_woody"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Pipeline of the necessary steps in R and outside of R

## 1 Import data

First the traits table and the original phylogeny (tree) have to be read into R in the right format to be able to use them. 

### The project root and the main tree
```{r define data root and import tree}
#load required package
library(ape)

#import and define data
project_root <- 'C:/Users/Esther/Documents/Master/Stage/'
tree_file <- paste(c(project_root, 'ALLMB.tre'), collapse = '')
tree <- read.tree(tree_file)
```

### Import derived woodiness data
```{r import derived woodiness data}
#import and define trait tables
woody_tsv_file <- paste(c(project_root, 'sdmdl-angiosperm-data/data/woody.tsv'), collapse = '')
woody_tsv <- read.table(woody_tsv_file, sep = '\t', header = TRUE)
```

## 2 Recode dependent variables to binary form

The data of the dependent characters (the plant traits) have to be recoded. This is because a binary 
input is prefered for this analysis due to technical reasons. BUt at the moment, the three plant traits 
all have 3 or 4 states For secondary woodiness, there are 3 trait-states: derived woody (0), non-woody (1), 
ancestral woody (2). 
This will be turned into: derived woody (0) and non-woody (1). Because we want to know more about this 
precise shift between non-woody and derived woodiness. So all the plants with ancestral woodiness (3) as 
trait, have to be eliminated from the analysis. 

```{r cut out state 2}
#define new sub-woody, which contains binary trait values (tsv)
woody_tsv_bi <- subset(woody_tsv, (woody_tsv$woody == 0 | woody_tsv$woody ==1))
```

## 3 Tidy data

### Match the names
Make the taxa names the same in the table as in the tree; the difference is that the tsv file has white spaces where the tree file has underscores (_). 
```{r match names}
#load required package 
library(stringi)

#replace the spaces in the taxa names with '_' 
woody_tsv_bi$allmb_name <- stri_replace_all_charclass(woody_tsv_bi$allmb_name, "\\p{WHITE_SPACE}", "_")
```

## 4 Combine the data

The trait states, the selected subset of taxa names, and the GIS information all have to be combined into one file. 

### Import GIS data
Import the data and make sure the taxa names are the same in this table as in the tsv file and in the tree file,
so with _ between the words instead of whitespaces. 
```{r import GIS data}
#import GIS data
GIS_file <- paste(c(project_root, 'niche_traits_merged.csv'), collapse = '')
GIS <- read.table(GIS_file, sep = ',', header = TRUE)

#change names of GIS data 
GIS_names <- as.character(unlist(c(GIS[1])))
GIS_names <- gsub('  ', '. ', GIS_names)
GIS_names <- stri_replace_all_charclass(GIS_names, "\\p{WHITE_SPACE}", "_")
#there is a discrepancy with several names that should have a - instead of a _ in some places. 

#change those specific GIS names:
GIS_names[GIS_names == "Acaena_novae_zelandiae"] <- "Acaena_novae-zelandiae"
GIS_names[GIS_names == "Impatiens_noli_tangere"] <- "Impatiens_noli-tangere"  
GIS_names[GIS_names == "Myrrhidendron_donnell_smithii" ] <- "Myrrhidendron_donnell-smithii" 
GIS_names[GIS_names == "Silene_flos_cuculi"] <- "Silene_flos-cuculi"
GIS_names[GIS_names == "Pithecellobium_unguis_cati"] <- "Pithecellobium_unguis-cati"
GIS_names[GIS_names == "Hyacinthoides_non_scripta" ] <- "Hyacinthoides_non-scripta" 
GIS_names[GIS_names == "Alisma_plantago_aquatica"] <- "Alisma_plantago-aquatica" 
GIS_names[GIS_names == "Apera_spica_venti"] <- "Apera_spica-venti"
GIS_names[GIS_names == "Arctostaphylos_uva_ursi"] <- "Arctostaphylos_uva-ursi"
GIS_names[GIS_names == "Blitum_bonus_henricus"] <- "Blitum_bonus-henricus"
GIS_names[GIS_names == "Begonia_meyeri_johannis"] <- "Begonia_meyeri-johannis"

#put the changed names back into the GIS table
GIS$allmb_name <- GIS_names
```

### Tidy data
Make the right selection of taxa out of the GIS variables table and the tree. 
```{r trim GIS table}
#select the taxa that should be kept in the GIS file
keep_GIS_woody <- as.character(unlist(c(woody_tsv_bi[1])))
#define the final GIS file for woodiness
GIS_woody <- GIS[GIS$allmb_name %in% keep_GIS_woody,]

#select the names that have to be kept in the tree based on the GIS file
keep_tips_woody <- as.character(unlist(c(GIS_woody[1]))) 
#define the final tree for mycorrhiza
tips_to_drop_woody <- setdiff(tree$tip.label, keep_tips_woody)
tree_woody <- drop.tip(tree, tips_to_drop_woody)
```

### Add the binary trait state to the GIS table
```{r add binary state to GIS table}
#make new tables for both woodiness states separately 
GIS_woody_0 <- woody_tsv_bi[woody_tsv_bi$woody == "0",]
GIS_woody_1 <- woody_tsv_bi[woody_tsv_bi$woody == "1",]

#add a column to GIS_woody that contains the binary trait values
GIS_woody$woody[GIS_woody$allmb_name %in% GIS_woody_0$allmb_name] <- "0"  
GIS_woody$woody[GIS_woody$allmb_name %in% GIS_woody_1$allmb_name] <- "1"  

#change order of the columns for esthetics
GIS_woody <- GIS_woody[,c(1,43, 2:42)]

#there are less plants in this GIS file than in the tree or the tsv file, this makes sense as it could be the case that there are not enough occurrences for each plant in the ALLMB tree, and thus not enough GIS data. 
```

## 5 Correct for colinearity in the predictor variables

There will probably be some colinearity in the GIS-data. This could lead to instability and bias in the model, 
so a correction will have to be done. This can be done using the variance inflation factor (of the usdm package)
or a correlation matrix. 
Only the combination that results in the best model is shown, see VIF_analysis_myco" for all the used combinations.

### VIF-analysis
```{r VIF analysis}
#load required package
library(usdm)

#fill in the arguments of the function and execute it 
vif(GIS_woody[,c(3:43)])

#choose variables to delete that have a higher VIF value than 10, but make sure it makes sense in terms of biology, resulting in:
# 24a. With bio5
vif(GIS_woody[,c(3:5, 15, 17, 20, 22:24, 26, 29, 38, 39, 42, 43)])
```

## 6 Model selection: Define formula and select model 

The linear model will have to be made, with the trait-presence-column as the dependent variable, and the 
GIS data frames as the predictor variables. The function phylostep of the phylolm package will be used. 

### Define input formulas
Define the formulas that will be used as input formula in the phylostep function, based on the results of the VIF analysis. 
Only the formula that results in the best model is shown, see "Model_selection_crops_elaborated" for all the used formulas. 
```{r define formula}
# 24a. With bio5
woody_formula_24a <- woody ~ Aspect + BulkDensity + ClayPercentage + monthCountByTemp10 + PETDriestQuarter + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio2 + bio5 + bio14 + bio15 + bio18 + bio19
```

### Change data to match the wanted data format
There are some flaws in the GIS data frame that make it unfit for the phylostep function. First of all
the row names have to match the tip.lables of the tree, and secondly the trait value (0/1) has to be
seen as a numeric value. 
```{r}
#make allmb_name column into the row names of this table
rownames(GIS_woody) <- GIS_woody[,1]
GIS_woody[,1] <- NULL

#make sure the values in the 'woody' column are seen as numeric values
GIS_woody$woody <- as.numeric(GIS_woody$woody)
#check whether it worked
is.numeric(GIS_woody$woody)
```

### Perform model selection
What formula gives us the best result (in terms of the AIC value)?
```{r}
#source function
library(phylolm)
source('phyloglmstep.R')

#fill in arguments and execute function
phyloglmstep(woody_formula_24a, starting.formula = NULL, data=GIS_woody, phy=tree_woody, 
                         method="logistic_MPLE",
                         direction = "forward", trace = 0,
                         btol = 100, log.alpha.bound = 8, start.beta=NULL, 
                         start.alpha=NULL, boot = 0, full.matrix = TRUE,
                         k=2)
```

## 7 Phylogenetic generalized linear modelling analysis

After this, the phylogenetic generalized linear modeling analysis will be done, with which we can optimize the model by looking at the significance of the different values.  
There is a function in the phylolm package that allows us to do this: phyloglm. 

```{r phylogenetic generalized linear modelling}
#construct the best model
woody_best_model <- woody ~ 1 + monthCountByTemp10 + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio5 + bio14

#fill in the arguments for the phyloglm function and execute it
model_woody <- phyloglm(formula = woody_best_model, data = GIS_woody, phy = tree_woody, method = "logistic_MPLE", btol = 100, log.alpha.bound = 8)
summary(model_woody) 
```

### Second round
The bio14 variable was not significant, so we will leave that out and do the phyloglm optimization again. 

```{r}
#construct best model
woody_best_model2<- woody ~ 1 + monthCountByTemp10 + PETWettestQuarter + OrganicCarbon + PhCaCL + Slope + bio5

#fill in arguments and execute function
model_woody2 <- phyloglm(formula = woody_best_model2, data = GIS_woody, phy = tree_woody, method = "logistic_MPLE", btol = 100, log.alpha.bound = 8)

summary(model_woody2)
```

### Define final model
```{r}
final_model_woody <- woody ~ 1.78041 - 2.20071*monthCountByTemp10 + 0.28208*PETWettestQuarter + 1.28682*OrganicCarbon - 1.07275*PhCaCL - 0.98976*Slope + 1.48267*bio5
```

## 8 Model evaluation

To see whether the best model we constructed, is still in compliance with the assumptions that have to be met when performing a GLM. There is a package that can do this for you, called 'performance'. We'll do a comprehensive model check, with function 'check_model()'
```{r}
#install and load required package 
library(performance)
library(see)
library(gridExtra)

#check model
check_model(final_model_woody)
```

1. Check for multicolinearity
2. Non-normality of residuals and outliers
3. Homogeneity of variance (scale-location)
4. Check for influential observations

## 9 Plot results

```{r boxplots}
#load required package
library(dplyr)

#define two subsets
GIS_filtered_non_woody <- GIS_woody %>%
    select(woody, monthCountByTemp10, PETWettestQuarter, OrganicCarbon,
    PhCaCL, Slope, bio5) %>%
    subset(woody == 0)
GIS_filtered_woody <- GIS_woody %>%
    select(woody, monthCountByTemp10, PETWettestQuarter, OrganicCarbon,
    PhCaCL, Slope, bio5) %>%
    subset(woody == 1)

#make boxplots
boxplot(GIS_filtered_woody$monthCountByTemp10,GIS_filtered_non_woody$monthCountByTemp10,
    main = "Number of months with mean temp greater than 10 degrees C",
    at = c(1,2),
    names = c("derived woody", "herbaceous"),
    las = 2,
    col = c("orange","red"),
    border = "brown",
    horizontal = TRUE,
    notch = TRUE
)
boxplot(GIS_filtered_woody$PETWettestQuarter,GIS_filtered_non_woody$PETWettestQuarter,
    main = "Mean monthly potential evapotranspiration of wettest quarter",
    at = c(1,2),
    names = c("derived woody", "herbaceous"),
    las = 2,
    col = c("orange","red"),
    border = "brown",
    horizontal = TRUE,
    notch = TRUE
)
boxplot(GIS_filtered_woody$OrganicCarbon,GIS_filtered_non_woody$OrganicCarbon,
    main = "Soil organic carbon content",
    at = c(1,2),
    names = c("derived woody", "herbaceous"),
    las = 2,
    col = c("orange","red"),
    border = "brown",
    horizontal = TRUE,
    notch = TRUE
)
boxplot(GIS_filtered_woody$PhCaCL,GIS_filtered_non_woody$PhCaCL,
    main = "Soil pH",
    at = c(1,2),
    names = c("derived woody", "herbaceous"),
    las = 2,
    col = c("orange","red"),
    border = "brown",
    horizontal = TRUE,
    notch = TRUE
)
boxplot(GIS_filtered_woody$Slope,GIS_filtered_non_woody$Slope,
    main = "Slope",
    at = c(1,2),
    names = c("derived woody", "herbaceous"),
    las = 2,
    col = c("orange","red"),
    border = "brown",
    horizontal = TRUE,
    notch = TRUE
)
boxplot(GIS_filtered_woody$bio5,GIS_filtered_non_woody$bio5,
    main = "Max temperature of warmest month",
    at = c(1,2),
    names = c("derived woody", "herbaceous"),
    las = 2,
    col = c("orange","red"),
    border = "brown",
    horizontal = TRUE,
    notch = TRUE
)
```

## 10 Biological interpretation

In the end, the final models and the predictor variables that are included will have to be interpreted biologically. 
With this knowledge, the thesis can be completed. 
