---
title: "Pipeline"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Pipeline of the necessary steps in R and outside of R

## 1 Import data

First the traits table and the original tree have to be read into R in the right format to be able to use them. 

### The project root and the main tree
```{r define data root and import tree}
#load required package
library(ape)

#import and define data
project_root <- 'C:/Users/Esther/Documents/Master/Stage/'
tree_file <- paste(c(project_root, 'ALLMB.tre'), collapse = '')
tree <- read.tree(tree_file)
```


### Secondary woodiness
```{r import secondary woodiness}
#load required package
library(ape)

#import and define trait tables
woody_nex_file <- paste(c(project_root, 'sdmdl-angiosperm-data/data/woody.nex'), collapse = '')
woody_nex <- read.nexus(woody_nex_file)

woody_tsv_file <- paste(c(project_root, 'sdmdl-angiosperm-data/data/woody.tsv'), collapse = '')
woody_tsv <- read.table(woody_tsv_file, sep = '\t', header = TRUE)
  
```

### Mycorrhiza
```{r import mycorrhiza}
#load required package
library(ape)

#import and define trait table 
myco_nex_file <- paste(c(project_root, 'sdmdl-angiosperm-data/data/myco.nex'), collapse = '')
myco_nex <- read.nexus(myco_nex_file)

myco_tsv_file <- paste(c(project_root, 'sdmdl-angiosperm-data/data/myco_taxa.tsv'), collapse = '')
myco_tsv <- read.table(myco_tsv_file, sep = '\t', header = TRUE)

myco_idk_file <- paste(c(project_root, 'sdmdl-angiosperm-data/data/myco.tsv'), collapse = '')
myco_idk <- read.table(myco_idk_file, sep = '\t', header = TRUE)

```

### Domestication 
```{r import domestication}
#load required package
library(ape)
#define trait table 
crops_tsv_file <- paste(c(project_root, 'sdmdl-angiosperm-data/data/crops.tsv'), collapse = '')
crops_tsv <- read.table(crops_tsv_file, sep = '\t', header = TRUE)
```

## 3 Recode dependent variables to binary form

The data of the dependent characters (the plant traits) have to be recoded. This is because a binary 
input is prefered for this analysis due to technical reasons, while all the data has 3 or 4 states at the
moment. 

### Secondary woodiness
For secondary woodiness, there are 3 trait-states: derived woody (0), non-woody (1), ancestral woody (2). 
This will be turned into: derived woody (0) and non-woody (1). As this is the actual shift between 
non-woody and derived woodiness. So all the plants with ancestral woodiness (3) as trait, have to be 
eliminated from the analysis. 
```{r binary woodiness tsv}
#load required package 
library(ape)

#import the trait tables for woody
woody_tsv

#define new sub-woody, which contains binary trait values (tsv)
woody_tsv_bi <- subset(woody_tsv, (woody_tsv$woody == 0 | woody_tsv$woody ==1))

```

How to do this in the nexus file (not necessary):
```{r binary woodiness nex}
#load required package 
library(ape)

#import the trait tables for woody
woody_nex
woody_tsv

#define the taxathat should not be included (with value 2)
woody_tsv_2 <- subset(woody_tsv, (woody_tsv$woody == 2)) 
unwanted_tips_woody <- as.character(unlist(c(woody_tsv_2[1]))) 
unwanted_tips_woody <- sub(" ", "_", unwanted_tips_woody) #to make the names the same as in the swo tree. 

#make the subset in the original woody.nex file 
woody_bi_wrong <- drop.tip(woody_nex, unwanted_tips_woody)
#the previous command yielded a wrong dataset (as the name already tells us), it seems like the names still 
#do not match completely, we'll have to fix that. For now this is a different solution that also works: 
unwanted_numbers_woody <- as.numeric(row.names(woody_tsv_2))

#so define new sub-woody, which contains binary trait values (both nex and tsv)
woody_nex_bi <- drop.tip(woody_nex, unwanted_numbers_woody)

#the root edge is not present anymore in swo_bin, but that is okay. 
```


### Mycorrhiza
For mycorrhiza, there are 4 trait-states: EM, AM, both, neither. This will be turned into: AM (0) and 
the rest together (1). So in this case nothing will be eliminated from the analysis, just (1)(2)(3) have 
to be collapsed together. 
```{r binary mycorrhiza tsv}
#load required package
library(ape)

#load required trait table 
myco_tsv

#create new column
myco_tsv["myco_state_int"] <- NA

#add binary values to column based on names 
myco_tsv$myco_state_int[myco_tsv$myco_state == "arbuscular"] <- "0"
myco_tsv$myco_state_int[myco_tsv$myco_state == "ecto"] <- "1"
myco_tsv$myco_state_int[myco_tsv$myco_state == "both"] <- "1"
myco_tsv$myco_state_int[myco_tsv$myco_state == "neither"] <- "1"

#now myco_tsv_bi contains an added column with binary state values
```

How to do this in the nexus file (not necessary):
```{r binary mycorrhiza nex}

```


### Domestication
For domestication, there are 3 trait-states: progenitor of domesticated (1), only domesticated (2) 
and semi-domesticated (3). 
This will be turned into: progenitor of domesticated and semi-domesticated together (1). This 
eliminates the species that just exist in their domesticated form, which have already been altered 
by human selection, so they can't tell us anything about pre-adaptations. 
```{r}
#load necessary packages and data 
library(ape)
crops_tsv

#cut out all the plants that have trait state 2 from the tsv file 
crops_tsv_1_3 <- subset(crops_tsv, (crops_tsv$domestication == 1 | crops_tsv$domestication == 3))

#combine states 1 and 3
crops_tsv_1_3$domestication_binary[crops_tsv_1_3$domestication == "1"] <- "1"
crops_tsv_1_3$domestication_binary[crops_tsv_1_3$domestication == "3"] <- "1"

```

The domestication database is not complete yet, some non-domesticated plants have to be added to 
serve as the 0 state. These plants will come from the main tree, based on their distance on the
pylogeny to the plants already included in the domestication traits table. 

```{r}
#load domestication trait table and the tree
crops_tsv_1_3
tree

#make a sub-selection of the tree based on earlier selected plants (based on location in phylogeny)
#Maybe: show node label for that specific tip label (taxa), then show taxa for that node, pick a taxa from that node (check whether it is in the tsv document on github), add if it's not on github


#add this sub-selection of the tree to the domestication trait table 
crops_tsv_bi <- 

#add a 0 to the domestication_binary and the domestication column for these new plants

  
```


## 4 Tidy data

4a. 
The tree and the traits-tables have to match, but there are plants in the phylogeny (tree) that are 
not in the traits table ->cut these out of the tree. It could be true that the taxa names don't 
match between the tree file and the trait-table, in that case not the same amount of taxa will still
be in the tree as there are in the trait table. If that is the case, this has to be fixed!

### Secondary woodiness
```{r match tips sec. wood.}
#load required packages and data
library(ape)
library(stringi)
tree 
woody_tsv_bi

#define the final tree for woodiness
keep_tips_woody_tsv_bi <- as.character(unlist(c(woody_tsv_bi[1]))) 
keep_tips_woody_tsv_bi <- stri_replace_all_charclass(keep_tips_woody_tsv_bi, "\\p{WHITE_SPACE}", "_")

tips_to_drop_woody <- setdiff(tree$tip.label, keep_tips_woody_tsv_bi)
woody_tree <- drop.tip(tree, tips_to_drop_woody)

#it works! the woody_tree now also contains only 749 tips/taxa.  
```

### Mycorrhiza 
```{r match tips myc.}
#load required packages and data
library(ape)
library(stringi)
tree
myco_tsv 

#define the final tree for myc
keep_tips_myco_tsv_bi <- as.character(unlist(c(myco_tsv[1]))) 
keep_tips_myco_tsv_bi <- stri_replace_all_charclass(keep_tips_myco_tsv_bi, "\\p{WHITE_SPACE}", "_")

tips_to_drop_myco <- setdiff(tree$tip.label, keep_tips_myco_tsv_bi)
myco_tree <- drop.tip(tree, tips_to_drop_myco)

```

### Domestication 
```{r match tips dom.}
#load required packages and data
library(ape)
tree 
crops_tsv_bi 

#define the final tree for dom
keep_tips_crops_tsv_bi <- as.character(unlist(c(crops_tsv_bi[1]))) 
keep_tips_crops_tsv_bi <- stri_replace_all_charclass(keep_tips_crops_tsv_bi, "\\p{WHITE_SPACE}", "_")

tips_to_drop_crops <- setdiff(tree$tip.label, keep_tips_crops_tsv_bi)
crops_tree <- drop.tip(tree, tips_to_drop_crops)
```

4b. Make sure the names match 
### Example for checking the names if they don't match: Secondary woodiness
```{r matching woody}
#load package and required data
package(ape)
keep_tips_woody_tsv_bi
woody_tree

#make vector of the names that do not match
keep_tips_woody_tsv_bi
nomatch_names_woody <- setdiff(keep_tips_woody_tsv_bi, woody_tree$tip.label)

#print the names
print(nomatch_names_woody) 

#it turns out that not all the white spaces were substituted by _ with the sub(" ", "_") command
#so tried a different command: 
keep_tips_woody_tsv_bi <- stri_replace_all_charclass(keep_tips_woody_tsv_bi, "\\p{WHITE_SPACE}", "_")

#this worked! now implemented in the code above 
```


## 5 Combine the data

The trait states, the selected subset of taxa names, and the GIS information all have to be combined into one file. 

### Import data
Import the data and make sure the taxa names are the same in this table as in the tsv file and in the tree file,
so with _ instead of spaces 
```{r}
#import GIS data
GIS_file <- paste(c(project_root, 'niche_traits_merged.csv'), collapse = '')
GIS <- read.table(GIS_file, sep = ',', header = TRUE)

#change names of GIS data 
GIS_names <- as.character(unlist(c(GIS[1])))
GIS_names <- stri_replace_all_charclass(GIS_names, "\\p{WHITE_SPACE}", "_")
GIS$allmb_name <- GIS_names
```

### Tidy data
Make the right selection of taxa out of the GIS variables table. 
#### Woodiness
```{r}
#define the taxa that should be selected from the GIS file
keep_GIS_woody <- as.character(unlist(c(woody_tsv_bi[1]))) 

#make the selection in the GIS file
GIS_woody <- GIS[GIS$allmb_name %in% keep_GIS_woody,]
```
#### Mycorrhiza
```{r}
#define the taxa that should be selected from the GIS file
keep_GIS_myco <- as.character(unlist(c(myco_tsv[1]))) 

#make the selection in the GIS file
GIS_myco <- GIS[GIS$allmb_name %in% keep_GIS_myco,]

```
#### Crops
```{r}
#define the taxa that should be selected from the GIS file
keep_GIS_crops <- as.character(unlist(c(crops_tsv_bi[1]))) 

#make the selection in the GIS file
GIS_crops <- GIS[GIS$allmb_name %in% keep_GIS_crops,]
```

### Add the binary trait state to the GIS table

#### Woodiness
```{r}
#select the taxa of woody_tsv_bi that are in GIS_woody
woody_tsv_bi_GIS <- woody_tsv_bi[woody_tsv_bi$allmb_name %in% GIS_woody$allmb_name,]

#add a column to GIS_woody that contains the binary trait values
GIS_woody$woody <- 

```

#### Mycorrhiza
```{r}
#select the taxa of woody_tsv_bi that are in GIS_woody
myco_tsv_GIS <- myco_tsv[myco_tsv$allmb_name %in% GIS_myco$allmb_name,]

#add a column to GIS_myco that contains the binary trait values
GIS_myco$mycorrhiza <- 

```

#### Domestication
```{r}
#select the taxa of woody_tsv_bi that are in GIS_woody
crops_tsv_bi_GIS <- crops_tsv_bi[crops_tsv_bi$allmb_name %in% GIS_crops$allmb_name,]

#add a column to GIS_woody that contains the binary trait values
GIS_crops$domestication <- 

```


## 6 Correct for colinearity in the predictor variables

There will probably be some colinearity in the GIS-data. This could lead to instability and bias in the model, 
so a correction will have to be done. This can be done using the variance inflation factor (of the usdm package)
or a correlation matrix. 

### VIF-analysis
```{r}
# import GIS data 

#load sdm package
library(usdm)

#execute VIF function 

```

### Biological interpretation
For each of the separate research question, check whether the predictor variables that are still present and
not present anymore, still make sense in terms of biological interpretation. 


## 7 Model selection: Define formula and select model 

The linear model will have to be made, with the trait-presence-column as the dependent variable, and the 
GIS data frames as the predictor variables. The function phylostep of the phylolm package will be used. 

### Secondary woodiness
```{r}
#define formula components

#construct formula

```


### Mycorrhiza
```{r}
#define formula components

#construct formula

```


### Domestication
```{r}
#define formula components

#construct formula

```


## 8 Analysis

After this, the phylogenetic generalized linear modeling analysis will be done, which optimizes the model. 
There is a function in the phylolm package that allows us to do this: phyloglm. 

### Secondary woodiness
```{r}
#fill in the arguments for the phyloglm function and execute it

```


### Mycorrhiza
```{r}
#fill in the arguments for the phyloglm function and execute it

```


### Domestication
```{r}
#fill in the arguments for the phyloglm function and execute it

```


## 9 Biological interpretation

In the end, the final models and the predictor variables that are included will have to be interpreted biologically. 
With this knowledge, the thesis can be completed. 



