---
title: "Pipeline_myco"
output: html_document
---

```{r setup, i---
title: "Pipeline"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Pipeline of the necessary steps in R and outside of R

## 1 Import data

First the traits table and the original tree have to be read into R in the right format to be able to use them. 

### The project root and the main tree
```{r define data root and import tree}
#load required package
library(ape)

#import and define data
project_root <- 'C:/Users/Esther/Documents/Master/Stage/'
tree_file <- paste(c(project_root, 'ALLMB.tre'), collapse = '')
tree <- read.tree(tree_file)
```

### Mycorrhiza
```{r import mycorrhiza}
#import and define trait table 
myco_tsv_file <- paste(c(project_root, 'sdmdl-angiosperm-data/data/myco_taxa.tsv'), collapse = '')
myco_tsv <- read.table(myco_tsv_file, sep = '\t', header = TRUE)
```

## 3 Recode dependent variables to binary form

The data of the dependent characters (the plant traits) have to be recoded. This is because a binary 
input is prefered for this analysis due to technical reasons, while all the data has 3 or 4 states at the
moment. 

### Mycorrhiza
For mycorrhiza, there are 4 trait-states: EM, AM, both, neither. This will be turned into: AM (0) and 
the rest together (1). So in this case nothing will be eliminated from the analysis, just (1)(2)(3) have 
to be collapsed together. 
```{r binary mycorrhiza tsv}
#create new column
myco_tsv["myco_state_int"] <- NA

#add binary values to column based on names 
myco_tsv$myco_state_int[myco_tsv$myco_state == "arbuscular"] <- "0"
myco_tsv$myco_state_int[myco_tsv$myco_state == "ecto"] <- "1"
myco_tsv$myco_state_int[myco_tsv$myco_state == "both"] <- "1"
myco_tsv$myco_state_int[myco_tsv$myco_state == "neither"] <- "1"

#now myco_tsv_bi contains an added column with binary state values
```

## 4 Tidy data

4a. 
The tree and the traits-tables have to match, but there are plants in the phylogeny (tree) that are 
not in the traits table ->cut these out of the tree. It could be true that the taxa names don't 
match between the tree file and the trait-table, in that case not the same amount of taxa will still
be in the tree as there are in the trait table. If that is the case, this has to be fixed!

### Mycorrhiza 
```{r match tips myc.}
#load required packages and data
library(stringi)

#define the final tree for myc
keep_tips_myco_tsv_bi <- as.character(unlist(c(myco_tsv[1]))) 
keep_tips_myco_tsv_bi <- stri_replace_all_charclass(keep_tips_myco_tsv_bi, "\\p{WHITE_SPACE}", "_")

tips_to_drop_myco <- setdiff(tree$tip.label, keep_tips_myco_tsv_bi)
myco_tree <- drop.tip(tree, tips_to_drop_myco)

```

4b. 
Make the taxa names the same in the table as in the tree. 
```{r}
#replace the spaces in the taxa names for _ 
myco_tsv$allmb_name <- stri_replace_all_charclass(myco_tsv$allmb_name, "\\p{WHITE_SPACE}", "_")

```

## 5 Correct for colinearity in the predictor variables

There will probably be some colinearity in the GIS-data. This could lead to instability and bias in the model, 
so a correction will have to be done. This can be done using the variance inflation factor (of the usdm package)
or a correlation matrix. 

### VIF-analysis
```{r}
# import GIS data 


#load sdm package
library(usdm)

#execute VIF function 

```

### Biological interpretation
For each of the separate research question, check whether the predictor variables that are still present and
not present anymore, still make sense in terms of biological interpretation. 


## 6 Model selection: Define formula and select model 

The linear model will have to be made, with the trait-presence-column as the dependent variable, and the 
GIS data frames as the predictor variables. The function phylostep of the phylolm package will be used. 

### Mycorrhiza
```{r}
#define formula components

#construct formula

```

## 7 Analysis

After this, the phylogenetic generalized linear modeling analysis will be done, which optimizes the model. 
There is a function in the phylolm package that allows us to do this: phyloglm. 

### Mycorrhiza
```{r}
#fill in the arguments for the phyloglm function and execute it

```

## 8 Biological interpretation

In the end, the final models and the predictor variables that are included will have to be interpreted biologically. 
With this knowledge, the thesis can be completed. 


