---
title: "VIF_analysis_myco"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# VIF-analysis of the mycorrhiza dataset

Each of the three traits has a GIS table with different values for each variable. These variables come from 4 different datasets. Within and inbetween these datasets is a lot of correlation, which we want to eliminate as this can cause bias in our model. It depends on the trait which of these variables are the most important, or will give us the most information. This means that although the same pipeline is followed for this section, the outcomes will differ because of biological reasoning. 

## General code

The general code for the VIF analysis is quite easy, but it will be changed every step, because everytime a new value will be left out. 
```{r}
#load required package
library(usdm)

#general code for VIF analysis
vif(GIS_myco[,c(3:43)])
```

## Variables

To have an overview of the variables and more explanation about what they mean, it's best to load a file that contains that information.
```{r}
#load variable overview
project_root <- 'C:/Users/Esther/Documents/Master/Stage/'
variable_file <- paste(c(project_root, 'GIS/Variables_overview.md'), collapse = '')
variable_overview <- read.delim(variable_file)
variable_overview <- variable_overview[1:52,1:2]
```

## Hypothesis

Following the hypothesis about mycorrhiza, the following variables will most probably be important for the analysis (and thus best be not deleted): 
1. Moist related variables: climaticMoistureIndex, embergerQ, any PET/rain variable
2. Soil related variables: Bulk density, Clay percentage, pH CaCL, Organic carbon 
3. Temperature related variables: continentality, BIO8-11, BIO5&BIO6 OR maxTempColdestMonth & minTempWarmestMonth, thermInd

## Consecutive rounds of the VIF analysis 

Do consecutive rounds of the VIF analysis, while leaving out one variable each round. Guidelines:
1. In the end, the VIF value has to be lower than 10 for each variable, so variables have to be deleted until this is the case. 
2. Preferably delete the one with the highest VIF value each round, this will speed up the decrease of all the VIF values, but:  
3. Make sure that it makes sense biologically and makes sense in the context of the specific trait. Especially tread lightly around the variables we selected previously.

### VIF-analysis purely based on highest VIF value (formula 1)
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), BIO14 (38), PETseasonality (18)
vif(GIS_myco[,c(3:5, 7, 14, 20, 22:24, 26, 28, 39, 42, 43)])

#result is 5 left of the variables we wanted
```

### Round one
```{r}
#all variables
vif(GIS_myco[,c(3:43)])
```
BIO5, BIO6 and BIO7 all are infinite. 
  BIO5 = Max Temperature of Warmest Month
  BIO6 = Min Temperature of Coldest Month
  BIO7 = Temperature Annual Range (BIO5-BIO6)
BIO7 is a combination of BIO5 and BIO6, which explains the correlated VIF value. Best step would be to exclude BIO7 from the analysis.  

### Round two
```{r}
#excluding: BIO7 (31)
vif(GIS_myco[,c(3:30, 32:43)])
```
BIO11 has the highest VIF value.
  BIO11 = Mean Temperature of Coldest Quarter
It is okay to leave this variable out, if there are still other temperature variables left in the end.

### Round three
```{r}
#excluding: BIO7 (31), BIO11 (35)
vif(GIS_myco[,c(3:30, 32:34, 36:43)])
```
continentality has the highest VIF value.
  continentality = Average temp. of warmest and coldest month
It is okay to leave this variable out, if there are still other temperature variables left in the end.

### Round four
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9)
vif(GIS_myco[,c(3:8, 10:30, 32:34, 36:43)])
```
BIO10 has the highest VIF value. 
  BIO10 = Mean Temperature of Warmest Quarter.
It is okay to leave this variable out (correlates with BIO11), if there are still other temperature variables left in the end.

### Round five
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO10 (34)
vif(GIS_myco[,c(3:8, 10:30, 32, 33, 36:43)])
```
maxTempColdestMonth has the highest vIF value.
  maxTempColdestMonth = Maximum temp. of the coldest month
It is okay to leave this variable out, if there are still other temperature variables left in the end.

### Round six
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO10 (34), maxTempColdestMonth (13)
vif(GIS_myco[,c(3:8, 10:12, 14:30, 32, 33, 36:43)])
```
BIO6 has the highest VIF value.
  BIO6 = Min Temperature of Coldest Month
It is okay to leave this variable out, if there are still other temperature variables left in the end.

### Round seven
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO10 (34), maxTempColdestMonth (13), BIO6
vif(GIS_myco[,c(3:8, 10:12, 14:29, 32, 33, 36:43)])
```
BIO1 has the highest VIF value.
  BIO1 = Annual Mean Temperature
It is okay to leave this variable out, if there are still other temperature variables left in the end.

### Round eight
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO10 (34), maxTempColdestMonth (13), BIO6 (30), BIO1 (25)
vif(GIS_myco[,c(3:8, 10:12, 14:24, 26:29, 32, 33, 36:43)])
```
annualPET has the highest VIF value.
  annualPET = Annual potential evapotranspiration
It is okay to leave this variable out, if there are still other precipitation variables left in the end.

### Round nine
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO10 (34), maxTempColdestMonth (13), BIO6 (30), BIO1 (25), annualPET (6)
vif(GIS_myco[,c(3:5, 7, 8, 10:12, 14:24, 26:29, 32, 33, 36:43)])
```
growingDegDays0 has the highest VIF value.
  growingDegDays0 = Sum of months with temperatures greater than 0 degrees
It is okay to leave this variable out, is not important for this analysis.

### Round ten
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO10 (34), maxTempColdestMonth (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11)
vif(GIS_myco[,c(3:5, 7, 8, 10, 12, 14:24, 26:29, 32, 33, 36:43)])
```
PETColdestQuarter has the highest VIF value.
  PETColdestQuarter = Mean monthly PET of coldest quarter
It is okay to leave this variable out, if there are still other precipitation-related variables left in the end.

### Round eleven
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO10 (34), maxTempColdestMonth (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16)
vif(GIS_myco[,c(3:5, 7, 8, 10, 12, 14, 15, 17:24, 26:29, 32, 33, 36:43)])
```
thermicityIndex has the highest VIF value.
  thermInd = Compensated thermicity index
It is okay to leave this variable out, if there are still other temperature variables left in the end.

### Round twelve
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO10 (34), maxTempColdestMonth (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21)
vif(GIS_myco[,c(3:5, 7, 8, 10, 12, 14, 15, 17:20, 22:24, 26:29, 32, 33, 36:43)])
```
BIO13 has the highest VIF value.
  BIO13 = Precipitation of Wettest Month
It is okay to leave this variable out, if there are still other precipitation variables left in the end.

### Round thirteen
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO10 (34), maxTempColdestMonth (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37)
vif(GIS_myco[,c(3:5, 7, 8, 10, 12, 14, 15, 17:20, 22:24, 26:29, 32, 33, 36, 38:43)])
```
BIO17 has the highest VIF value.
  BIO17 = Precipitation of Driest Quarter
It is okay to leave this variable out, if there are still other precipitation variables left in the end.

### Round fourteen
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO10 (34), maxTempColdestMonth (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41)
vif(GIS_myco[,c(3:5, 7, 8, 10, 12, 14, 15, 17:20, 22:24, 26:29, 32, 33, 36, 38:40, 42, 43)])
```
BIO5 has the highest VIF value
  BIO5 = Max Temperature of Warmest Month
It is okay to leave this variable out, if there are still other temperature variables left in the end.

### Round fifteen
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO10 (34), maxTempColdestMonth (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29)
vif(GIS_myco[,c(3:5, 7, 8, 10, 12, 14, 15, 17:20, 22:24, 26:28, 32, 33, 36, 38:40, 42, 43)])
```
BIO9 has the highest VIF value.
  BIO9 = Mean Temperature of Driest Quarter
It is okay to leave this variable out, if there are still other temperature variables left in the end.

### Round sixteen
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO10 (34), maxTempColdestMonth (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33)
vif(GIS_myco[,c(3:5, 7, 8, 10, 12, 14, 15, 17:20, 22:24, 26:28, 32, 36, 38:40, 42, 43)])
```
BIO12 has the highest VIF value.
  BIO12 = Annual Precipitation
It is okay to leave this variable out, if there are still other precipitation variables left in the end.

### Round seventeen
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO10 (34), maxTempColdestMonth (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36)
vif(GIS_myco[,c(3:5, 7, 8, 10, 12, 14, 15, 17:20, 22:24, 26:28, 32, 38:40, 42, 43)])
```
BIO3 has the highest VIF value.
  BIO3 = Isothermality (BIO2/BIO7) (* 100)
It is okay to leave this variable out, if there are still other temperature variables left in the end.

### Round eightteen
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO10 (34), maxTempColdestMonth (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27)
vif(GIS_myco[,c(3:5, 7, 8, 10, 12, 14, 15, 17:20, 22:24, 26, 28, 32, 38:40, 42, 43)])
```
growingDegDays5 has the highest VIF value.
  growingDegDays5 = Sum of months with temperatures greater than 5 degrees
It is okay to leave this variable out, if there are still other temperature variables left in the end.

### Round nineteen
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO10 (34), maxTempColdestMonth (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12)
vif(GIS_myco[,c(3:5, 7, 8, 10, 14, 15, 17:20, 22:24, 26, 28, 32, 38:40, 42, 43)])
```
BIO8 has the highest VIF value.
  BIO8 = Mean Temperature of Wettest Quarter
It is okay to leave this variable out, if there are still other temperature variables left in the end.

### Round twenty
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO10 (34), maxTempColdestMonth (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32)
vif(GIS_myco[,c(3:5, 7, 8, 10, 14, 15, 17:20, 22:24, 26, 28, 38:40, 42, 43)])
```
PETWarmestQuarter has the highest VIF value.
  PETWarmestQuarter = Mean monthly PET of warmest quarter
It is okay to leave this variable out, if there are still other precipitation-related variables left in the end.

### Round twenty-one
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO10 (34), maxTempColdestMonth (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19) 
vif(GIS_myco[,c(3:5, 7, 8, 10, 14, 15, 17, 18, 20, 22:24, 26, 28, 38:40, 42, 43)])
```
BIO16 has the highest VIF value
  BIO16 = Precipitation of Wettest Quarter
It is okay to leave this variable out, if there are still other precipitation variables left in the end.

### Round twenty-two
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO10 (34), maxTempColdestMonth (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40)
vif(GIS_myco[,c(3:5, 7, 8, 10, 14, 15, 17, 18, 20, 22:24, 26, 28, 38, 39, 42, 43)])
```
monthCountByTemp10 has the highest VIF value.
  monthCountByTemp10 = Sum of months with temperatures greater than 10 degrees
It is okay to leave this variable out, if there are still other temperature variables left in the end.

### Round twenty-three
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO10 (34), maxTempColdestMonth (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15)
vif(GIS_myco[,c(3:5, 7, 8, 10, 14, 17, 18, 20, 22:24, 26, 28, 38, 39, 42, 43)])
```
embergerQ has the highest VIF value. Followed by climaticMoistureIndex. 
  embergerQ = Emberger’s pluviothermic quotient 
    Moist and temperature combined (?)
  climaticMoistureIndex = Metric of relative wetness and aridity
This variable is quite important, when looking at the hypothesis. So we'll delete climaticMoistureIndex.

### Round twenty-four
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO10 (34), maxTempColdestMonth (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), climaticMoistureIndex (8)
vif(GIS_myco[,c(3:5, 7, 10, 14, 17, 18, 20, 22:24, 26, 28, 38, 39, 42, 43)])
```
embergerQ has the highest VIF value. Followed by PETDriestQuarter
  embergerQ = Emberger’s pluviothermic quotient
  PETDriestQuarter = Mean monthly PET of driest quarter
embergerQ has a lot of information, when looking at the hypothesis. So delete PETDriestQuarter

### Round twenty-five
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO10 (34), maxTempColdestMonth (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), climaticMoistureIndex (8), PETDriestQuarter (17)
vif(GIS_myco[,c(3:5, 7, 10, 14, 18, 20, 22:24, 26, 28, 38, 39, 42, 43)])
```
embergerQ has the highest VIF value. Followed by BIO14
  embergerQ = Emberger’s pluviothermic quotient
  BIO14 = Precipitation of Driest Month
embergerQ has a lot of information, when looking at the hypothesis. So delete BIO14 (as long as there are still precipitation related variables in the end)

### Round twenty-six
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO10 (34), maxTempColdestMonth (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), climaticMoistureIndex (8), PETDriestQuarter (17), BIO14 (38)
vif(GIS_myco[,c(3:5, 7, 10, 14, 18, 20, 22:24, 26, 28, 39, 42, 43)])
```
embergerQ has the highest VIF value. Followed by PETseasonality
  embergerQ = Emberger’s pluviothermic quotient
  PETseasonality = Monthly variability in potential evapotranspiration
embergerQ has a lot of information, when looking at the hypothesis. So delete PETseasonality (as long as there are still precipitation related variables in the end). 

### Round twenty-seven
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO10 (34), maxTempColdestMonth (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), climaticMoistureIndex (8), PETDriestQuarter (17), BIO14 (38), PETseasonality (18)
vif(GIS_myco[,c(3:5, 7, 10, 14, 20, 22:24, 26, 28, 39, 42, 43)])
```
embergerQ has the highest VIF value, followed by BIO19
  embergerQ = Emberger’s pluviothermic quotient
  BIO19 = Precipitation of Coldest Quarter
In my opinion, embergerQ is important for the model, so I'll keep that variable and delete BIO19.

### Round twenty-eight: With embergerQ (formula 2)
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO10 (34), maxTempColdestMonth (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), climaticMoistureIndex (8), PETDriestQuarter (17), BIO14 (38), PETseasonality (18), BIO19 (43)
vif(GIS_myco[,c(3:5, 7, 10, 14, 20, 22:24, 26, 28, 39, 42)])
```
Result is only VIF values of below 10. 

## Other options

### 3. With climaticMoistureIndex
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), BIO14 (38), PETseasonality (18), aridityIndexThornthwaite (7), BIO18 (42)
vif(GIS_myco[,c(3:5, 8, 14, 20, 22:24, 26, 28, 39, 43)])
```

### 4. With continentality
```{r}
#excluded: BIO7 (31), BIO11(35), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), BIO14 (38), PETseasonality (18), BIO4 (28)
vif(GIS_myco[,c(3:5, 7, 9, 14, 20, 22:24, 26, 39, 42, 43)])
```

### 5. With thermicityIndex
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), BIO14 (38), PETseasonality (18), BIO4 (28), minTempWarmest (14), aridityIndexThornthwaite (7)
vif(GIS_myco[,c(3:5, 20:24, 26, 38, 39, 42, 43)])
```

### 6. With bio8
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), BIO14 (38), PETseasonality (18), PETWettestQuarter (20), minTempWarmest (14)
vif(GIS_myco[,c(3:5, 7, 22:24, 26, 28, 32, 39, 42, 43)])
```

### 7. With bio9
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), BIO14 (38), PETseasonality (18), bio4 (28)
vif(GIS_myco[,c(3:5, 7, 14, 20, 22:24, 26, 33, 39, 42, 43)])
```

### 8. With BIO8 and BIO9 
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO12 (36), BIO3 (27), growingDegDays5 (12), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), BIO14 (38), PETseasonality (18), minTempWarmest (14), PETWettestQuarter (20), BIO4 (28)
vif(GIS_myco[,c(3:5, 7, 22:24, 26, 32, 33, 39, 42, 43)])
```

### 9. With BIO8, BIO9, and EmbergerQ
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO12 (36), BIO3 (27), growingDegDays5 (12), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), PETDriestQuarter (17), climaticMoistureIndex (8), BIO14 (38), PETseasonality (18), minTempWarmest (14), PETWettestQuarter (20), BIO4 (28), aridityIndexThornthwaite (7), BIO19 (43)
vif(GIS_myco[,c(3:5, 10, 22:24, 26, 32, 33, 39, 41, 42)])
vif(GIS_myco[,c(3:5, 10, 22:24, 26, 32, 33, 38, 39, 42)])
#added 38, 41
```

### 10. With BIO8, BIO9, and climaticMoistureIndex
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO12 (36), BIO3 (27), growingDegDays5 (12), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), BIO14 (38), PETseasonality (18), minTempWarmest (14), PETWettestQuarter (20), BIO4 (28), aridityIndexThornthwaite (7), BIO18 (42), BIO2 (26)
vif(GIS_myco[,c(3:5, 8, 22:24, 32, 33, 38, 39, 43)])
vif(GIS_myco[,c(3:5, 8, 22:24, 27, 32, 33, 39, 41, 43)])
#added 27, 38, 41
vif(GIS_myco[,c(3:5, 8, 22:24, 27, 32, 33, 38, 39, 43)])
#added a third option 
```

### 11. With embergerQ and thermicityIndex
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO10 (34), maxTempColdestMonth (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), climaticMoistureIndex (8), PETDriestQuarter (17), BIO14 (38), PETseasonality (18), BIO19 (43), BIO4 (28), minTempWarmest (14), aridityIndexThornthwaite (7), PETWettestQuarter (20), OrganicCarbon (22), ClayPercentage (5)
vif(GIS_myco[,c(3, 4, 10, 21, 23, 24, 26, 38, 39, 42)])
vif(GIS_myco[,c(3, 4, 10, 20:24, 26, 39, 41, 42)])
#added 20, 22, 38, 41
vif(GIS_myco[,c(3, 4, 10, 20:24, 26, 38, 39, 42)])
#added a third option
```

### 12. With climaticMoistureIndex and thermicityIndex
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), BIO14 (38), PETseasonality (18), aridityIndexThornthwaite (7), BIO18 (42), BIO4 (28), minTempWarmest (14)
vif(GIS_myco[,c(3:5, 8, 20:24, 26, 39, 43)])
```

## Second selection

### 13. With annualPET
```{r}
#excluded: BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), PETseasonality (18), bio4 (28), aridityIndexThornthwaite (7), PETWettestQuarter (20), bio2 (26)
vif(GIS_myco[,c(3:6, 14, 18, 22:24, 31, 38, 39, 42, 43)])
#added 18, 31, 38
```

### 14. With growingDegDays0
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), BIO14 (38), PETseasonality (18), bio4 (28), minTempWarmest (14), aridityIndexThornthwaite (7)
vif(GIS_myco[,c(3:5, 11, 20, 22:24, 26, 38, 39, 42, 43)])
#added 38
```

### 15. With growingDegDays5
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), BIO14 (38), PETseasonality (18), bio4 (28), aridityIndexThornthwaite (7)
vif(GIS_myco[,c(3:5, 12, 14, 20, 22:24, 26, 38, 39, 42, 43)])
#added 38
```

### 16. With maxTempColdest
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), BIO14 (38), PETseasonality (18), bio4 (28)
vif(GIS_myco[,c(3:5, 7, 13, 14, 20, 22:24, 26, 39, 42, 43)])
```

### 17. With monthCountByTemp10
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), BIO14 (38), PETseasonality (18), minTempWarmest (14), aridityIndexThornthwaite (7), bio4 (28)
vif(GIS_myco[,c(3:5, 7, 14, 15, 20, 22:24, 26, 39, 42, 43)])
```

### 18. With PETColdestQuarter
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), BIO14 (38), PETseasonality (18), bio4 (28), aridityIndexThornthwaite (7)
vif(GIS_myco[,c(3:5, 14, 16, 20, 22:24, 26, 38, 39, 42, 43)])
#added 38
```

### 19. With PETDriestQuarter
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), climaticMoistureIndex (8), BIO14 (38), PETseasonality (18), PETWettestQuarter (20), aridityIndexThornthwaite (7)
vif(GIS_myco[,c(3:5, 14, 17, 22:24, 26, 28, 38, 39, 42, 43)])
#added 38
```

### 20. With PETseasonality
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), BIO14 (38), PETWettestQuarter (20)
vif(GIS_myco[,c(3:5, 7, 14, 18, 22:24, 26, 28, 39, 42, 43)])
```

### 21. With PETWarmestQuarter
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), BIO14 (38), PETseasonality (18), bio2 (26)
vif(GIS_myco[,c(3:5, 7, 14, 19, 20, 22:24, 28, 39, 42, 43)])
```

### 22. With bio1
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), BIO14 (38), PETseasonality (18), minTempWarmest (14), bio4 (28)
vif(GIS_myco[,c(3:5, 7, 18, 20, 22:26, 39, 42, 43)])
#added 18
```

### 23. With bio3
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), BIO14 (38), PETseasonality (18), bio4 (28)
vif(GIS_myco[,c(3:5, 7, 14, 20, 22:24, 26, 27, 39, 42, 43)])
```

### 24. With bio5 
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), BIO14 (38), PETseasonality (18), minTempWarmest (14)
vif(GIS_myco[,c(3:5, 7, 20, 22:24, 26, 28, 29, 39, 42, 43)])
```

### 25. With bio6
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), BIO14 (38), PETseasonality (18), bio4 (28)
vif(GIS_myco[,c(3:5, 7, 14, 20, 22:24, 26, 30, 39, 42, 43)])
```

### 26. With bio7
```{r}
#excluded: BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), BIO14 (38), PETseasonality (18), bio4 (28)
vif(GIS_myco[,c(3:5, 7, 14, 20, 22:24, 26, 31, 39, 42, 43)])
```

### 27. With bio10
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), BIO14 (38), PETseasonality (18), minTempWarmest (14)
vif(GIS_myco[,c(3:5, 7, 20, 22:24, 26, 28, 34, 39, 42, 43)])
```

### 28. With bio11
```{r}
#excluded: BIO7 (31), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), BIO14 (38), PETseasonality (18), bio4 (28)
vif(GIS_myco[,c(3:5, 7, 14, 20, 22:24, 26, 35, 39, 42, 43)])
```

### 29. With bio12 
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), BIO14 (38), PETseasonality (18), bio18 (42), aridityIndexThornthwaite (7), minTempWarmest (14)
vif(GIS_myco[,c(3:5, 18, 20, 22:24, 26, 28, 36, 39, 43)])
#added 18
```

### 30. With bio13
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21),  BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), BIO14 (38), PETseasonality (18), bio18 (42)
vif(GIS_myco[,c(3:5, 7, 14, 20, 22:24, 26, 28, 37, 39, 43)])
```

### 31. With bio14 
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), PETseasonality (18), aridityIndexThornthwaite (7)
vif(GIS_myco[,c(3:5, 14, 20, 22:24, 26, 28, 38, 39, 42, 43)])
```

### 32. With bio16
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO17 (41), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), BIO14 (38), PETseasonality (18), bio18 (42)
vif(GIS_myco[,c(3:5, 7, 14, 20, 22:24, 26, 28, 39, 40, 43)])
```

### 33. With bio17
```{r}
#excluded: BIO7 (31), BIO11(35), continentality (9), BIO10 (34), maxTempColdest (13), BIO6 (30), BIO1 (25), annualPET (6), growingDegDays0 (11), PETColdestQuarter (16), thermicityIndex (21), BIO13 (37), BIO5 (29), BIO9 (33), BIO12 (36), BIO3 (27), growingDegDays5 (12), BIO8 (32), PETWarmestQuarter (19), BIO16 (40), monthCountByTemp10 (15), embergerQ (10), PETDriestQuarter (17), climaticMoistureIndex (8), BIO14 (38), PETseasonality (18), aridityIndexThornthwaite (7), bio18 (42)
vif(GIS_myco[,c(3:5, 14, 20, 22:24, 26, 28, 37, 39, 41, 43)])
#added 37
```