---
title: "VIF_analysis_crops"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# VIF-analysis of the domestication dataset

Each of the three traits has a GIS table with different values for each variable. These variables come from 4 different datasets. Within and inbetween these datasets is a lot of correlation, which we want to eliminate as this can cause bias in our model. It depends on the trait which of these variables are the most important, or will give us the most information. This means that although the same pipeline is followed for this section, the outcomes will differ because of biological reasoning. 

## General code

The general code for the VIF analysis is quite easy, but it will be changed every step, because everytime a new value will be left out. 
```{r}
#load required package
library(usdm)

#general code for VIF analysis
vif(GIS_woody[,c(3:43)])
```

## Variables

To have an overview of the variables and more explanation about what they mean, it's best to load a file that contains that information.
```{r}
#load variable overview
project_root <- 'C:/Users/Esther/Documents/Master/Stage/'
variable_file <- paste(c(project_root, 'GIS/Variables_overview.md'), collapse = '')
variable_overview <- read.delim(variable_file)
variable_overview <- variable_overview[1:52,1:2]
```

## Hypothesis

Following the hypothesis about domestication, the following variables will most probably be important for the analysis (and thus best be not deleted): 
1. PETseasonality = Monthly variability in potential evapotranspiration
2. BIO2 = Mean Diurnal Range (Mean of monthly (max temp - min temp))
3. BIO4 = Temperature Seasonality (standard deviation *100)
4. BIO15 = Precipitation Seasonality (Coefficient of Variation)
5. BIO5&BIO6 = Max Temperature of Warmest month, and Min Temperature of Coldest Month
    OR: maxTempColdestMonth & minTempWarmestMonth
    OR: BIO7 = Temperature Annual Range (BIO5-BIO6)
    OR: BIO3 (if possible)

## Consecutive rounds of the VIF analysis 

Do consecutive rounds of the VIF analysis, while leaving out one variable each round. Guidelines:
1. In the end, the VIF value has to be lower than 10 for each variable, so variables have to be deleted until this is the case. 
2. Preferably delete the one with the highest VIF value each round, this will speed up the decrease of all the VIF values, but:  
3. Make sure that it makes sense biologically and makes sense in the context of the specific trait. Especially tread lightly around the variables we selected previously.

### Round one
```{r}
#all variables
vif(GIS_woody[,c(3:43)])
```


### Round two
```{r}
#excluding: 
vif(GIS_woody[,c(3:43)])
```


### Round three
```{r}
#excluding:
vif(GIS_woody[,c(3:43)]) 
```


