---
title: "VIF_analysis_crops"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# VIF-analysis of the domestication dataset

Each of the three traits has a GIS table with different values for each variable. These variables come from 4 different datasets. Within and inbetween these datasets is a lot of correlation, which we want to eliminate as this can cause bias in our model. It depends on the trait which of these variables are the most important, or will give us the most information. This means that although the same pipeline is followed for this section, the outcomes will differ because of biological reasoning. 

## General code

The general code for the VIF analysis is quite easy, but it will be changed every step, because everytime a new value will be left out. 
```{r}
#load required package
library(usdm)

#general code for VIF analysis
vif(GIS_crops[,c(3:43)])
```

## Variables

To have an overview of the variables and more explanation about what they mean, it's best to load a file that contains that information.
```{r}
#load variable overview
project_root <- 'C:/Users/Esther/Documents/Master/Stage/'
variable_file <- paste(c(project_root, 'GIS/Variables_overview.md'), collapse = '')
variable_overview <- read.delim(variable_file)
variable_overview <- variable_overview[1:52,1:2]
```

## Hypothesis

Following the hypothesis about domestication, the following variables will most probably be important for the analysis (and thus best be not deleted): 
1. PETseasonality = Monthly variability in potential evapotranspiration
2. BIO2 = Mean Diurnal Range (Mean of monthly (max temp - min temp))
3. BIO4 = Temperature Seasonality (standard deviation *100)
4. BIO15 = Precipitation Seasonality (Coefficient of Variation)
5. BIO5&BIO6 = Max Temperature of Warmest month, and Min Temperature of Coldest Month
    OR: maxTempColdestMonth & minTempWarmestMonth
    OR: BIO7 = Temperature Annual Range (BIO5-BIO6)
    OR: BIO3 (if possible)

## Consecutive rounds of the VIF analysis 

Do consecutive rounds of the VIF analysis, while leaving out one variable each round. Guidelines:
1. In the end, the VIF value has to be lower than 10 for each variable, so variables have to be deleted until this is the case. 
2. Preferably delete the one with the highest VIF value each round, this will speed up the decrease of all the VIF values, but:  
3. Make sure that it makes sense biologically and makes sense in the context of the specific trait. Especially tread lightly around the variables we selected previously.

## VIF-analysis purely based on highest VIF value
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), PETWarmestQuarter (19)
vif(GIS_crops[,c(3:5, 14, 17, 20, 22:24, 37, 38, 43)]) 
```

## VIF analysis combined with biological interpretation

### Round one
```{r}
#all variables
vif(GIS_woody[,c(3:43)])
```
BIO5, BIO6 and BIO7 all are infinite. 
  BIO5 = Max Temperature of Warmest Month
  BIO6 = Min Temperature of Coldest Month
  BIO7 = Temperature Annual Range (BIO5-BIO6)
BIO7 is a combination of BIO5 and BIO6, which explains the correlated VIF value. Best step would be to exclude BIO7 from the analysis, or delete both BOIO5 and BIO6. If I delete BIO7, I want to make sure I keep both BIO5 and BIO6

### Round two
```{r}
#excluding: BIO7 (column 31)
vif(GIS_crops[,c(3:30, 32:43)])
```
BIO11 has the highest value. 
  BIO11 = Mean Temperature of Coldest Quarter
It is okay to delete this variable, is not about the extremes/oscillations. 

### Round three
```{r}
#excluding: BIO7 (31), BIO11 (35)
vif(GIS_crops[,c(3:30, 32:34, 36:43)]) 
```
Continentality has the highest VIF value.
  continentality = Average temp. of warmest and coldest month
It is okay to delete this variable, is not about the extremes/oscillations.

### Round four
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9)
vif(GIS_crops[,c(3:8, 10:30, 32:34, 36:43)]) 
```
BIO1 has the highest VIF value
  BIO1 = Annual Mean Temperature
It is okay to delete this variable, is not about the extremes/oscillations.
  
### Round five
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25)
vif(GIS_crops[,c(3:8, 10:24, 26:30, 32:34, 36:43)]) 
```
BIO4 has the highest VIF value
   BIO4 = Temperature Seasonality (standard deviation *100)
BIO4 is a variable that could be important. We'll leave it out for now, but try to add it back later. 

### Round six
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28)
vif(GIS_crops[,c(3:8, 10:24, 26, 27, 29, 30, 32:34, 36:43)]) 
```
BIO10 has the highest VIF value
   BIO10 = Mean Temperature of Warmest Quarter
It is okay to delete this variable, is not about the extremes/oscillations.

### Round seven
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34)
vif(GIS_crops[,c(3:8, 10:24, 26, 27, 29, 30, 32, 33, 36:43)]) 
```
annualPET has the highest VIF value
   annualPET Annual potential evapotranspiration
It is okay to delete this variable, is not about the extremes/oscillations.

### Round eight
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6)
vif(GIS_crops[,c(3:5, 7, 8, 10:24, 26, 27, 29, 30, 32, 33, 36:43)]) 
```
growingDegDays0 has the highest VIF value
   growingDegDays0 = Sum of months with temperatures greater than 0 degrees
It is okay to delete this variable, is not about the extremes/oscillations.

### Round nine
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11)
vif(GIS_crops[,c(3:5, 7, 8, 10, 12:24, 26, 27, 29, 30, 32, 33, 36:43)]) 
```
maxTempColdestMonth has the highest VIF value
   maxTempColdestMonth = Maximum temp. of the coldest month
It is okay to delete this variable, is not about the extremes/oscillations.

### Round ten
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13)
vif(GIS_crops[,c(3:5, 7, 8, 10, 12, 14:24, 26, 27, 29, 30, 32, 33, 36:43)]) 
```
thermicityIndex has the highest VIF value
   thermicityIndex = Compensated thermicity index
It is okay to delete this variable, is not about the extremes/oscillations.

### Round eleven
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21)
vif(GIS_crops[,c(3:5, 7, 8, 10, 12, 14:20, 22:24, 26, 27, 29, 30, 32, 33, 36:43)]) 
```
PETColdestQuarter has the highest VIF value
   PETColdestQuarter = Mean monthly PET of coldest quarter
It is okay to delete this variable, is not about the extremes/oscillations.

### Round twelve
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16)
vif(GIS_crops[,c(3:5, 7, 8, 10, 12, 14, 15, 17:20, 22:24, 26, 27, 29, 30, 32, 33, 36:43)]) 
```
BIO17 has the highest VIF value
   BIO17 = Precipitation of Driest Quarter
It is okay to delete this variable, is not about the extremes/oscillations.

### Round thirteen
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41)
vif(GIS_crops[,c(3:5, 7, 8, 10, 12, 14, 15, 17:20, 22:24, 26, 27, 29, 30, 32, 33, 36:40, 42, 43)]) 
```
BIO12 has the highest VIF value
  BIO12 = Annual Precipitation
It is okay to delete this variable, is not about the extremes/oscillations.

### Round fourteen
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36)
vif(GIS_crops[,c(3:5, 7, 8, 10, 12, 14, 15, 17:20, 22:24, 26, 27, 29, 30, 32, 33, 37:40, 42, 43)]) 
```
BIO8 has the highest VIF value
  BIO8 = Mean Temperature of Wettest Quarter
It is okay to delete this variable, is not about the extremes/oscillations.

### Round fifteen
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32)
vif(GIS_crops[,c(3:5, 7, 8, 10, 12, 14, 15, 17:20, 22:24, 26, 27, 29, 30, 33, 37:40, 42, 43)]) 
```
BIO6 has the highest VIF value
  BIO6 = Min Temperature of Coldest Month
It is okay to delete this variable, it is not the most important temperature variable, and BIO2 is still included.

### Round sixteen
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30)
vif(GIS_crops[,c(3:5, 7, 8, 10, 12, 14, 15, 17:20, 22:24, 26, 27, 29, 33, 37:40, 42, 43)]) 
```
BIO5 has the highest VIF value
  BIO5 = Max Temperature of Warmest Month
It is okay to delete this variable, it is not the most important temperature variable, and BIO2 is still included.

### Round seventeen
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29)
vif(GIS_crops[,c(3:5, 7, 8, 10, 12, 14, 15, 17:20, 22:24, 26, 27, 33, 37:40, 42, 43)]) 
```
BIO16 has the highest VIF value
  BIO16 = Precipitation of Wettest Quarter
It is okay to delete this variable, is not about the extremes/oscillations.

### Round eightteen
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40)
vif(GIS_crops[,c(3:5, 7, 8, 10, 12, 14, 15, 17:20, 22:24, 26, 27, 33, 37:39, 42, 43)]) 
```
growingDegDays5 has the highest VIF value
  growingDegDays5 = Sum of months with temperatures greater than 5 degrees
It is okay to delete this variable, is not about the extremes/oscillations.

### Round nineteen
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12)
vif(GIS_crops[,c(3:5, 7, 8, 10, 14, 15, 17:20, 22:24, 26, 27, 33, 37:39, 42, 43)]) 
```
climaticMoistureIndex  has the highest VIF value
  climaticMoistureIndex = Metric of relative wetness and aridity
It is okay to delete this variable, is not about the extremes/oscillations.

### Round twenty
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8)
vif(GIS_crops[,c(3:5, 7, 10, 14, 15, 17:20, 22:24, 26, 27, 33, 37:39, 42, 43)]) 
```
BIO3 has the highest VIF value
  BIO3 = Isothermality (BIO2/BIO7) (* 100)
It is okay to delete this variable, is not about the extremes/oscillations.

### Round twenty-one
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27)
vif(GIS_crops[,c(3:5, 7, 10, 14, 15, 17:20, 22:24, 26, 33, 37:39, 42, 43)]) 
```
BIO9 has the highest VIF value
  BIO9 Mean Temperature of Driest Quarter
It is okay to delete this variable, is not about the extremes/oscillations.

### Round twenty-two
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33)
vif(GIS_crops[,c(3:5, 7, 10, 14, 15, 17:20, 22:24, 26, 37:39, 42, 43)]) 
```
embergerQ has the highest VIF value
  embergerQ Emberger’s pluviothermic quotient 
It is okay to delete this variable, is not about the extremes/oscillations.

### Round twenty-three
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10)
vif(GIS_crops[,c(3:5, 7, 14, 15, 17:20, 22:24, 26, 37:39, 42, 43)]) 
```
BIO18 has the highest VIF value
  BIO18 Precipitation of Warmest Quarter
It is okay to delete this variable, is not about the extremes/oscillations.

### Round twenty-four
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42)
vif(GIS_crops[,c(3:5, 7, 14, 15, 17:20, 22:24, 26, 37:39, 43)]) 
```
aridityIndexThornthwaite has the highest VIF value
  aridityIndexThornthwaite = Thornthwaite aridity index
It is okay to delete this variable, is not about the extremes/oscillations.

### Round twenty-five
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7)
vif(GIS_crops[,c(3:5, 14, 15, 17:20, 22:24, 26, 37:39, 43)]) 
```
BIO2 has the highest VIF value
  BIO2 = Mean Diurnal Range (Mean of monthly (max temp - min temp))
BIO2 is a variable that could be important. We'll leave it out for now, but try to add it back later. 

### Round twenty-six
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26)
vif(GIS_crops[,c(3:5, 14, 15, 17:20, 22:24, 37:39, 43)]) 
```
PETseasonality and monthCountByTemp10 have the highest VIF values
  PETseasonality = Monthly variability in potential evapotranspiration
  monthCountByTemp10 = Sum of months with temperatures greater than 10 degrees
PETseasonality could be important for this research question, as it implies oscillations. monthCountByTemp10 is less important. 

### Round twenty-seven
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), monthCountByTemp10 (15)
vif(GIS_crops[,c(3:5, 14, 17:20, 22:24, 37:39, 43)]) 
```
PETseasonality and PETDriestQuarter have the highest VIF values
  PETseasonality = Monthly variability in potential evapotranspiration
  PETDriestQuarter = Mean monthly PET of driest quarter
PETseasonality could be important for this research question, as it implies oscillations. PETDriestQuarter is less important. 

### Round twenty-eight
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), monthCountByTemp10 (15), PETDriestQuarter (17)
vif(GIS_crops[,c(3:5, 14, 18:20, 22:24, 37:39, 43)]) 
```
PETseasonality, BIO13 and BIO15 have the highest VIF values
  PETseasonality = Monthly variability in potential evapotranspiration
  BIO13 = Precipitation of Wettest Month
  BIO15 = Precipitation Seasonality (Coefficient of Variation)
PETseasonality and BIO15 could be important for this research question, as they imply oscillations. BIO13 is less important. 

### Round twenty-nine
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), monthCountByTemp10 (15), PETDriestQuarter (17), BIO13 (37)
vif(GIS_crops[,c(3:5, 14, 18:20, 22:24, 38, 39, 43)]) 
```
PETWarmestQuarter and BIO15 have the highest VIF values
  PETWarmestQuarter = Mean monthly PET of warmest quarter
  BIO15 = Precipitation Seasonality (Coefficient of Variation)
BIO15 could be important for this research question, as it implies oscillations. PETWarmestQuarter is less important. 

### Round thirty: With PETseasonality
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), monthCountByTemp10 (15), PETDriestQuarter (17), BIO13 (37), PETWarmestQuarter (19)
vif(GIS_crops[,c(3:5, 14, 18, 20, 22:24, 38, 39, 43)]) 
```
All VIF values are now beneath 10.  

## VIF analysis, other combinations:

### With BIO4 (and BIO15)
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), PETWarmestQuarter (19), BIO13 (37), PETDriestQuarter (17), BIO14 (38)
vif(GIS_crops[,c(3:5, 14, 20, 22:24, 28, 39, 43)]) 
```

### With BIO2
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), PETseasonality (18), monthCountByTemp10 (15), PETWarmestQuarter (19)
vif(GIS_crops[,c(3:5, 14, 17, 20, 22:24, 26, 37, 38, 43)]) 
```

### With BIO15
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), PETWarmestQuarter (19), BIO19 (43), BIO14 (38)
vif(GIS_crops[,c(3:5, 14, 17, 20, 22:24, 37, 39)]) 
```

### With BIO2 and PET seasonality
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), monthCountByTemp10 (15), PETDriestQuarter (17), BIO13 (37), PETWarmestQuarter (19), BIO15 (39)
vif(GIS_crops[,c(3:5, 14, 18, 20, 22:24, 26, 38, 43)]) 
```

### With BIO4 and PET seasonality
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), monthCountByTemp10 (15), PETDriestQuarter (17), BIO13 (37), PETWarmestQuarter (19), BIO14 (38), BIO15 (39), PhCaCL (23), minTempWarmest (14), BulkDensity (4)
vif(GIS_crops[,c(3, 5, 18, 20, 22, 24, 28, 43)]) 
```

### With BIO2 and BIO4 
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), PETWarmestQuarter (19), PETDriestQuarter (17), BIO13 (37)
vif(GIS_crops[,c(3:5, 14, 20, 22:24, 26, 28, 38, 43)]) 
```

### With BIO2 and BIO15
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), PETseasonality (18), monthCountByTemp10 (15), PETWarmestQuarter (19), BIO19 (43), BIO13 (37), BIO14 (38)
vif(GIS_crops[,c(3:5, 14, 17, 20, 22:24, 26, 39)]) 
```

### With BIO2, BIO4, and BIO15 
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), PETseasonality (18), monthCountByTemp10 (15), PETWarmestQuarter (19), PETDriestQuarter (17), BIO13 (37), BIO14 (38)
vif(GIS_crops[,c(3:5, 14, 20, 22:24, 26, 28, 39, 43)]) 
```

### With BIO2, BIO4 and PETseasonality
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), monthCountByTemp10 (15), PETDriestQuarter (17), BIO13 (37), PETWarmestQuarter (19), BIO15 (39), PETWettestQuarter (20), BIO14 (38), PhCaCL (23), minTempWarmest (14), BulkDensity (4), BIO19 (43), ClayPercentage (5), OrganicCarbon (22), Slope (24), Aspect (3)
vif(GIS_crops[,c(18, 26, 28)]) 
```
NOT POSSIBLE!! 

### With BIO2, BIO4 and BIO15 
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), PETseasonality (18), monthCountByTemp10 (15), PETWarmestQuarter (19), PETDriestQuarter (17), BIO13 (37), BIO14 (38)
vif(GIS_crops[,c(3:5, 14, 20, 22:24, 26, 28, 39, 43)]) 
```

## Second try 

This yielded models with a lot of soil variables, while we expected other variables to have an effect as well. So we're going to do more VIF analysis, while pushing in other variables. 

### With annualPET (6)
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), PETWarmestQuarter (19), PETDriestQuarter (17)
vif(GIS_crops[,c(3:5, 6, 14, 20, 22:24, 37, 38, 43)]) 
```

### aridityIndexThornthwaite 
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), PETWarmestQuarter (19), bio14 (38), PETDriestQuarter (17)
vif(GIS_crops[,c(3:5, 7, 14, 20, 22:24, 37, 43)]) 
```

### climaticMoistureIndex 
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), PETWarmestQuarter (19), bio13 (37)
vif(GIS_crops[,c(3:5, 8, 14, 17, 20, 22:24, 38, 43)]) 
```

### continentality 
```{r}
#excluding: BIO7 (31), BIO11 (35), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), PETWarmestQuarter (19), PETDriestQuarter (17)
vif(GIS_crops[,c(3:5, 9, 14, 20, 22:24, 37, 38, 43)]) 
```

### embergerQ 
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), BIO18 (42), aridityIndexThornth, bio19 (43)waite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), PETWarmestQuarter (19), bio13 (37), bio14 (38)
vif(GIS_crops[,c(3:5, 10, 14, 17, 20, 22:24)]) 
```

### growingDegDays0 
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), PETWarmestQuarter (19), PETDriestQuarter (17), bio13 (37)
vif(GIS_crops[,c(3:5, 11, 14, 20, 22:24, 38, 43)]) 
```

### maxTempColdestMonth 
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), PETWarmestQuarter (19), PETDriestQuarter (17)
vif(GIS_crops[,c(3:5, 13, 14, 20, 22:24, 37, 38, 43)]) 
```

### PETColdestQuarter 
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), PETWarmestQuarter (19), PETDriestQuarter (17)
vif(GIS_crops[,c(3:5, 14, 16, 20, 22:24, 37, 38, 43)]) 
```

### PETWarmestQuarter 
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), bio19 (43), bio13 (37), BulkDensity (4)
vif(GIS_crops[,c(3, 5, 14, 17, 19, 20, 22:24, 38)]) 
```

### thermicityIndex
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), PETWarmestQuarter (19), PETDriestQuarter (17), bio13 (37)
vif(GIS_crops[,c(3:5, 14, 20, 21, 22:24, 38, 43)]) 
```

### bio1 
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), PETWarmestQuarter (19), PETDriestQuarter (17), bio13 (37)
vif(GIS_crops[,c(3:5, 14, 20, 22:24, 25, 38, 43)]) 
```

### bio3 
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), PETWarmestQuarter (19), PETDriestQuarter (17)
vif(GIS_crops[,c(3:5, 14, 20, 22:24, 27, 37, 38, 43)]) 
```

### bio5
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), PETWarmestQuarter (19), minTempWarmest (14)
vif(GIS_crops[,c(3:5, 17, 20, 22:24, 29, 37, 38, 43)]) 
```

### bio6
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), PETWarmestQuarter (19), PETDriestQuarter (17)
vif(GIS_crops[,c(3:5, 14, 20, 22:24, 30, 37, 38, 43)]) 
```

### bio7
```{r}
#excluding: BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), PETWarmestQuarter (19), bio13 (37)
vif(GIS_crops[,c(3:5, 14, 17, 20, 22:24, 31, 38, 43)]) 
```

### bio8
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), PETWarmestQuarter (19), PETWettestQuarter (20)
vif(GIS_crops[,c(3:5, 14, 17, 22:24, 32, 37, 38, 43)]) 
```

### bio9
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), PETWarmestQuarter (19), PETDriestQuarter (17)
vif(GIS_crops[,c(3:5, 14, 20, 22:24, 33, 37, 38, 43)]) 
```

### bio10
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), PETWarmestQuarter (19), minTempWarmest (14)
vif(GIS_crops[,c(3:5, 17, 20, 22:24, 34, 37, 38, 43)]) 
```

### bio11
```{r}
#excluding: BIO7 (31), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), PETWarmestQuarter (19), PETDriestQuarter (17)
vif(GIS_crops[,c(3:5, 14, 20, 22:24, 35, 37, 38, 43)]) 
```

### bio12
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), PETWarmestQuarter (19), bio13 (37), bio14 (38)
vif(GIS_crops[,c(3:5, 14, 17, 20, 22:24, 36, 43)]) 
```

### bio16
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), PETWarmestQuarter (19), bio13 (37)
vif(GIS_crops[,c(3:5, 14, 17, 20, 22:24, 38, 40, 43)]) 
```

### bio17
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), BIO18 (42), aridityIndexThornthwaite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), PETWarmestQuarter (19), bio14 (38), bio19 (43)
vif(GIS_crops[,c(3:5, 14, 17, 20, 22:24, 37, 41)]) 
```

### bio18
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), aridityIndexThornthwaite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), PETWarmestQuarter (19), bio19 (43), bio13 (37)
vif(GIS_crops[,c(3:5, 14, 17, 20, 22:24, 38, 42)]) 
```

### With things
```{r}
#excluding: BIO7 (31), BIO11 (35), continentality (9), BIO1 (25), BIO4 (28), BIO10 (34), annualPET (6), growingDegDays0 (11), maxTempColdestMonth (13), thermicityIndex (21), PETColdestQuarter (16), BIO17 (41), BIO12 (36), BIO8 (32), BIO6 (30), BIO5 (29), BIO16 (40), growingDegDays5 (12), climaticMoistureIndex (8), BIO3 (27), BIO9 (33), embergerQ (10), aridityIndexThornthwaite (7), BIO2 (26), PETseasonality (18), monthCountByTemp10 (15), BIO15 (39), PETWarmestQuarter (19), bio19 (43), bio13 (37)
vif(GIS_crops[,c(3:5, 14, 20, 22:24, 26, 28, 39, 40, 43)]) 
```
